### 2020-11-10 paper/paper_panarchy.Rmd
first commit: setting up the project

---
title: "Panarchy: ripples of a boundary concept"
author: |
  | \small \textbf{Juan C. Rocha}
  | \small Stockholm Resilience Centre
  | \small Stockholm University
  | \small \texttt{\href{mailto:juan.rocha@su.se}{\nolinkurl{juan.rocha@su.se}}}

abstract: |
  Ecosystems around the world are at riks of critical transitions due to increasing anthropogenic preasures and climate change [@Bolt:2018ib; @Rocha:2018gn; @Rockstrom:2009p4681; @Steffen:2018ku]. However, it is not clear where this risks are higher or where ecosystems are more vulnerable. Here I measure resilience of primary production proxies for marine and terrestrial ecosystems globally.

bibliography: references.bib
csl: nature.csl
output:
  pdf_document:
    dev: pdf
    keep_tex: no
    toc: no
    latex_engine: xelatex
  word_document:
    toc: no
citecolor: blue
documentclass: article
font: Arial
fontsize: 10pt
keywords:
- cooperation
- thresholds
- risk
- uncertainty
- regime shifts
linkcolor: blue
urlcolor: blue
header-includes:
- \usepackage{dcolumn, rotating, longtable, lineno, float, fontspec, array, tabularx}
- \setmainfont{Arial}
- \linenumbers
- \renewcommand{\abstractname}{}
- \graphicspath{{figures/}}
- \usepackage[margin=2.5cm]{geometry}
---

# Outline
Intro

# Introduction
(800w)

# Methods
(800w)

# Results
(1000w)

# Discussion
(1000w)
- Prospects of applying ML
- False negatives and false positives.
- Signals of critical transtions that are strong enought to impact primary productivity. The data does not allow to detect higher resolution changes in the ecological hierarchy such as regime shifts in populations or communities unless they impact aggregated primary productivy as observed from remote sensing.
- Many regime shift types are not detected: e.g. mangroves, salt marshes, mountain glaciers, etc. Higher resolution data might help, but
- updates: ecological resilience observatory.

# Conclusion
(400w)

# References {-}
\bibliography{references}

\pagebreak

# Supplementary Material

### 2020-11-11 paper/paper_panarchy.Rmd
first draft intro and methods

@@
 
-bibliography: references.bib
+bibliography: panarchy.bib
 csl: nature.csl
@@
 fontsize: 10pt
-keywords:
-- cooperation
-- thresholds
-- risk
-- uncertainty
-- regime shifts
+
 linkcolor: blue
@@
 Intro
+-  What is Panarchy and how it has influenced the scholarly of sustainability scientist in the last 2 decades?
+    - Key features: adaptive cycles nested in scales, poverty and rigidity traps.
+    - How have these concepts developed in the last 2 decades?
+    - What are the key frontiers of research?
+
+- Method
+    - Literature review (coded by humans)
+    - topic modelling on WoS searches and all papers that have been cited Holling's summary
+- Results
 
@@
 (800w)
+Panarchy encapsulates a set of concepts that have inspired the work of environmental scientist and practitioners for almost two decades.
+It is the product of a group of scientists interested in solving environmental issues by integrating insights from the social and natural sciences, as well as the arts and humanities.
+Panarchy was, and still is, a boundary object that has enable collaborations and nurture new scientific frameworks.
+The ideas put forward have found applications in field studies, mathematical models, participatory work, and scenario development.
+It has inspired resilience assessments and guided decision making.
+In this article we pay tribute to the book by studying how the concepts and metaphors proposed have been further developed in the academic literature.
+We also document the criticism to the concepts and identify key research frontiers.
+
+Panarchy is not a theory of what it is, but a metaphor of what might be [@Gunderson:2002vk].
+It is not a predictive tool, but aims to understand adaptive change.
+Panarchy posits that systems are organised in nested hierarchies across space and time.
+Each level of the hierarchy is a subsystem that can be in different phases of adaptive cycles.
+These phases are exploitation, conservation, release, and reorganization.
+They are inspired in mathematical models used in economics and ecology, but broadly describe patterns of growth, collapse, and recovery that are common to populations, ecological communities, markets, or political organizations.
+While each subsystem in scale can be at different phases of the cycle, they can influence each other through cross-scale interactions namely revolt and remember.
+
+The adaptive cycle is hypothesized to exist in a three-dimensional space that integrates three key systemic features: its potential, connectivity, and resilience [@Gunderson:2002vk].
+The potential refers to the available capital to the system, which can be approximated as nutrients and carbon captured by a diverse forest, or the human capital of skills and knowledge accumulated to run an economy.
+Connectivity is a proxy of the structure of the system, it is the network of interactions and strength between its elements.
+Resilience is the capacity of any system to absorb disturbance and reorganise while undergoing change so as to still retain essentially the same function, structure, feedbacks, and therefore identity [@Anonymous:2016fv].
+
+The three-dimensional space has corners with attractors that can stagnate the adaptive cycle: poverty and rigidity traps [@Gunderson:2002vk].
+In panarchy, poverty traps are described as maladptive states where potential, connectivity and resilience is low.
+Poverty traps are a series of feedback mechanisms that reinforce inpoverish states [@Bowles:2006hx], limiting the capacity to innovate and increase potential.
+The opposite corner where potential, connectivity and resilience are high are also a maladaptive space denoted as rigidity traps.
+In that space there is little space for experimentation and innovation such a very conservative and bureaucratic systems.
+
+Panarchy offers a rich conceptual framework to think of environmental problems.
+While inspired in a few mathematical constructs (e.g.
+cycles, traps, scaling laws), it is general enough to invite scholars from multiple disciplinary backgrounds, ontologies and epistemologies, to collaborate around research questions and applied problems.
+It is a boundary object.
+The book itself presents a series of case studies where geographers, economists, political scientist, and ecologist have found the framework useful.
+However, to what extent have these concepts gained empirical support beyond the metaphor, how these concepts have evolved, and are we at a stage where we can start deriving theories remind open questions.
+For example, the adaptive cycle is a useful metaphor to look back at history and organize events and periods around phases.
+It reminds challenging however to narrow down in what phase is a system in the present, or what would be its expected trajectory in the future.
+In other words, it seems useufl in retrospective studies, but its application more limited for prospective ones.
+
+Here we review the academic literature of the last two decades with the aim of tracing how these concepts have evolved, in what type of problems are they found useful, and what remain key frontiers of research.
+
+[Figure 1 -- Panarchy conceptual figure? ]
 
@@
 
+To answer these questions we combined an automated literature review based on topic modelling with human-coded document analysis.
+
+*Data*: We used the Web of Science, Scopus, and GoogleScholar to survey academic literature that has used or referenced works that trace back to Panarchy [@Gunderson:2002vk].
+We extracted complete records from the Web of Science that matched the search for "(panarchy OR adaptive cycle) AND resilience" (N=XX).
+The data was combined with all papers (N = 1379) that cited the inaugural paper that introduced the book to the academic community [Holling].
+Records with missing abstracts were dropped (N=XX), and records with missing year were set to 2020 given that they are accepted manuscripts with digital object identifier (doi) scheduled to be published later in 2020 or 2021.
+
+To prepare the data for topic modelling, we constructed a document term matrix with documents in rows and words in columns.
+Here our unit of analysis for the document are the abstracts retrieved, and the matrix contain the count of words per document.
+We removed stop words (e.g.
+"the", "a") and digits from the matrix, as well as a list of words that were over represented in our data (e.g.
+"paper", "study", "aim") that are common in the scientific literature but do not inform us about the topics the papers are about.
+
+*Topic models*: are an unsupervised statistical technique to reduce the dimensionality of a corpus of data (typically but not necessarily text) into topics [@Blai].
+Here a topic is a latent variable that ranks words with high probability of appearing together within the same document.
+Documents in turn can be described by the probability that is is described by a particular set of topics.
+Since they are (posterior) probabilities, the sum of the probability of all words for any given topic should be one, and the sum of the probability of all topics for any given document should also be one.
+An iterative process or algorithm is what allows the model to learn the ranking of words that best explain topics, and the ranking of topics that best explain documents.
+
+The underlying statistical technique for this machine learning approach is called Latent Dirichlet Allocation (LDA) [@Blei].
+It allocates probabilites to latent variables (topics) based on the distribution of words in text data, assuming a multivariate continuous (Dirichlet) distribution.
+We compared three LDA algorithms: correlated topic models (CTM), variational expectation maximization (VEM), and Gibbs sampling (Gibbs), by assessing their performance against their log-likelihood estimation, entropy and perplexity [@tmodels].
+These metrics enable us to choose which algorithm best fit our data, what is the optimal number of topics to fit, and how to avoid overfitting.
+
+*Document analysis*: We complemented our unsupervised approach with document analysis [] by coding an additional set of categorical variables for a subsample of papers (N = XX).
+We annotated qualitative aspects such as use of the adaptive cycle, identification of its phases, whether the paper is conceptual, modelling or empirical; or the temporal and spatial scales of the case study.
+We also identified what aspects of panarchy were most used in the papers.
+For example whether there is emphasis on cross-scale interactions, or poverty and rigidity traps.
+Text annotations were used to capture potential criticisms as well as the methods used.
+The code book is available in the supplementary material (archived in XX repository).
+
 # Results
@@
 
+Panarchy, the book published in 2002, has been cited over 7200 times according to Google Scholar.
+The scientific paper that introduced the book to the scientific community [@HOlling] has received 1715 citations in the Web of Science, and 1923 in Scopus.
+Roughly half of the citations comes from environmental (28%) and social sciences (22%), while the less represented fields are computer science (2.1%), arts and humanities (2.6%).
+
 # Discussion
 (1000w)
-- Prospects of applying ML
-- False negatives and false positives.
-- Signals of critical transtions that are strong enought to impact primary productivity.
-The data does not allow to detect higher resolution changes in the ecological hierarchy such as regime shifts in populations or communities unless they impact aggregated primary productivy as observed from remote sensing.
-- Many regime shift types are not detected: e.g. mangroves, salt marshes, mountain glaciers, etc. Higher resolution data might help, but
-- updates: ecological resilience observatory.
 

### 2020-11-23 paper/paper_panarchy.Rmd
first draft of the paper + scripts for figures and analysis

@@
 author: |
-  | \small \textbf{Juan C. Rocha}
-  | \small Stockholm Resilience Centre
-  | \small Stockholm University
-  | \small \texttt{\href{mailto:juan.rocha@su.se}{\nolinkurl{juan.rocha@su.se}}}
+    | \small Juan Rocha^1,2,3^, Linda Luvuno^4^, Jesse Rieb^5^, Erin Crockett^6^,  Katja Malmborg^1^, Michael Schoon^7^, Garry Peterson^1^
+    | \small
+    | \small ^1^Stockholm Resilience Centre, Stockholm University, Kräftriket 2B, 10691 Stockholm
+    | \small ^2^Future Earth, The Swedish Royal Academy of Sciences, Lilla Frescativägen 4A, 104 05 Stockholm
+    | \small ^3^South American Institute for Resilience and Sustainability Studies, Maldonado, Uruguay
+    | \small ^4^...
+    | \small ^5^...
+    | \small ^6^...
+    | \small ^7^...
 
 abstract: |
-  Ecosystems around the world are at riks of critical transitions due to increasing anthropogenic preasures and climate change [@Bolt:2018ib; @Rocha:2018gn; @Rockstrom:2009p4681; @Steffen:2018ku].
-However, it is not clear where this risks are higher or where ecosystems are more vulnerable.
-Here I measure resilience of primary production proxies for marine and terrestrial ecosystems globally.
+  Panarchy has marked the academic career of many scholars and practitioners facing environmental problems.
+Initially introduced as a conceptual framework and set of metaphors, panarchy has gained the attention of scholars across many disciplines and its ideas continue to inspire further conceptual developments.
+To what extent these concepts have developed, theorized and empirically grounded are questions we address by reviewing the literature with qualitative methods and machine learning.
+We find that the adaptive cycle is the feature of panarchy that has attracted more attention.
+Challenges remain in empirically grounding the metaphor, but recent theoretical and empirical work offer some avenues for future research.
 
 bibliography: panarchy.bib
-csl: nature.csl
+# csl: nature.csl
 output:
@@
 
-# Outline
-Intro
--  What is Panarchy and how it has influenced the scholarly of sustainability scientist in the last 2 decades?
-    - Key features: adaptive cycles nested in scales, poverty and rigidity traps.
-    - How have these concepts developed in the last 2 decades?
-    - What are the key frontiers of research?
+# Introduction
 
-- Method
-    - Literature review (coded by humans)
-    - topic modelling on WoS searches and all papers that have been cited Holling's summary
-- Results
-
-# Introduction
-(800w)
-Panarchy encapsulates a set of concepts that have inspired the work of environmental scientist and practitioners for almost two decades.
+Panarchy encapsulates a set of concepts that have inspired the work of environmental scientists and practitioners for almost two decades.
 It is the product of a group of scientists interested in solving environmental issues by integrating insights from the social and natural sciences, as well as the arts and humanities.
-Panarchy was, and still is, a boundary object that has enable collaborations and nurture new scientific frameworks.
+Panarchy was, and still is, a boundary object that has enabled collaborations and nurtured new scientific frameworks.
 The ideas put forward have found applications in field studies, mathematical models, participatory work, and scenario development.
@@
 It is not a predictive tool, but aims to understand adaptive change.
-Panarchy posits that systems are organised in nested hierarchies across space and time.
+Panarchy posits that systems are organized in nested hierarchies across space and time.
 Each level of the hierarchy is a subsystem that can be in different phases of adaptive cycles.
-These phases are exploitation, conservation, release, and reorganization.
+These phases are exploitation, conservation, release, and reorganization, the latter characterized by events of creative destruction (Fig 1).
 They are inspired in mathematical models used in economics and ecology, but broadly describe patterns of growth, collapse, and recovery that are common to populations, ecological communities, markets, or political organizations.
-While each subsystem in scale can be at different phases of the cycle, they can influence each other through cross-scale interactions namely revolt and remember.
+While each subsystem in scale can be at different phases of the cycle, they can influence each other through cross-scale interactions called revolt and remember.
 
 The adaptive cycle is hypothesized to exist in a three-dimensional space that integrates three key systemic features: its potential, connectivity, and resilience [@Gunderson:2002vk].
-The potential refers to the available capital to the system, which can be approximated as nutrients and carbon captured by a diverse forest, or the human capital of skills and knowledge accumulated to run an economy.
+The potential refers to the available capital to the system, which can be approximated as nutrients and carbon captured by a forest, or the human capital of skills and knowledge accumulated to run the economy.
 Connectivity is a proxy of the structure of the system, it is the network of interactions and strength between its elements.
-Resilience is the capacity of any system to absorb disturbance and reorganise while undergoing change so as to still retain essentially the same function, structure, feedbacks, and therefore identity [@Anonymous:2016fv].
+Resilience is the capacity of any system to absorb disturbance and reorganize while undergoing change so as to still retain essentially the same function, structure, feedbacks, and therefore identity [@Anonymous:2016fv].
 
 The three-dimensional space has corners with attractors that can stagnate the adaptive cycle: poverty and rigidity traps [@Gunderson:2002vk].
-In panarchy, poverty traps are described as maladptive states where potential, connectivity and resilience is low.
-Poverty traps are a series of feedback mechanisms that reinforce inpoverish states [@Bowles:2006hx], limiting the capacity to innovate and increase potential.
-The opposite corner where potential, connectivity and resilience are high are also a maladaptive space denoted as rigidity traps.
-In that space there is little space for experimentation and innovation such a very conservative and bureaucratic systems.
+In panarchy, poverty traps are described as maladaptive states where potential, connectivity and resilience are low.
+Poverty traps are a series of feedback mechanisms that reinforce impoverished states [@Bowles:2006hx], limiting the capacity to innovate and increase potential.
+The opposite corner, where potential, connectivity and resilience are high, is also a maladaptive space denoted rigidity traps.
+In that corner there is little space for experimentation and innovation.
+Examples include very conservative and bureaucratic systems.
 
@@
 It is a boundary object.
-The book itself presents a series of case studies where geographers, economists, political scientist, and ecologist have found the framework useful.
-However, to what extent have these concepts gained empirical support beyond the metaphor, how these concepts have evolved, and are we at a stage where we can start deriving theories remind open questions.
+The book itself presents a series of case studies where geographers, economists, political scientists, and ecologists have found the framework useful.
+However, to what extent have these concepts gained empirical support beyond the metaphor, or how these concepts have evolved, remind open questions.
 For example, the adaptive cycle is a useful metaphor to look back at history and organize events and periods around phases.
-It reminds challenging however to narrow down in what phase is a system in the present, or what would be its expected trajectory in the future.
-In other words, it seems useufl in retrospective studies, but its application more limited for prospective ones.
+It remains challenging however to narrow down in what phase is a system in the present, or what would be its expected trajectory in the future.
+In other words, it seems useful in retrospective studies, but its application is more limited for prospective ones.
+Are we at a stage where we can start distilling theory that enables case study comparisons, deriving functional forms, or draw theoretical expectations and predictions?
 
@@
 
-[Figure 1 -- Panarchy conceptual figure? ]
+\begin{figure*}[h]
+\centering
+\includegraphics[width = 3in, height = 3in]{fig1_panarchy}
+\caption{\textbf{Panarchy} is an heuristic of nested adaptive cycles that serves to represent a variety of systems and environmental problems.
+Adaptive cycles at different scales of the hierarchy can be connected through remember and revolt cross-scale interactions.
+Reprinted with permission from Island Press}
+\label{fig:fig1}
+\end{figure*}
 
 # Methods
-(800w)
 
-To answer these questions we combined an automated literature review based on topic modelling with human-coded document analysis.
+To answer these questions we combined an automated literature review based on topic modeling [@Blei:2012dk; @Griffiths:2004ey] with human-coded document analysis [@Bryman:2008va].
+All data and code to replicate the analysis is available at: [https://github.com/juanrocha/panarchy](https://github.com/juanrocha/panarchy) and a public online repository (add figshare link when finished).
 
 *Data*: We used the Web of Science, Scopus, and GoogleScholar to survey academic literature that has used or referenced works that trace back to Panarchy [@Gunderson:2002vk].
-We extracted complete records from the Web of Science that matched the search for "(panarchy OR adaptive cycle) AND resilience" (N=XX).
-The data was combined with all papers (N = 1379) that cited the inaugural paper that introduced the book to the academic community [Holling].
-Records with missing abstracts were dropped (N=XX), and records with missing year were set to 2020 given that they are accepted manuscripts with digital object identifier (doi) scheduled to be published later in 2020 or 2021.
+We extracted complete records from the Scopus database that matched the search for "(panarchy OR adaptive cycle)" (N = 595), or the search "(panarchy OR adaptive cycle) AND resilience" (N = 278).
+The data was combined with all papers (N = 1923) that cited the inaugural paper that introduced the book to the academic community [@Holling:2001cu].
+Records with missing abstracts were dropped (N = 191), and records with missing year were set to 2020 given that they are accepted manuscripts with digital object identifier (doi) scheduled to be published later in 2020 or 2021.
 
-To prepare the data for topic modelling, we constructed a document term matrix with documents in rows and words in columns.
-Here our unit of analysis for the document are the abstracts retrieved, and the matrix contain the count of words per document.
+To prepare the data for topic modeling, we constructed a document term matrix with documents in rows and words in columns.
+Here our unit of analysis for the document are the abstracts retrieved, and the matrix contains the count of words per document.
 We removed stop words (e.g.
@@
 
-*Topic models*: are an unsupervised statistical technique to reduce the dimensionality of a corpus of data (typically but not necessarily text) into topics [@Blai].
+*Topic models*: are an unsupervised statistical technique to reduce the dimensionality of a corpus of data (typically but not necessarily text) into topics [@Blei:2012dk].
 Here a topic is a latent variable that ranks words with high probability of appearing together within the same document.
-Documents in turn can be described by the probability that is is described by a particular set of topics.
+Documents in turn can be described by the probability distribution of a particular set of topics.
 Since they are (posterior) probabilities, the sum of the probability of all words for any given topic should be one, and the sum of the probability of all topics for any given document should also be one.
@@
 
-The underlying statistical technique for this machine learning approach is called Latent Dirichlet Allocation (LDA) [@Blei].
-It allocates probabilites to latent variables (topics) based on the distribution of words in text data, assuming a multivariate continuous (Dirichlet) distribution.
-We compared three LDA algorithms: correlated topic models (CTM), variational expectation maximization (VEM), and Gibbs sampling (Gibbs), by assessing their performance against their log-likelihood estimation, entropy and perplexity [@tmodels].
-These metrics enable us to choose which algorithm best fit our data, what is the optimal number of topics to fit, and how to avoid overfitting.
+The underlying statistical technique for this machine learning approach is called Latent Dirichlet Allocation (LDA) [@Blei:2003tn].
+It allocates probabilities to latent variables (topics) based on the distribution of words in text data, assuming a multivariate continuous (Dirichlet) distribution.
+We compared three LDA algorithms: correlated topic models (CTM), variational expectation maximization (VEM), and Gibbs sampling (Gibbs), by assessing their performance against their log-likelihood estimation, entropy and perplexity [@Grun:2011tb].
+These metrics enable us to choose which algorithm best fits our data, what is the optimal number of topics to fit, and how to avoid overfitting.
 
-*Document analysis*: We complemented our unsupervised approach with document analysis [] by coding an additional set of categorical variables for a subsample of papers (N = XX).
-We annotated qualitative aspects such as use of the adaptive cycle, identification of its phases, whether the paper is conceptual, modelling or empirical; or the temporal and spatial scales of the case study.
+*Document analysis*: We complemented our unsupervised approach with document analysis [@Bryman:2008va] by coding an additional set of categorical variables for a sample of papers (N = 41).
+We annotated qualitative aspects such as use of the adaptive cycle, identification of its phases, whether the paper is conceptual, modeling or an empirical study.
+When empirical, we record the temporal and spatial scales of the case study.
 We also identified what aspects of panarchy were most used in the papers.
@@
 # Results
-(1000w)
 
 Panarchy, the book published in 2002, has been cited over 7200 times according to Google Scholar.
-The scientific paper that introduced the book to the scientific community [@HOlling] has received 1715 citations in the Web of Science, and 1923 in Scopus.
+The scientific paper that introduced the book to the scientific community [@Holling:2001cu] has received 1715 citations in the Web of Science, and 1923 in Scopus.
 Roughly half of the citations comes from environmental (28%) and social sciences (22%), while the less represented fields are computer science (2.1%), arts and humanities (2.6%).
+
+A first exploration on the data suggest that content related to the adaptive cycle or general resilience has slightly decreased, while issues on development and management had gained attention (Fig 2).
+This simple exploration is not necessarily the best fit for our data, but serves to exemplify how topic models work.
+A topic is a set of words that are ranked according to a probability that they represent an underlying content of the document (Fig 1A).
+For example, the words "resilience", "systems", and "social" have a high probability to capture the content of topic 5.
+Similarly, there is a cluster of papers whose content is largely dominated by topic 1 on the adaptive cycle (gamma > 0.6), featuring it as one of the most active areas of research despite its relative decline in recent years (Fig 2).
+
+\begin{figure*}[h]
+\centering
+\includegraphics[width = 5in, height = 3in]{fig2_five_topics_time}
+\caption{\textbf{A simple model with 5 topics} Topics are latent variables that rank words according to the probability of the word explaining the topic (A).
+Documents are composed by topics in different proportions which allows the identification of clusters of documents around the same topics, or how the topic content change over time (B).
+Each dot in the lower pane is an abstract of a paper (N = 2177) organized by the year it was published against each of the topic content.}
+\label{fig:fig2}
+\end{figure*}
+
+Comparing algorithm performance reveals that Gibbs sampling better fits the data.
+As a rule of thumb, an ideal method should maximize entropy and the log-likelihood estimation while minimizing perplexity [@Grun:2011tb].
+Entropy is a measure of order or disorder of a system, in the context of topic models it measures how evenly spread is the topic distribution.
+Perplexity measures the uncertainty of predicting a single word, so if the model performance were the same as random, perplexity will approximate the vocabulary size (16 728 words).
+Gibbs sampling maximises both entropy and likelihood when compared to other alternatives.
+The second best is the variational-expectation maximisation algorithm (VEM) when $\alpha$ is not set constant.
+$\alpha$ is an hyper parameter that weights the evenness of topic distribution.
+A lower $\alpha$ than default values indicates that the documents can be described by rather fewer topics, or that its distribution is very uneven.
+In fact, we observe that increasing the number of topics from our first 5 topics exploration (Fig 2) does increase entropy at expense of reducing $\alpha$, meaning that despite the larger number of topics, the main content of a document is captured by a few of them.
+The log-likelihood maximization stops around 25-50 topics.
+Thus, we restrict the rest of our analysis to 25 topics.
+
+\begin{figure*}[h]
+\centering
+\includegraphics[width = 4in, height = 3.5in]{fig3_topicnumber}
+\caption{\textbf{Algorithm and number of topic selection} Gibbs sampling maximizes entropy and log-likelihood estimation, making it a suitable algorithm for our data (A).
+Increasing the number of topics (from 5-100) shows that $\alpha$ decreases, suggesting that despite the larger number of topics, a few of them suffice to describe most papers (B).
+Log-likelihood is maximized for 50 topics followed closely with 25.
+While 50 topics is marginally better, for visualization purposes we continue our analysis with 25.
+Note that perplexity cannot be calculated for Gibbs sampling, hence the missing value in A and absence in B.}
+\label{fig:fig3}
+\end{figure*}
+
+Most topics maintain a relatively constant development over time (Fig 3 and 4).
+We do not observe strong trends, however some topics have gained attention.
+For example, topic 11 on complex systems theory, or topic 10 on ecosystem services appear constant and well scattered across time.
+In contrast, research on business innovation (topic 20), urban infrastructure (topic 3), sustainability (topic 1), or archaeology (topic 5) have gained attention in recent years.
+Topic 8 is outlier, with a selection of papers ranking high on content related to network infrastructure and performance possibly from engineering disciplines.
+It is the only topic with a clear set of papers separated from the rest of the collection with a decreasing trend.
+Topic 14 on agricultural practices and conservation, as well as 17 on modeling have a large share of the publications in recent years (see SM Fig 1).
+
+\begin{figure*}[h]
+\centering
+\includegraphics[width = 5in, height = 5in]{fig4_topics25}
+\caption{\textbf{Words probabilities per topics} $\beta$ is the probability of a word belonging to a topic (x-axis).
+Each plot summarizes the top ten words that best describe each of the topics of one of our top models.}
+\label{fig:fig4}
+\end{figure*}
+
+\begin{figure*}[h]
+\centering
+\includegraphics[width = 5in, height = 5in]{fig5_topics25_time}
+\caption{\textbf{Change of topic content over time} Each panel shows the topic content for each of the abstracts analyzed (N = 2177) and a loess smooth shows its change over time.
+$\gamma$ is the posterior probability of a topic describing a document.
+SM Fig 1 complement this figure with a summary of the proportion of papers per year per topic.}
+\label{fig:fig5}
+\end{figure*}
+
+The human-coded document analysis reveals that the most common panarchy feature in the literature is the adaptive cycle (81%, N = 42) followed by cross-scale interactions (Fig 6).
+Poverty and rigidity traps are the less studied features in our sample, some of them using slightly different terminology such as "lock-ins".
+The bulk of the papers analyzed are conceptual papers, many without a method or discernible research question.
+Roughly half of the papers are conceptual or restrict their use of panarchy as a metaphor.
+Over 40% of papers in our sample go a step ahead and use panarchy as an empirical construct, for example by attempting to identify the phases of the adaptive cycle (61%).
+Out of the 22 empirical cases, 6 are at the time scale of centuries, 7 on decades, 6 on years, and 1 in weeks; while 4 were city studies, 17 regional, and 1 national.
+Empirical papers are dominated by qualitative methods (77%, N = 22), generally retrospective historical reviews.
+
+The high level of conceptualization but lack of empirically grounded hypotheses or theory testing comes across as one of the major criticisms.
+Nonetheless, several of the papers reviewed attempted to identify the adaptive cycle either through qualitative or quantitative methods.
+Identification of adaptive cycles has found applications in a wide range of disciplines and research problems, from delimitation of periods in archaeology and anthropology [@Redman:2003bd], to financial crises in Europe [@Castell:2020dg], or traffic jams in China [@Zeng:2020hu].
+Recent empirical tests of the adaptive cycle innovate on the types of data and methods that one can use.
+Information transfer methods based on entropy have been suggested to approximate relevant components of a system and the empirical proxies of the axis where the adaptive cycle unfolds: potential, connectedness, and resilience [@Castell:2020dg].
+Percolation methods combined with big data have been shown useful to test hypothesis of regime shifts in urban systems (originally proposed in Panarchy [@Gunderson:2002vk]), and derive at which temporal and spatial scales does the adaptive cycle emerge [@Zeng:2020hu].
+
+\begin{figure*}[h]
+\centering
+\includegraphics[width = 6in, height = 3in]{fig6_qual_results}
+\caption{\textbf{Qualitative results} Document analysis was used to disentangle the different panarchy features addressed by a sub sample of papers (N = 42).
+Over half of the papers are conceptual work, and most empirical papers fall into qualitative methods.}
+\label{fig:fig6}
+\end{figure*}
 
@@
 
+Given the strong dominance of conceptual papers, it is not surprising that one of the key frontiers of research is to empirically ground such conceptualizations in a way that one can test hypotheses and advance theory beyond metaphors.
+Metaphors have played an important role as boundary objects enabling interdisciplinary dialogues.
+We observe it on the wide applications that panarchy has inspired, from anthropology to engineering.
+On the empirical realm, the adaptive cycle is the panarchy dimension that researchers engage the most with, but it is dominated by qualitative and retrospective studies.
+That means, we still lack the theory and methods to gather observations and be able to decide in which phase of the adaptive cycle a system currently is, or what is the probability of it transitioning to a new phase?
+
+The criticism of over conceptualization is not unique to the literature engaging Panarchy ideas.
+A recent review of sustainability science mapped the different schools of thought that the discipline has developed over the last decades [@Clark:2020jd].
+A similar conclusion was reached, where too many conceptual frameworks have been developed, but fewer are the empirical attempts to confront the frameworks against data to falsify hypotheses.
+The review emphasizes the problem of measurement and observation [@Clark:2020jd].
+In the context of panarchy, recent work has put forwards an information theory based approach that enables the identification of adaptive cycles in the European financial crises and a grassland ecosystems [@Castell:2020dg].
+The authors identify phases of the adaptive cycle, but fail to find support for the difference in speed between the forward and backward loops originally proposed in Panarchy.
+
+The panarchy dimensions that received less attention in our qualitative analysis are creative destruction, rigidity and poverty traps.
+But this is only because of the scope of our data: papers that have cited Panarchy or Holling's paper on the matter [@Gunderson:2002vk; @Holling:2001cu].
+Concepts such as creative destruction or poverty traps precedes Panarchy and therefore have been theorized and empirically grounded outside the panarchy stream of thinking.
+For example, the theory of poverty traps dates back to the 1950's in economics, it has received both theoretical development [@Bowles:2006hx] as well as empirical grounding [@Banerjee:2015ja; @Banerjee:2012tq] that has enable researchers and governments distinguish what kind of interventions are likely to reduce poverty.
+Here we encourage similar empirical efforts in formalizing theory and empirical support for the adaptive cycles, their cross-scale interactions (remember and revolt), or how the hierarchical nested nature of complex systems enhance of erode resilience in social-ecological systems.
+
+This paper aimed to study the use of panarchy related concepts and their evolution since the publication of the book.
+Systematic literature reviews often suffer from the limitation of restricting sample size to a subset of what is readable on the time frame of a project -- dozens maybe hundreds of papers.
+Qualitative analysis offers rich insights into the papers but is limited by sample size.
+Topic modeling enabled us to complement the analysis to all papers reported in major scientific databases.
+It has the advantage of reproducibility, but offers little insights about the dimensions of panarchy used, methods, or criticisms.
+Here we show that both methods combined are a powerful combination.
+Future studies could benefit from including gray literature such as theses, non-governmental organizations reports, government agencies reports, or non-english literature.
+Previous studies have shown that the use of full text instead of abstracts can also offer additional insights on the automated analyses, for example in attributing impacts of ecosystem services from regime shifts in social-ecological systems [@Rocha:2015tk].
+Replication attempts do however face the challenge of accessing full text when many of the papers are behind pay-walls.
+
 # Conclusion
@@
 
-# References {-}
-\bibliography{references}
+# References
+
+<div id="refs"></div>
 
@@
 
-# Supplementary Material
+# Supplementary Material {#sec:SM}
+
+\renewcommand\thefigure{S\arabic{figure}}
+\renewcommand\thetable{S\arabic{table}}
+\setcounter{table}{0}
+\setcounter{figure}{0}
+
+\begin{figure*}[h]
+\centering
+\includegraphics[width = 6in, height = 5in]{SM_topics25_timeline2}
+\caption{\textbf{Proportion of papers per year per topic.} }
+\label{fig:figS1}
+\end{figure*}

### 2020-11-23 01-read_data.R

@@ -1,0 +1,60 @@
+#  Topic modelling of "panarchy" data
+#  source data: scopus
+#  date: 20201114
+#  Juan Rocha
+#  
+# libraries for modeling
+library(tidytext)
+library(tidyverse)
+
+
+
+files <- list.files("data/") %>% str_subset(pattern = ".bib")
+files <- paste0("data/", files) 
+
+dat <- map(.x = files, .f = function(x){
+    x %>% 
+        bib2df::bib2df(separate_names = TRUE) %>%
+        janitor::clean_names()
+})
+
+dat <- bind_rows(dat)
+
+dat ## contains 2796 entries, of which 2357 are unique.
+
+dat %>% select(category, bibtexkey, author, journal, key, title, year, doi, abstract, coden, document_type) %>% unique()
+
+## Deciding which selection to use: needs to be unique and have an abstract
+dat %>% pull(bibtexkey) %>% unique() %>%  length() # 2357
+dat %>% pull(doi) %>% unique() %>%  length() # 2142
+dat %>% unique() # 2420
+dat %>% filter(!is.na(abstract)) # 2605
+dat %>% pull(abstract) %>% unique() %>% length() # 2192
+
+
+
+#### document-term matrix ####
+
+## extra words
+too_words <- tibble(
+    word = c("paper", "study", "aim", "aims", "objective", "purpose")
+)
+
+## creates the document term matrix
+dtm <- dat %>% 
+    select(abstract, title, year) %>%
+    filter(!is.na(abstract)) %>%
+    unique() %>% # 2193
+    unnest_tokens(word, abstract) %>% 
+    filter(!is.na(word)) %>% ## Books don't have abstract so they get dropped (n = 8)
+    anti_join(stop_words) %>% 
+    anti_join(too_words) %>%
+    filter(!str_detect(word, "[:digit:]")) %>%
+    group_by(title) %>%
+    count(word, sort = TRUE) %>%
+    cast_dtm(document = title, term = word, value = n)
+
+dtm
+
+save(dtm, file = "data/dtm.Rdata")
+

### 2020-11-23 02-model.R

@@ -1,0 +1,47 @@
+#  Topic modelling of "panarchy" data
+#  source data: scopus
+#  date: 20201114
+#  Juan Rocha
+#
+# libraries for modeling
+library(tidytext)
+library(tidyverse)
+library(tm)
+library (topicmodels)
+library(lda)
+
+load("data/dtm.Rdata")
+
+SEED <- 2010
+
+## Choosing best algorithm
+k <- 10
+
+tset.TM <- list (
+    VEM0 = LDA(dtm, k=k, control = list ( seed = SEED)),
+    VEM_fixed= LDA(dtm, k=k, control= list (estimate.alpha = F, seed = SEED)),
+    Gibbs = LDA (dtm, k=k, method ="Gibbs", control = list (seed = SEED, burnin= 1000, thin = 100, iter= 1000)),
+    CTM = CTM (dtm, k=k, control = list(seed = SEED, var= list (tol= 10^-4), em= list (tol = 10^-3))))
+
+sapply (tset.TM[1:3], slot, "alpha")
+
+#Finding number of topics
+k <- c(5,10,25,50,100)
+
+topicNumber.TM <- map(
+    .x = k,
+    .f = function(x) {
+    LDA(dtm, k = x, control= list (seed = SEED), method = "Gibbs")
+    })
+
+
+save(tset.TM, topicNumber.TM, file = "data/models_gibbs.RData")
+
+#### Lance Gunderson pubs ####
+
+lance <- bib2df::bib2df(
+    file = "data/scopus_gunderson_201121.bib",
+    separate_names = TRUE) %>%
+    janitor::clean_names()
+
+lance %>% pull(doi) %>% unique()

### 2020-11-23 03-figueres.R

@@ -1,0 +1,199 @@
+library(tidytext)
+library(tidyverse)
+library(tm)
+library (topicmodels)
+library(lda)
+
+load("data/models_gibbs.RData")
+
+#Finding number of topics
+k <- c(5,10,25,50,100)
+
+## visualizations:
+df_topic_number <- tibble(
+    topic_number = k,
+    entropy = map_dbl (topicNumber.TM, function (x)
+        mean(apply(posterior(x)$topics, 1, function (z) - sum(z * log(z)))) # maximize Entropy
+    ),
+    alpha = map_dbl(topicNumber.TM, slot, "alpha"),
+    log_lik = map_dbl(topicNumber.TM, logLik) #,  #maximize loglik
+    #perplexity = map_dbl(topicNumber.TM, perplexity) #minimize perplexity
+)
+
+#### algorithm selection ####
+df_stats <- tibble(
+    model = names(lapply(tset.TM, logLik)),
+    loglik = as.numeric(lapply(tset.TM, logLik)), #maximize loglik
+    entropy = lapply (tset.TM, function (x) 
+        mean(apply(posterior(x)$topics,
+                   1, function (z) - sum(z * log(z))))) %>% as.numeric()#maximize ENTROPY
+    
+)
+
+
+perp <-  lapply(tset.TM[c(1,2,4)], perplexity)
+perp$Gibbs <- NA
+# pretty names:
+df_stats$model <- c("VEM alpha", "VEM fixed", "Gibbs", "CTM")
+
+df_stats %>%
+    add_column(perplexity = as.numeric(perp[c(1,2,4,3)])) %>% #minimize perplexity
+    pivot_longer(cols = 2:4, names_to = "measure", values_to = "value") %>%
+    ggplot(aes(model, value)) + 
+    geom_col() + facet_wrap(.~measure, scales = "free") +
+    theme_light(base_size = 6) 
+
+# ggsave(filename = "figures/fig1_algorithm_selection.png",
+#        device = "png", width = 5, height = 2)
+
+# J201117: the best algo is Gibbs sampling!
+
+#### number of topics ####
+df_topic_number %>%
+    # mutate(alpha_log = log10(alpha)) %>%
+    pivot_longer(cols = 2:last_col(), names_to = "measure", values_to = "value") %>%
+    # filter(measure != "alpha") %>%
+    ggplot(aes(as.factor(topic_number), value)) +
+    geom_col() + 
+    labs(x = "Number of topics") +
[…]

### 2020-12-18 paper/paper_panarchy.Rmd
major update on manuscript with co-authors input. New figures.

@@
     | \small ^3^South American Institute for Resilience and Sustainability Studies, Maldonado, Uruguay
-    | \small ^4^...
-    | \small ^5^...
-    | \small ^6^...
-    | \small ^7^...
+    | \small ^4^.Center for Complex Systems in Transition, Stellenbosch University, South Africa
+    | \small ^5^Department of Geography, McGill University, 805 rue Sherbrooke ouest, Montréal (Québec) H3A 0B9, Canada
+    | \small ^6^Department of Natural Resource Sciences, McGill University, 21111 Lakeshore Road, Sainte-11 Anne-de-Bellevue, QC, H9X 3V9, Canada
+    | \small ^7^School of Sustainability, Arizona State University, Tempe, Arizona, USA
 
 abstract: |
-  Panarchy has marked the academic career of many scholars and practitioners facing environmental problems.
+  How do social-ecological systems change over time?
+In 2002 Holling and colleagues proposed the concept of Panarchy, which presented social-ecological systems as an interacting set of adaptive cycles, each of which is produced by the dynamic tensions between novelty and efficiency at multiple scales.
 Initially introduced as a conceptual framework and set of metaphors, panarchy has gained the attention of scholars across many disciplines and its ideas continue to inspire further conceptual developments.
-To what extent these concepts have developed, theorized and empirically grounded are questions we address by reviewing the literature with qualitative methods and machine learning.
-We find that the adaptive cycle is the feature of panarchy that has attracted more attention.
+Almost twenty years after this concept was introduced we review how it has been used, tested, extended and revised.
+We do this by using  qualitative methods and machine learning.
+We find that the adaptive cycle is the feature of panarchy that has attracted the most attention.
 Challenges remain in empirically grounding the metaphor, but recent theoretical and empirical work offer some avenues for future research.
@@
 
+Almost two decades ago, the edited book “Panarchy: Understanding Transformations in Human and Natural Systems”  [@Gunderson:2002vk], presented a synthetic perspective of how a group of leading social-ecological researchers associated with the Resilience Alliance understood change in social-ecological systems.
+The concept of Panarchy was a key focus on this influential book.
+In the late 1990s and early 2000s, the Resilience Alliance was a productive, innovative highly collaborative group of environmental scientists that was focused on bridging social and natural sciences as well as theory and practice.
+They were very focused on addressing social-ecological problems by combining insights from the social and natural sciences, as well as the arts and humanities [@Parker:2012eo].
+
 Panarchy encapsulates a set of concepts that have inspired the work of environmental scientists and practitioners for almost two decades.
-It is the product of a group of scientists interested in solving environmental issues by integrating insights from the social and natural sciences, as well as the arts and humanities.
+It was a central concept produced by the Resilience Alliance, an international group of influential academics focused on the resilience of social-ecological systems in theory and practice [@Parker:2012eo].
+The Panarchy concept builds on Holling’s adaptive cycle [@biosphere:hJ4c_Ua8] by extending the idea across spatial and temporal scales.
+
 Panarchy was, and still is, a boundary object that has enabled collaborations and nurtured new scientific frameworks.
-The ideas put forward have found applications in field studies, mathematical models, participatory work, and scenario development.
-It has inspired resilience assessments and guided decision making.
+The ideas put forward have been applied in field studies, mathematical models, participatory work, and scenario development.
+Panarchy has inspired resilience assessments and guided decision making.
 In this article we pay tribute to the book by studying how the concepts and metaphors proposed have been further developed in the academic literature.
-We also document the criticism to the concepts and identify key research frontiers.
+We also document criticism of the concepts and identify key research frontiers.
 
@@
 It is not a predictive tool, but aims to understand adaptive change.
-Panarchy posits that systems are organized in nested hierarchies across space and time.
-Each level of the hierarchy is a subsystem that can be in different phases of adaptive cycles.
+The adaptive cycle is hypothesized to exist within a three-dimensional space defined by three properties: potential, connectivity, and resilience [@Gunderson:2002vk].
+Potential refers to the capital available capital to the system, for example nutrients and carbon captured by a forest, or human capital - skills and knowledge accumulated to run the economy.
+Connectivity is a proxy of the structure of the system; it is the network of interactions and strength between its elements.
+Resilience is the capacity of any system to absorb disturbance and reorganize while undergoing change so as to retain essentially the same identity, as determined by its function, structure, and feedbacks [@Anonymous:2016fv].
+
+Panarchy posits that systems are organized in nested hierarchies across space and time, where each level of the hierarchy is a subsystem that can be in a different phase of the adaptive cycle.
 These phases are exploitation, conservation, release, and reorganization, the latter characterized by events of creative destruction (Fig 1).
-They are inspired in mathematical models used in economics and ecology, but broadly describe patterns of growth, collapse, and recovery that are common to populations, ecological communities, markets, or political organizations.
-While each subsystem in scale can be at different phases of the cycle, they can influence each other through cross-scale interactions called revolt and remember.
+They are inspired by mathematical models used in economics and ecology, but broadly describe patterns of growth, collapse, and recovery that are common to populations, ecological communities, markets, or political organizations.
+While each subsystem in the hierarchy can be at a different phase of the cycle, these subsystems can influence each other through cross-scale interactions called “revolt” and “remember.”
 
-The adaptive cycle is hypothesized to exist in a three-dimensional space that integrates three key systemic features: its potential, connectivity, and resilience [@Gunderson:2002vk].
-The potential refers to the available capital to the system, which can be approximated as nutrients and carbon captured by a forest, or the human capital of skills and knowledge accumulated to run the economy.
-Connectivity is a proxy of the structure of the system, it is the network of interactions and strength between its elements.
-Resilience is the capacity of any system to absorb disturbance and reorganize while undergoing change so as to still retain essentially the same function, structure, feedbacks, and therefore identity [@Anonymous:2016fv].
-
-The three-dimensional space has corners with attractors that can stagnate the adaptive cycle: poverty and rigidity traps [@Gunderson:2002vk].
-In panarchy, poverty traps are described as maladaptive states where potential, connectivity and resilience are low.
-Poverty traps are a series of feedback mechanisms that reinforce impoverished states [@Bowles:2006hx], limiting the capacity to innovate and increase potential.
-The opposite corner, where potential, connectivity and resilience are high, is also a maladaptive space denoted rigidity traps.
+The three-dimensional space has corners with attractors that can cause the adaptive cycle to stagnate: poverty and rigidity traps [@Gunderson:2002vk].
+Poverty traps are described in panarchy as maladaptive states where potential, connectivity and resilience are low.
+Poverty traps are a series of feedback mechanisms that reinforce impoverished states [@Bowles:2006hx], limiting the system’s capacity to innovate and increase potential.
+The opposite corner, where potential, connectivity and resilience are high, is another maladaptive space called a rigidity trap.
 In that corner there is little space for experimentation and innovation.
@@
 
-Panarchy offers a rich conceptual framework to think of environmental problems.
-While inspired in a few mathematical constructs (e.g.
+Panarchy offers a rich conceptual framework for understanding environmental problems.
+While inspired by a few mathematical constructs (e.g.
 cycles, traps, scaling laws), it is general enough to invite scholars from multiple disciplinary backgrounds, ontologies and epistemologies, to collaborate around research questions and applied problems.
-It is a boundary object.
-The book itself presents a series of case studies where geographers, economists, political scientists, and ecologists have found the framework useful.
-However, to what extent have these concepts gained empirical support beyond the metaphor, or how these concepts have evolved, remind open questions.
-For example, the adaptive cycle is a useful metaphor to look back at history and organize events and periods around phases.
-It remains challenging however to narrow down in what phase is a system in the present, or what would be its expected trajectory in the future.
-In other words, it seems useful in retrospective studies, but its application is more limited for prospective ones.
+As such, it is useful as a boundary object  and can be used empirically or metaphorically.
+The book itself presents a series of case studies where geographers, economists, political scientists, and ecologists have demonstrated the utility of the framework to their area of research.
+However, the extent to which these concepts have gained empirical support beyond the metaphor, or how they have evolved over time, remain open questions.
+For example, the adaptive cycle is a useful metaphor to look back at history and organize events and periods around phases, but it remains challenging to identify the phase of a system in real time, or to use panarchy to project future trajectories.
+In other words, panarchy is useful for retrospective studies, but its applications may be more limited for prospective ones.
 Are we at a stage where we can start distilling theory that enables case study comparisons, deriving functional forms, or draw theoretical expectations and predictions?
 
-Here we review the academic literature of the last two decades with the aim of tracing how these concepts have evolved, in what type of problems are they found useful, and what remain key frontiers of research.
+Here we review the academic literature of the last two decades to trace how these concepts have evolved, in what type of problems they have been applied and found useful, and finally what remains as the key frontiers of research.
 
@@
 To answer these questions we combined an automated literature review based on topic modeling [@Blei:2012dk; @Griffiths:2004ey] with human-coded document analysis [@Bryman:2008va].
-All data and code to replicate the analysis is available at: [https://github.com/juanrocha/panarchy](https://github.com/juanrocha/panarchy) and a public online repository (add figshare link when finished).
+All data and code to replicate the analysis are available at: [https://github.com/juanrocha/panarchy](https://github.com/juanrocha/panarchy) and a public online repository (add figshare link when finished).
 
@@
 We removed stop words (e.g.
-"the", "a") and digits from the matrix, as well as a list of words that were over represented in our data (e.g.
-"paper", "study", "aim") that are common in the scientific literature but do not inform us about the topics the papers are about.
+"the", "a") and digits from the matrix, as well as a list of words that were over represented in our data  and are common in the scientific literature but are unrelated to the papers’ topics(e.g.
+"paper", "study", "aim").
 
@@
 We compared three LDA algorithms: correlated topic models (CTM), variational expectation maximization (VEM), and Gibbs sampling (Gibbs), by assessing their performance against their log-likelihood estimation, entropy and perplexity [@Grun:2011tb].
+Entropy is a measure of order or disorder of a system; in the context of topic models it measures how evenly the topic distribution is spread.
+Perplexity measures the uncertainty of predicting a single word, so if the model performance were the same as random, perplexity would approximate the vocabulary size (16 728 words).
 These metrics enable us to choose which algorithm best fits our data, what is the optimal number of topics to fit, and how to avoid overfitting.
 
-*Document analysis*: We complemented our unsupervised approach with document analysis [@Bryman:2008va] by coding an additional set of categorical variables for a sample of papers (N = 41).
+*Document analysis*: We complemented our unsupervised approach with document analysis [@Bryman:2008va] by coding an additional set of categorical variables for a sample of highly cited papers (N = 41).
 We annotated qualitative aspects such as use of the adaptive cycle, identification of its phases, whether the paper is conceptual, modeling or an empirical study.
@@
 For example whether there is emphasis on cross-scale interactions, or poverty and rigidity traps.
-Text annotations were used to capture potential criticisms as well as the methods used.
+We used text annotations to capture potential criticisms as well as the methods used.
 The code book is available in the supplementary material (archived in XX repository).
@@
 
-Panarchy, the book published in 2002, has been cited over 7200 times according to Google Scholar.
+_Panarchy_, the book published in 2002, has been cited over 7200 times according to Google Scholar.
 The scientific paper that introduced the book to the scientific community [@Holling:2001cu] has received 1715 citations in the Web of Science, and 1923 in Scopus.
-Roughly half of the citations comes from environmental (28%) and social sciences (22%), while the less represented fields are computer science (2.1%), arts and humanities (2.6%).
+Roughly half of the citations have come from environmental (28%) and social sciences (22%).
+Computer science (2.1%)and arts and humanities (2.6%) have been the least represented.
 
@@
 
-\begin{figure*}[h]
-\centering
-\includegraphics[width = 5in, height = 3in]{fig2_five_topics_time}
-\caption{\textbf{A simple model with 5 topics} Topics are latent variables that rank words according to the probability of the word explaining the topic (A).
-Documents are composed by topics in different proportions which allows the identification of clusters of documents around the same topics, or how the topic content change over time (B).
-Each dot in the lower pane is an abstract of a paper (N = 2177) organized by the year it was published against each of the topic content.}
-\label{fig:fig2}
-\end{figure*}
-
-Comparing algorithm performance reveals that Gibbs sampling better fits the data.
+Comparing algorithm performance revealed that Gibbs sampling was the best fit for the data.
 As a rule of thumb, an ideal method should maximize entropy and the log-likelihood estimation while minimizing perplexity [@Grun:2011tb].
-Entropy is a measure of order or disorder of a system, in the context of topic models it measures how evenly spread is the topic distribution.
-Perplexity measures the uncertainty of predicting a single word, so if the model performance were the same as random, perplexity will approximate the vocabulary size (16 728 words).
-Gibbs sampling maximises both entropy and likelihood when compared to other alternatives.
-The second best is the variational-expectation maximisation algorithm (VEM) when $\alpha$ is not set constant.
-$\alpha$ is an hyper parameter that weights the evenness of topic distribution.
+Gibbs sampling maximised both entropy and likelihood with our data when compared to other alternatives.
+The second best was the variational-expectation maximisation algorithm (VEM) when $\alpha$ was not set constant.
+$\alpha$ is a hyper parameter that weights the evenness of topic distribution.
 A lower $\alpha$ than default values indicates that the documents can be described by rather fewer topics, or that its distribution is very uneven.
-In fact, we observe that increasing the number of topics from our first 5 topics exploration (Fig 2) does increase entropy at expense of reducing $\alpha$, meaning that despite the larger number of topics, the main content of a document is captured by a few of them.
-The log-likelihood maximization stops around 25-50 topics.
-Thus, we restrict the rest of our analysis to 25 topics.
+In fact, we observed that increasing the number of topics from 5 to 100 topics did increase entropy at expense of reducing $\alpha$, meaning that despite the larger number of topics, the main content of a document is still captured by a few of them.
+The log-likelihood maximization stopped at around 25-50 topics.
+Thus, we restricted the rest of our analysis to 25 topics.
 
@@
 \centering
-\includegraphics[width = 4in, height = 3.5in]{fig3_topicnumber}
+\includegraphics[width = 4in, height = 3.5in]{fig2_topicnumber}
 \caption{\textbf{Algorithm and number of topic selection} Gibbs sampling maximizes entropy and log-likelihood estimation, making it a suitable algorithm for our data (A).
@@
 Log-likelihood is maximized for 50 topics followed closely with 25.
-While 50 topics is marginally better, for visualization purposes we continue our analysis with 25.
+While 50 topics is marginally better, for visualization purposes we restrict our analysis to 25.
 Note that perplexity cannot be calculated for Gibbs sampling, hence the missing value in A and absence in B.}
-\label{fig:fig3}
+\label{fig:algorithm}
 \end{figure*}
 
-Most topics maintain a relatively constant development over time (Fig 3 and 4).
-We do not observe strong trends, however some topics have gained attention.
-For example, topic 11 on complex systems theory, or topic 10 on ecosystem services appear constant and well scattered across time.
-In contrast, research on business innovation (topic 20), urban infrastructure (topic 3), sustainability (topic 1), or archaeology (topic 5) have gained attention in recent years.
-Topic 8 is outlier, with a selection of papers ranking high on content related to network infrastructure and performance possibly from engineering disciplines.
-It is the only topic with a clear set of papers separated from the rest of the collection with a decreasing trend.
-Topic 14 on agricultural practices and conservation, as well as 17 on modeling have a large share of the publications in recent years (see SM Fig 1).
+A topic is a set of words that are ranked according to a probability that they represent an underlying content of the document (Fig 3).
+For example, the words "resilience", "vulnerability", “risk” and "disaster" have a high probability to capture the content of topic 25.
+Papers early in our time series (2001-3) have high content largely dominated by topics 25 on resilience, 24 on the adaptive cycle, and 18 on adaptive governance and management.
+Towards the end of the time series, topic 3 on urban systems and 9 on local communities and knowledge have spikes up to 9% of the content of each year (2019-2020).
+For comparison, if all topics were equally represented in the content, they would have 4% in the corpus (grey line in Fig 3B).
+
+Despite fluctuations, most topics showed a relatively constant level of interest over time (Fig 3, Fig S2).
+We did not observe strong trends, but some topics have gained a small amount of attention.
+For example, topic 1 on sustainability, or topic 10 on ecosystem services appeared consistently across time.
+In contrast, research on business innovation (topic 20), urban infrastructure (topic 3), or archaeology (topic 5) have gained attention in recent years.
+Topic 8 was an outlier, with a selection of papers ranking high on content related to network infrastructure and performance, possibly from engineering disciplines.
+It was the only topic with a set of papers that was clearly distinct from the rest of the collection, and showed a decreasing trend over time.
 
@@
 \centering
-\includegraphics[width = 5in, height = 5in]{fig4_topics25}
-\caption{\textbf{Words probabilities per topics} $\beta$ is the probability of a word belonging to a topic (x-axis).
-Each plot summarizes the top ten words that best describe each of the topics of one of our top models.}
-\label{fig:fig4}
+\includegraphics[width = 5in, height = 5in]{fig3_topics25}
+\caption{\textbf{Panarchy topics over time} The number of papers per year (A) with a maximum of 324 in 2019.
+The proportion of papers per year (B) and the relative proportion of topic content per year (C) do not show strong trends for the time window with most papers (gray area in A).
+Each topic is summarized in (D) by the top 10-words that best describe them according to the posterior probability of our  model fit.
+See FigS2 for a panel of time series per topics}
+\label{fig:topics25}
 \end{figure*}
 
-\begin{figure*}[h]
-\centering
-\includegraphics[width = 5in, height = 5in]{fig5_topics25_time}
-\caption{\textbf{Change of topic content over time} Each panel shows the topic content for each of the abstracts analyzed (N = 2177) and a loess smooth shows its change over time.
-$\gamma$ is the posterior probability of a topic describing a document.
-SM Fig 1 complement this figure with a summary of the proportion of papers per year per topic.}
-\label{fig:fig5}
-\end{figure*}
+The human-coded document analysis revealed that the most common feature of panarchy in the literature was the adaptive cycle (81%, N = 42) followed by cross-scale interactions (Fig 4).
+Poverty and rigidity traps were less studied features in our sample, even when accounting for slightly different terminology such as "lock-ins".
+The bulk of the papers analyzed were conceptual papers, many without a method or discernible research question.
+Roughly half of the papers were conceptual or only used panarchy as a metaphor.
+Over 40% of papers in our sample went a step further and used panarchy as an empirical construct, for example by attempting to identify the phases of the adaptive cycle (61%).
+Out of the 22 empirical cases, 6 were at the time scale of centuries, 7 on decades, 6 on years, and 1 in weeks.
+Spatially, 4 were at a city scale, 17 regional, and 1 national.
+Empirical papers were dominated by qualitative methods (77%, N = 22) and were generally retrospective historical reviews.
 
-The human-coded document analysis reveals that the most common panarchy feature in the literature is the adaptive cycle (81%, N = 42) followed by cross-scale interactions (Fig 6).
-Poverty and rigidity traps are the less studied features in our sample, some of them using slightly different terminology such as "lock-ins".
-The bulk of the papers analyzed are conceptual papers, many without a method or discernible research question.
-Roughly half of the papers are conceptual or restrict their use of panarchy as a metaphor.
-Over 40% of papers in our sample go a step ahead and use panarchy as an empirical construct, for example by attempting to identify the phases of the adaptive cycle (61%).
-Out of the 22 empirical cases, 6 are at the time scale of centuries, 7 on decades, 6 on years, and 1 in weeks; while 4 were city studies, 17 regional, and 1 national.
-Empirical papers are dominated by qualitative methods (77%, N = 22), generally retrospective historical reviews.
-
-The high level of conceptualization but lack of empirically grounded hypotheses or theory testing comes across as one of the major criticisms.
+The high level of conceptualization but lack of empirically grounded hypotheses or theory testing came across as one of the major limitations.
 Nonetheless, several of the papers reviewed attempted to identify the adaptive cycle either through qualitative or quantitative methods.
@@
 Information transfer methods based on entropy have been suggested to approximate relevant components of a system and the empirical proxies of the axis where the adaptive cycle unfolds: potential, connectedness, and resilience [@Castell:2020dg].
-Percolation methods combined with big data have been shown useful to test hypothesis of regime shifts in urban systems (originally proposed in Panarchy [@Gunderson:2002vk]), and derive at which temporal and spatial scales does the adaptive cycle emerge [@Zeng:2020hu].
+Percolation methods combined with big data have been shown to be useful to test hypotheses of regime shifts in urban systems (originally proposed in Panarchy [@Gunderson:2002vk]), and derive the temporal and spatial scales at which the adaptive cycle emerges [@Zeng:2020hu].
 
@@
 \centering
-\includegraphics[width = 6in, height = 3in]{fig6_qual_results}
+\includegraphics[width = 6in, height = 3in]{fig4_qual_results}
 \caption{\textbf{Qualitative results} Document analysis was used to disentangle the different panarchy features addressed by a sub sample of papers (N = 42).
 Over half of the papers are conceptual work, and most empirical papers fall into qualitative methods.}
-\label{fig:fig6}
+\label{fig:qual}
 \end{figure*}
@@
 We observe it on the wide applications that panarchy has inspired, from anthropology to engineering.
-On the empirical realm, the adaptive cycle is the panarchy dimension that researchers engage the most with, but it is dominated by qualitative and retrospective studies.
+In the empirical realm, the adaptive cycle is the panarchy dimension that researchers engage the most with, but it is dominated by qualitative and retrospective studies.
 That means, we still lack the theory and methods to gather observations and be able to decide in which phase of the adaptive cycle a system currently is, or what is the probability of it transitioning to a new phase?
@@
 A recent review of sustainability science mapped the different schools of thought that the discipline has developed over the last decades [@Clark:2020jd].
-A similar conclusion was reached, where too many conceptual frameworks have been developed, but fewer are the empirical attempts to confront the frameworks against data to falsify hypotheses.
+A similar conclusion was reached, where too many conceptual frameworks have been developed, but fewer empirical attempts try to test the frameworks against data to falsify hypotheses.
 The review emphasizes the problem of measurement and observation [@Clark:2020jd].
-In the context of panarchy, recent work has put forwards an information theory based approach that enables the identification of adaptive cycles in the European financial crises and a grassland ecosystems [@Castell:2020dg].
-The authors identify phases of the adaptive cycle, but fail to find support for the difference in speed between the forward and backward loops originally proposed in Panarchy.
+In the context of panarchy, recent work has developed an information theory based approach that enables the identification of adaptive cycles [@Castell:2020dg].
+The authors identify phases of the adaptive cycle  in the European financial crises and in grassland ecosystems, but fail to find support for the difference in speed between the forward and backward loops originally proposed in Panarchy.
 
 The panarchy dimensions that received less attention in our qualitative analysis are creative destruction, rigidity and poverty traps.
-But this is only because of the scope of our data: papers that have cited Panarchy or Holling's paper on the matter [@Gunderson:2002vk; @Holling:2001cu].
-Concepts such as creative destruction or poverty traps precedes Panarchy and therefore have been theorized and empirically grounded outside the panarchy stream of thinking.
-For example, the theory of poverty traps dates back to the 1950's in economics, it has received both theoretical development [@Bowles:2006hx] as well as empirical grounding [@Banerjee:2015ja; @Banerjee:2012tq] that has enable researchers and governments distinguish what kind of interventions are likely to reduce poverty.
-Here we encourage similar empirical efforts in formalizing theory and empirical support for the adaptive cycles, their cross-scale interactions (remember and revolt), or how the hierarchical nested nature of complex systems enhance of erode resilience in social-ecological systems.
+This may be at least partly due to the scope of our data: papers that have cited Panarchy or Holling's [@Gunderson:2002vk; @Holling:2001cu].
+Concepts such as creative destruction and poverty traps precede Panarchy and therefore have been theorized and empirically grounded outside the panarchy stream of thinking.
+For example, the theory of poverty traps dates back to the 1950's in economics, and has received both theoretical development [@Bowles:2006hx] as well as empirical grounding [@Banerjee:2015ja; @Banerjee:2012tq] that has enabled researchers and governments to distinguish what kind of interventions are likely to reduce poverty.
+Here we encourage similar empirical efforts in formalizing theory and empirical support for the adaptive cycles, their cross-scale interactions (remember and revolt), or how the hierarchical nested nature of complex systems enhance or erode resilience in social-ecological systems.
 
 This paper aimed to study the use of panarchy related concepts and their evolution since the publication of the book.
-Systematic literature reviews often suffer from the limitation of restricting sample size to a subset of what is readable on the time frame of a project -- dozens maybe hundreds of papers.
+Systematic literature reviews often suffer from the limitation of restricting sample size to a subset of what is readable on the time frame of a project -- between dozens and maybe hundreds of papers.
 Qualitative analysis offers rich insights into the papers but is limited by sample size.
+In our analysis, the qualitative insights could be biased towards a non-random selection of papers that aligned our research interests, or highly cited, review type of work.
 Topic modeling enabled us to complement the analysis to all papers reported in major scientific databases.
-It has the advantage of reproducibility, but offers little insights about the dimensions of panarchy used, methods, or criticisms.
+It has the advantage of reproducibility, and reduces bias, but offers little insights about the dimensions of panarchy used, methods, or criticisms.
 Here we show that both methods combined are a powerful combination.
@@
 (400w)
+
+We reviewed the academic literature of the last two decades to trace how Panarchy related concepts have evolved, in what type of problems they have been applied and found useful, and what remains as the key frontiers of research.
+Despite a growing body of literature, no topic seems to dominate the academic production of scholars using these concepts.
+The feature most used is the adaptive cycle, and the problems where it is found most useful is in studies with a historical perspective, either in short time horizons such as natural resource management, urban development, or conservation; all the way to archaeology and anthropology studies on the time horizon of centuries to millennia.
+Hierarchies in scale or maladaptive traps have received comparatively less attention.
+The frontiers of research point out to the problem of measurement and prediction.
+For example, how to observe and approximate resilience, connectedness and potential in the present; or how likely a current social-ecological system is to undergo a transition in the phases of the adaptive cycle, or a “memory” or “revolt” type of cross-scale interaction.
+While our survey of the literature offers a few options to empirically ground panarchy concepts, we believe the problem of measurement and prediction are fertile ground for future research efforts.
 
@@
 \centering
-\includegraphics[width = 6in, height = 5in]{SM_topics25_timeline2}
+\includegraphics[width = 6in, height = 5in]{SM_topics25_timeline}
 \caption{\textbf{Proportion of papers per year per topic.} }

### 2020-12-18 03-figueres.R

@@ -3,6 +3,8 @@ library(tidyverse)
 library(tm)
 library (topicmodels)
 library(lda)
+library(patchwork)
+
 
 load("data/models_gibbs.RData")
 
@@ -36,11 +38,12 @@ perp$Gibbs <- NA
 # pretty names:
 df_stats$model <- c("VEM alpha", "VEM fixed", "Gibbs", "CTM")
 
-df_stats %>%
+g1 <- df_stats %>%
     add_column(perplexity = as.numeric(perp[c(1,2,4,3)])) %>% #minimize perplexity
     pivot_longer(cols = 2:4, names_to = "measure", values_to = "value") %>%
     ggplot(aes(model, value)) + 
     geom_col() + facet_wrap(.~measure, scales = "free") +
+    labs(x = "Algorithm", y = "Value", tag = "A") +
     theme_light(base_size = 6) 
 
 # ggsave(filename = "figures/fig1_algorithm_selection.png",
@@ -49,24 +52,26 @@ df_stats %>%
 # J201117: the best algo is Gibbs sampling!
 
 #### number of topics ####
-df_topic_number %>%
+g2 <- df_topic_number %>%
     # mutate(alpha_log = log10(alpha)) %>%
     pivot_longer(cols = 2:last_col(), names_to = "measure", values_to = "value") %>%
     # filter(measure != "alpha") %>%
     ggplot(aes(as.factor(topic_number), value)) +
     geom_col() + 
-    labs(x = "Number of topics") +
+    labs(x = "Number of topics", y = "Value", tag = "B") +
     facet_wrap(.~measure, scales = "free", ncol = 4, nrow = 1) +
     theme_light(base_size = 6)
 
-# ggsave(filename = "figures/fig2_topicnumber.png",
-#        device = "png", width = 5, height = 2)
+g1/g2
+
+ggsave(filename = "paper/figures/fig3_topicnumber.png",
+       device = "png", width = 5, height = 4)
 
 
 #### heatmaps topics ####
 
 gplots::heatmap.2(
-    posterior(topicNumber.TM[[2]])$topics, 
+    posterior(topicNumber.TM[[3]])$topics, 
     key=T, trace="none", margins=c(5,5), 
     cexRow=0.7, cexCol=0.7, col="topo.colors", keysize=1, key.par = list(cex = 0.5),
     labRow = FALSE, key.title = NA, key.xlab = "probability", key.ylab = "count" )
[…]

### 2020-12-28 paper/paper_panarchy.Rmd
submission + pre-print preparation

@@
     | \small
-    | \small ^1^Stockholm Resilience Centre, Stockholm University, Kräftriket 2B, 10691 Stockholm
-    | \small ^2^Future Earth, The Swedish Royal Academy of Sciences, Lilla Frescativägen 4A, 104 05 Stockholm
-    | \small ^3^South American Institute for Resilience and Sustainability Studies, Maldonado, Uruguay
-    | \small ^4^.Center for Complex Systems in Transition, Stellenbosch University, South Africa
-    | \small ^5^Department of Geography, McGill University, 805 rue Sherbrooke ouest, Montréal (Québec) H3A 0B9, Canada
-    | \small ^6^Department of Natural Resource Sciences, McGill University, 21111 Lakeshore Road, Sainte-11 Anne-de-Bellevue, QC, H9X 3V9, Canada
-    | \small ^7^School of Sustainability, Arizona State University, Tempe, Arizona, USA
-
+    | \footnotesize ^1^Stockholm Resilience Centre, Stockholm University, Kräftriket 2B, 10691 Stockholm
+    | \footnotesize ^2^Future Earth, The Swedish Royal Academy of Sciences, Lilla Frescativägen 4A, 104 05 Stockholm
+    | \footnotesize ^3^South American Institute for Resilience and Sustainability Studies, Maldonado, Uruguay
+    | \footnotesize ^4^Center for Complex Systems in Transition, Stellenbosch University, South Africa
+    | \footnotesize ^5^Department of Geography, McGill University, 805 rue Sherbrooke ouest, Montréal (Québec) H3A 0B9, Canada
+    | \footnotesize ^6^Department of Natural Resource Sciences, McGill University, 21111 Lakeshore Road, 
+    | \footnotesize Sainte-11 Anne-de-Bellevue, QC, H9X 3V9, Canada
+    | \footnotesize ^7^School of Sustainability, Arizona State University, Tempe, Arizona, USA
 abstract: |
@@
 Almost twenty years after this concept was introduced we review how it has been used, tested, extended and revised.
-We do this by using  qualitative methods and machine learning.
+We do this by combining qualitative methods and machine learning.
+Document analysis was used to code panarchy features that are commonly used in the scientific literature (N = 42), a qualitative analysis that was complemented with topic modeling of 2177 documents.
 We find that the adaptive cycle is the feature of panarchy that has attracted the most attention.
 Challenges remain in empirically grounding the metaphor, but recent theoretical and empirical work offer some avenues for future research.
-
 bibliography: panarchy.bib
@@
     dev: pdf
-    keep_tex: no
+    keep_tex: yes
     toc: no
     latex_engine: xelatex
-  word_document:
+  rtf_document:
     toc: no
@@
 fontsize: 10pt
-
 linkcolor: blue
@@
 - \setmainfont{Arial}
-- \linenumbers
 - \renewcommand{\abstractname}{}
@@
 - \usepackage[margin=2.5cm]{geometry}
+
+## For Arxiv submission, use pdflatex as engine, comment out font, increase font size to 11, delete the package fontspec and the graphics path.
+Compile in latex so the bibliography is part of the tex file produced.
 ---
@@
 The concept of Panarchy was a key focus on this influential book.
-In the late 1990s and early 2000s, the Resilience Alliance was a productive, innovative highly collaborative group of environmental scientists that was focused on bridging social and natural sciences as well as theory and practice.
-They were very focused on addressing social-ecological problems by combining insights from the social and natural sciences, as well as the arts and humanities [@Parker:2012eo].
+In the late 1990s and early 2000s, the Resilience Alliance was a productive, innovative highly collaborative group of environmental scientists that was focussed on bridging social and natural sciences as well as theory and practice.
+They were very focussed on addressing social-ecological problems by combining insights from the social and natural sciences, as well as the arts and humanities [@Parker:2012eo].
 
@@
 
-Panarchy was, and still is, a boundary object that has enabled collaborations and nurtured new scientific frameworks.
-The ideas put forward have been applied in field studies, mathematical models, participatory work, and scenario development.
+Panarchy remains a boundary object that has inspired research topics, enabled collaborations and nurtured new scientific frameworks.
+The ideas put forward have been applied in field studies, archaeology, mathematical models, participatory work, and scenario development.
 Panarchy has inspired resilience assessments and guided decision making.
@@
 We also document criticism of the concepts and identify key research frontiers.
+
+Panarchy proposes that it is useful to conceptualize systems in terms of interacting adaptive cycles.
+The adaptive cycle was an idea first proposed by Holling based upon his experience working and studying managed ecosystems [@biosphere:hJ4c_Ua8].
+It was meant to be a conceptual tool that focussed attention on processes of destruction and reorganization, which had been neglected in comparison to those of growth and conservation.
+Considering these processes provides a more complex may to understand system dynamics.
+
+An adaptive cycle alternates between long periods of system aggregation, connection, and accumulation, and shorter periods of disruption and reorganization.
+The adaptive cycle exhibits two major phases.
+The first, often referred to as the front-loop, from r to K, is the slow, incremental phase of growth and accumulation.
+The second, referred to as the back-loop, from Omega to Alpha, is the rapid phase of reorganization leading to renewal.
+A system going into the back-loop can either remerge in a similar form, or may re-emerge as a new type of system with new boundaries and key components.
+Resilience researchers proposed that tensions between demands for organization and efficiency versus demands for novelty and diversity drive adaptive cycle type dynamics in many different types of complex systems
+[@Holling:2002wu].
 
@@
 
+\begin{figure*}[ht]
+\centering
+\includegraphics[width = 6in, height = 2.5in]{fig1_panarchy}
+\caption{\textbf{Panarchy} is an heuristic of nested adaptive cycles that serves to represent a variety of systems and environmental problems.
+Adaptive cycles (A) at different scales of the hierarchy (B) can be connected through remember and revolt cross-scale interactions.}
+\label{fig:fig1}
+\end{figure*}
+
 Panarchy posits that systems are organized in nested hierarchies across space and time, where each level of the hierarchy is a subsystem that can be in a different phase of the adaptive cycle.
-These phases are exploitation, conservation, release, and reorganization, the latter characterized by events of creative destruction (Fig 1).
+These phases are exploitation, conservation, release, and reorganization, the latter characterized by events of creative destruction (Fig.
+\ref{fig:fig1}).
 They are inspired by mathematical models used in economics and ecology, but broadly describe patterns of growth, collapse, and recovery that are common to populations, ecological communities, markets, or political organizations.
@@
 In that corner there is little space for experimentation and innovation.
-Examples include very conservative and bureaucratic systems.
+Examples include systems where ecological resilience has been expensively replaced by artificial processes to maintain the system such as levees, flood barriers, or chemical control of pests [@Holling:1996wx].
 
@@
 
-\begin{figure*}[h]
-\centering
-\includegraphics[width = 3in, height = 3in]{fig1_panarchy}
-\caption{\textbf{Panarchy} is an heuristic of nested adaptive cycles that serves to represent a variety of systems and environmental problems.
-Adaptive cycles at different scales of the hierarchy can be connected through remember and revolt cross-scale interactions.
-Reprinted with permission from Island Press}
-\label{fig:fig1}
-\end{figure*}
-
 # Methods
@@
 To answer these questions we combined an automated literature review based on topic modeling [@Blei:2012dk; @Griffiths:2004ey] with human-coded document analysis [@Bryman:2008va].
-All data and code to replicate the analysis are available at: [https://github.com/juanrocha/panarchy](https://github.com/juanrocha/panarchy) and a public online repository (add figshare link when finished).
+All data and code to replicate the analysis are available at: [https://github.com/juanrocha/panarchy](https://github.com/juanrocha/panarchy) and a public online repository ([10.6084/m9.figshare.13490919](10.6084/m9.figshare.13490919)).
 
@@
 We removed stop words (e.g.
-"the", "a") and digits from the matrix, as well as a list of words that were over represented in our data  and are common in the scientific literature but are unrelated to the papers’ topics(e.g.
+"the", "a") and digits from the matrix, as well as a list of words that were over represented in our data  and are common in the scientific literature but are unrelated to the papers’ topics (e.g.
 "paper", "study", "aim").
@@
 We used text annotations to capture potential criticisms as well as the methods used.
-The code book is available in the supplementary material (archived in XX repository).
+The code book is available in the supplementary material (archived in [10.6084/m9.figshare.13490919](10.6084/m9.figshare.13490919)).
 
@@
 
-_Panarchy_, the book published in 2002, has been cited over 7200 times according to Google Scholar.
+Panarchy, the book published in 2002, has been cited over 7200 times according to Google Scholar.
 The scientific paper that introduced the book to the scientific community [@Holling:2001cu] has received 1715 citations in the Web of Science, and 1923 in Scopus.
 Roughly half of the citations have come from environmental (28%) and social sciences (22%).
-Computer science (2.1%)and arts and humanities (2.6%) have been the least represented.
+Computer science (2.1%) and arts and humanities (2.6%) have been the least represented.
 
-A first exploration on the data suggest that content related to the adaptive cycle or general resilience has slightly decreased, while issues on development and management had gained attention (Fig 2).
-This simple exploration is not necessarily the best fit for our data, but serves to exemplify how topic models work.
-A topic is a set of words that are ranked according to a probability that they represent an underlying content of the document (Fig 1A).
-For example, the words "resilience", "systems", and "social" have a high probability to capture the content of topic 5.
-Similarly, there is a cluster of papers whose content is largely dominated by topic 1 on the adaptive cycle (gamma > 0.6), featuring it as one of the most active areas of research despite its relative decline in recent years (Fig 2).
+\begin{figure*}[ht]
+\centering
+\includegraphics[width = 5in, height = 4in]{fig2_topicnumber}
+\caption{\textbf{Algorithm and number of topic selection} Gibbs sampling maximizes entropy and log-likelihood estimation, making it a suitable algorithm for our data (A).
+Increasing the number of topics (from 5-100) shows that $\alpha$ decreases, suggesting that despite the larger number of topics, a few of them suffice to describe most papers (B).
+Log-likelihood is maximized for 50 topics followed closely with 25.
+While 50 topics is marginally better, for visualization purposes we restrict our analysis to 25.
+Note that perplexity cannot be calculated for Gibbs sampling, hence the missing value in A and absence in B.}
+\label{fig:algorithm}
+\end{figure*}
 
-Comparing algorithm performance revealed that Gibbs sampling was the best fit for the data.
+Comparing algorithm performance revealed that Gibbs sampling was the best fit for the data (Fig \ref{fig:algorithm}).
 As a rule of thumb, an ideal method should maximize entropy and the log-likelihood estimation while minimizing perplexity [@Grun:2011tb].
@@
 
-\begin{figure*}[h]
+\begin{figure*}[ht]
 \centering
-\includegraphics[width = 4in, height = 3.5in]{fig2_topicnumber}
-\caption{\textbf{Algorithm and number of topic selection} Gibbs sampling maximizes entropy and log-likelihood estimation, making it a suitable algorithm for our data (A).
-Increasing the number of topics (from 5-100) shows that $\alpha$ decreases, suggesting that despite the larger number of topics, a few of them suffice to describe most papers (B).
-Log-likelihood is maximized for 50 topics followed closely with 25.
-While 50 topics is marginally better, for visualization purposes we restrict our analysis to 25.
-Note that perplexity cannot be calculated for Gibbs sampling, hence the missing value in A and absence in B.}
-\label{fig:algorithm}
+\includegraphics[width = 5in, height = 5in]{fig3_topics25}
+\caption{\textbf{Panarchy topics over time} The number of papers per year (A) with a maximum of 324 in 2019.
+The proportion of papers per year (B) and the relative proportion of topic content per year (C) do not show strong trends for the time window with most papers (gray area in A).
+Each topic is summarized in (D) by the top 10-words that best describe them according to the posterior probability of our model fit.
+See FigS1 for a panel of time series per topics}
+\label{fig:topics25}
 \end{figure*}
 
-A topic is a set of words that are ranked according to a probability that they represent an underlying content of the document (Fig 3).
+A topic is a set of words that are ranked according to a probability that they represent an underlying content of the document (Fig \ref{fig:topics25}).
 For example, the words "resilience", "vulnerability", “risk” and "disaster" have a high probability to capture the content of topic 25.
@@
 Towards the end of the time series, topic 3 on urban systems and 9 on local communities and knowledge have spikes up to 9% of the content of each year (2019-2020).
-For comparison, if all topics were equally represented in the content, they would have 4% in the corpus (grey line in Fig 3B).
+For comparison, if all topics were equally represented in the content, they would have 4% in the corpus (grey line in Fig \ref{fig:topics25}B).
 
-Despite fluctuations, most topics showed a relatively constant level of interest over time (Fig 3, Fig S2).
+Despite fluctuations, most topics showed a relatively constant level of interest over time (Fig \ref{fig:topics25}, Fig \ref{fig:figS1}).
 We did not observe strong trends, but some topics have gained a small amount of attention.
@@
 
-\begin{figure*}[h]
-\centering
-\includegraphics[width = 5in, height = 5in]{fig3_topics25}
-\caption{\textbf{Panarchy topics over time} The number of papers per year (A) with a maximum of 324 in 2019.
-The proportion of papers per year (B) and the relative proportion of topic content per year (C) do not show strong trends for the time window with most papers (gray area in A).
-Each topic is summarized in (D) by the top 10-words that best describe them according to the posterior probability of our  model fit.
-See FigS2 for a panel of time series per topics}
-\label{fig:topics25}
-\end{figure*}
-
-The human-coded document analysis revealed that the most common feature of panarchy in the literature was the adaptive cycle (81%, N = 42) followed by cross-scale interactions (Fig 4).
+The human-coded document analysis revealed that the most common feature of panarchy in the literature was the adaptive cycle (81%, N = 42) followed by cross-scale interactions (Fig \ref{fig:qual}).
 Poverty and rigidity traps were less studied features in our sample, even when accounting for slightly different terminology such as "lock-ins".
@@
 
-\begin{figure*}[h]
+\begin{figure*}[ht]
 \centering
-\includegraphics[width = 6in, height = 3in]{fig4_qual_results}
-\caption{\textbf{Qualitative results} Document analysis was used to disentangle the different panarchy features addressed by a sub sample of papers (N = 42).
+\includegraphics[width = 5in, height = 2.5in]{fig4_qual_results}
+\caption{\textbf{Qualitative results} Document analysis was used to disentangle the different panarchy features addressed by a sub-sample of papers (N = 42).
 Over half of the papers are conceptual work, and most empirical papers fall into qualitative methods.}
@@
 # Discussion
-(1000w)
 
@@
 Topic modeling enabled us to complement the analysis to all papers reported in major scientific databases.
-It has the advantage of reproducibility, and reduces bias, but offers little insights about the dimensions of panarchy used, methods, or criticisms.
+It has the advantage of reproducibility and reduces sample bias, but offers little insights about the dimensions of panarchy used, methods, or criticisms.
 Here we show that both methods combined are a powerful combination.
-Future studies could benefit from including gray literature such as theses, non-governmental organizations reports, government agencies reports, or non-english literature.
+Future studies could benefit from including gray literature such as theses, books, non-governmental organizations reports, government agencies reports, or non-english literature.
 Previous studies have shown that the use of full text instead of abstracts can also offer additional insights on the automated analyses, for example in attributing impacts of ecosystem services from regime shifts in social-ecological systems [@Rocha:2015tk].
@@
 # Conclusion
-(400w)
 
@@
 Despite a growing body of literature, no topic seems to dominate the academic production of scholars using these concepts.
-The feature most used is the adaptive cycle, and the problems where it is found most useful is in studies with a historical perspective, either in short time horizons such as natural resource management, urban development, or conservation; all the way to archaeology and anthropology studies on the time horizon of centuries to millennia.
+The feature most used is the adaptive cycle, and the problems where it is found most useful is in studies with a historical perspective, either in short time horizons such as natural resource management, urban development, or conservation; all the way to archeology and anthropology studies on the time horizon of centuries to millennia.
 Hierarchies in scale or maladaptive traps have received comparatively less attention.
@@
 While our survey of the literature offers a few options to empirically ground panarchy concepts, we believe the problem of measurement and prediction are fertile ground for future research efforts.
+
+# Acknowledgements
+
+JCR acknowledges financial support from Formas research grant (942-2015-731), the Stockholm Resilience Centre, and Stockholm University.
 

### 2020-12-28 03-figueres.R

@@ -4,6 +4,7 @@ library(tm)
 library (topicmodels)
 library(lda)
 library(patchwork)
+library(png)
 
 
 load("data/models_gibbs.RData")
@@ -42,10 +43,13 @@ g1 <- df_stats %>%
     add_column(perplexity = as.numeric(perp[c(1,2,4,3)])) %>% #minimize perplexity
     pivot_longer(cols = 2:4, names_to = "measure", values_to = "value") %>%
     ggplot(aes(model, value)) + 
-    geom_col() + facet_wrap(.~measure, scales = "free") +
+    geom_col() + 
+    # scale_y_continuous(labels = scales::label_scientific) +
+    facet_wrap(.~measure, scales = "free_y") +
     labs(x = "Algorithm", y = "Value", tag = "A") +
-    theme_light(base_size = 6) 
-
+    theme_light(base_size = 8) + 
+    theme(axis.text.x = element_text(size = 5))
+g1
 # ggsave(filename = "figures/fig1_algorithm_selection.png",
 #        device = "png", width = 5, height = 2)
 
@@ -58,14 +62,15 @@ g2 <- df_topic_number %>%
     # filter(measure != "alpha") %>%
     ggplot(aes(as.factor(topic_number), value)) +
     geom_col() + 
+    # scale_y_continuous(labels = scales::label_scientific) +
     labs(x = "Number of topics", y = "Value", tag = "B") +
     facet_wrap(.~measure, scales = "free", ncol = 4, nrow = 1) +
-    theme_light(base_size = 6)
+    theme_light(base_size = 8)
 
 g1/g2
 
-ggsave(filename = "paper/figures/fig3_topicnumber.png",
-       device = "png", width = 5, height = 4)
+ggsave(filename = "paper/figures/fig2_topicnumber.png",
+       device = "png", width = 5, height = 4, dpi = 400)
 
 
 #### heatmaps topics ####
@@ -98,7 +103,7 @@ g_25 <- df_topics25 %>%
     theme(axis.text.x = element_text(size = 5))
 
 
-
+g_25
 
 ggsave("fig3_topics25_time.png", device = "png", width = unit(5,"in"), height = unit(5,"in"))
 
@@ -162,7 +167,7 @@ g_prop <- df_documents %>%
     geom_line(aes(color = as.factor(topic)), show.legend = FALSE, size = 0.25) +
[…]

### 2021-05-19 paper/paper_panarchy.Rmd
revised version addressing reviewer comments, plus typo correction on figures

@@
 
-Almost two decades ago, the edited book “Panarchy: Understanding Transformations in Human and Natural Systems” [@Gunderson:2002vk], presented a synthetic perspective of how a group of leading social-ecological researchers associated with the Resilience Alliance understood change in social-ecological systems.
-The concept of Panarchy was a key focus on this influential book.
-In the late 1990s and early 2000s, the Resilience Alliance was a productive, innovative highly collaborative group of environmental scientists that was focussed on bridging social and natural sciences as well as theory and practice.
-They were very focussed on addressing social-ecological problems by combining insights from the social and natural sciences, as well as the arts and humanities [@Parker:2012eo].
+Almost two decades ago, the edited book "Panarchy: Understanding Transformations
+in Human and Natural Systems" [@Gunderson:2002vk], presented a synthetic
+perspective of how a group of social-ecological researchers associated with the
+Resilience Alliance understood change in social-ecological systems. The concept
+of Panarchy was a key focus on this influential book. In the late 1990s and
+early 2000s, the Resilience Alliance was a productive, innovative, and highly
+collaborative group of interdisciplinary scientists that focused on bridging
+social and natural sciences as well as theory and practice. They were addressing
+social-ecological problems by combining insights from the social and natural
+sciences, as well as the arts and humanities [@Parker:2012eo].
 
-Panarchy encapsulates a set of concepts that have inspired the work of environmental scientists and practitioners for almost two decades.
-It was a central concept produced by the Resilience Alliance, an international group of influential academics focused on the resilience of social-ecological systems in theory and practice [@Parker:2012eo].
-The Panarchy concept builds on Holling’s adaptive cycle [@biosphere:hJ4c_Ua8] by extending the idea across spatial and temporal scales.
-
-Panarchy remains a boundary object that has inspired research topics, enabled collaborations and nurtured new scientific frameworks.
-The ideas put forward have been applied in field studies, archaeology, mathematical models, participatory work, and scenario development.
-Panarchy has inspired resilience assessments and guided decision making.
-In this article we pay tribute to the book by studying how the concepts and metaphors proposed have been further developed in the academic literature.
+The Panarchy concept builds on Holling's adaptive cycle [@biosphere:hJ4c_Ua8] by
+extending the idea across spatial and temporal scales. It remains a boundary
+object that has inspired research topics, enabled collaborations and nurtured
+new scientific frameworks. The ideas put forward have been applied in field
+studies, archaeology, mathematical models, participatory work, and scenario
+development. Panarchy has inspired resilience assessments and guided decision
+making. In this article we pay tribute to the book by studying how the concepts
+and metaphors proposed have been further developed in the academic literature.
 We also document criticism of the concepts and identify key research frontiers.
 
-Panarchy proposes that it is useful to conceptualize systems in terms of interacting adaptive cycles.
-The adaptive cycle was an idea first proposed by Holling based upon his experience working and studying managed ecosystems [@biosphere:hJ4c_Ua8].
-It was meant to be a conceptual tool that focussed attention on processes of destruction and reorganization, which had been neglected in comparison to those of growth and conservation.
-Considering these processes provides a more complex may to understand system dynamics.
+Panarchy proposes that it is useful to conceptualize systems in terms of
+interacting adaptive cycles. The adaptive cycle was an idea first proposed by
+Holling based upon his experience working and studying managed ecosystems
+[@biosphere:hJ4c_Ua8]. It was meant to be a conceptual tool that focused
+attention on processes of destruction and reorganization, which had been
+neglected in comparison to those of growth and conservation. Considering these
+processes provides a more complex way to understand system dynamics.
 
-An adaptive cycle alternates between long periods of system aggregation, connection, and accumulation, and shorter periods of disruption and reorganization.
-The adaptive cycle exhibits two major phases.
-The first, often referred to as the front-loop, from r to K, is the slow, incremental phase of growth and accumulation.
-The second, referred to as the back-loop, from Omega to Alpha, is the rapid phase of reorganization leading to renewal.
-A system going into the back-loop can either remerge in a similar form, or may re-emerge as a new type of system with new boundaries and key components.
-Resilience researchers proposed that tensions between demands for organization and efficiency versus demands for novelty and diversity drive adaptive cycle type dynamics in many different types of complex systems
-[@Holling:2002wu].
+An adaptive cycle alternates between long periods of system aggregation,
+connection, and accumulation, and shorter periods of disruption and
+reorganization. The adaptive cycle exhibits four phases arranged into two major
+loops (Fig. \ref{fig:fig1}). The phases are exploitation ($r$), conservation
+($K$), release ($\Omega$), and reorganization ($\alpha$), the latter
+characterized by events of creative destruction (Fig. \ref{fig:fig1}). The first
+loop, often referred to as the front-loop, from $r$ to $K$, is the slow,
+incremental phase of growth and accumulation. The second loop, referred to as
+the back-loop, from $\Omega$ to $\alpha$, is the rapid phase of reorganization
+leading to renewal. A system going into the back-loop can either remain in a
+similar form, or may transition to a new type of system with new boundaries and
+key components [@Holling:2002wu]. Resilience researchers proposed that tensions
+between demands for organization and efficiency versus demands for novelty and
+diversity drive adaptive cycle type dynamics in many different types of complex
+systems [@Holling:2002wu].
 
-Panarchy is not a theory of what it is, but a metaphor of what might be [@Gunderson:2002vk].
-It is not a predictive tool, but aims to understand adaptive change.
-The adaptive cycle is hypothesized to exist within a three-dimensional space defined by three properties: potential, connectivity, and resilience [@Gunderson:2002vk].
-Potential refers to the capital available capital to the system, for example nutrients and carbon captured by a forest, or human capital - skills and knowledge accumulated to run the economy.
-Connectivity is a proxy of the structure of the system; it is the network of interactions and strength between its elements.
-Resilience is the capacity of any system to absorb disturbance and reorganize while undergoing change so as to retain essentially the same identity, as determined by its function, structure, and feedbacks [@Anonymous:2016fv].
+Panarchy is not a theory of what it is, but a metaphor of what might be
+[@Gunderson:2002vk]. It is not a predictive tool, but aims to understand
+adaptive change. The adaptive cycle is hypothesized to exist within a
+three-dimensional space defined by three properties: potential, connectivity,
+and resilience [@Gunderson:2002vk]. Potential refers to the capital available to
+the system, for example nutrients and carbon captured by a forest, or human
+capital - skills and knowledge accumulated to run the economy. Connectivity is a
+proxy of the structure of the system; it is the network of interactions and
+strength between its elements. Resilience is the capacity of any system to
+absorb disturbance and reorganize while undergoing change so as to retain
+essentially the same identity, as determined by its function, structure, and
+feedbacks [@Anonymous:2016fv].
 
+```{=tex}
 \begin{figure*}[ht]
@@
 \end{figure*}
+```
+Panarchy posits that systems are organized in nested hierarchies across space
+and time, where each level of the hierarchy is a subsystem that can be in a
+different phase of the adaptive cycle. These phases are inspired by mathematical
+models used in economics and ecology, but broadly describe patterns of growth,
+collapse, and recovery that are common to populations, ecological communities,
+markets, or political organizations. While each subsystem in the hierarchy can
+be at a different phase of the adaptive cycle, these subsystems can influence
+each other through cross-scale interactions called *"revolt"* and *"remember"*
+[@Gunderson:2002vk].
 
-Panarchy posits that systems are organized in nested hierarchies across space and time, where each level of the hierarchy is a subsystem that can be in a different phase of the adaptive cycle.
-These phases are exploitation, conservation, release, and reorganization, the latter characterized by events of creative destruction (Fig.
-\ref{fig:fig1}).
-They are inspired by mathematical models used in economics and ecology, but broadly describe patterns of growth, collapse, and recovery that are common to populations, ecological communities, markets, or political organizations.
-While each subsystem in the hierarchy can be at a different phase of the cycle, these subsystems can influence each other through cross-scale interactions called “revolt” and “remember.”
+The three-dimensional space has corners with attractors that can cause the
+adaptive cycle to stagnate: poverty and rigidity traps [@Gunderson:2002vk].
+Poverty traps are described in panarchy as maladaptive states where potential,
+connectivity and resilience are low. Poverty traps are a series of feedback
+mechanisms that reinforce impoverished states [@Bowles:2006hx; @Maru:2012dm;
+@Allison:2004te], limiting the system's capacity to innovate and increase
+potential. The opposite corner, where potential, connectivity and resilience are
+high, is another maladaptive space called a rigidity trap. In that corner there
+is little space for experimentation and innovation. Examples include systems
+where ecological resilience has been expensively replaced by artificial
+processes to maintain the system such as levees, flood barriers, or chemical
+control of pests [@Holling:1996wx].
 
-The three-dimensional space has corners with attractors that can cause the adaptive cycle to stagnate: poverty and rigidity traps [@Gunderson:2002vk].
-Poverty traps are described in panarchy as maladaptive states where potential, connectivity and resilience are low.
-Poverty traps are a series of feedback mechanisms that reinforce impoverished states [@Bowles:2006hx], limiting the system’s capacity to innovate and increase potential.
-The opposite corner, where potential, connectivity and resilience are high, is another maladaptive space called a rigidity trap.
-In that corner there is little space for experimentation and innovation.
-Examples include systems where ecological resilience has been expensively replaced by artificial processes to maintain the system such as levees, flood barriers, or chemical control of pests [@Holling:1996wx].
+Panarchy offers a rich conceptual framework for understanding environmental
+problems. While inspired by a few mathematical constructs (e.g. cycles, traps,
+scaling laws), it is general enough to invite scholars from multiple
+disciplinary backgrounds, ontologies and epistemologies, to collaborate around
+research questions and applied problems. As such, it is useful as a boundary
+object and can be used empirically or metaphorically. The book itself presents a
+series of case studies where geographers, economists, political scientists, and
+ecologists have demonstrated the utility of the framework to their area of
+research. However, the extent to which these concepts have gained empirical
+support beyond the metaphor, or how they have evolved over time, remain open
+questions. For example, the adaptive cycle is a useful metaphor to look back at
+history and organize events and periods around phases, but it remains
+challenging to identify the phase of a system in real time, or to use panarchy
+to project future trajectories.
 
-Panarchy offers a rich conceptual framework for understanding environmental problems.
-While inspired by a few mathematical constructs (e.g.
-cycles, traps, scaling laws), it is general enough to invite scholars from multiple disciplinary backgrounds, ontologies and epistemologies, to collaborate around research questions and applied problems.
-As such, it is useful as a boundary object  and can be used empirically or metaphorically.
-The book itself presents a series of case studies where geographers, economists, political scientists, and ecologists have demonstrated the utility of the framework to their area of research.
-However, the extent to which these concepts have gained empirical support beyond the metaphor, or how they have evolved over time, remain open questions.
-For example, the adaptive cycle is a useful metaphor to look back at history and organize events and periods around phases, but it remains challenging to identify the phase of a system in real time, or to use panarchy to project future trajectories.
-In other words, panarchy is useful for retrospective studies, but its applications may be more limited for prospective ones.
-Are we at a stage where we can start distilling theory that enables case study comparisons, deriving functional forms, or draw theoretical expectations and predictions?
-
-Here we review the academic literature of the last two decades to trace how these concepts have evolved, in what type of problems they have been applied and found useful, and finally what remains as the key frontiers of research.
+Here we review the academic literature of the last two decades to trace how
+these concepts have evolved, in what type of problems they have been applied and
+found useful, and finally what remains as the key frontiers of research.
 
@@
 
-To answer these questions we combined an automated literature review based on topic modeling [@Blei:2012dk; @Griffiths:2004ey] with human-coded document analysis [@Bryman:2008va].
-All data and code to replicate the analysis are available at: [https://github.com/juanrocha/panarchy](https://github.com/juanrocha/panarchy) and a public online repository ([10.6084/m9.figshare.13490919](10.6084/m9.figshare.13490919)).
+To answer these questions we combined an automated literature review based on
+topic modeling [@Blei:2012dk; @Griffiths:2004ey] with human-coded document
+analysis [@Bryman:2008va].
 
-*Data*: We used the Web of Science, Scopus, and GoogleScholar to survey academic literature that has used or referenced works that trace back to Panarchy [@Gunderson:2002vk].
-We extracted complete records from the Scopus database that matched the search for "(panarchy OR adaptive cycle)" (N = 595), or the search "(panarchy OR adaptive cycle) AND resilience" (N = 278).
-The data was combined with all papers (N = 1923) that cited the inaugural paper that introduced the book to the academic community [@Holling:2001cu].
-Records with missing abstracts were dropped (N = 191), and records with missing year were set to 2020 given that they are accepted manuscripts with digital object identifier (doi) scheduled to be published later in 2020 or 2021.
+*Data*: We used the Web of Science, Scopus, and GoogleScholar to survey academic
+literature that has used or referenced works that trace back to Panarchy
+[@Gunderson:2002vk]. We extracted complete records from the Scopus database that
+matched the search for *"panarchy OR adaptive cycle"* (N = 595), or the search
+*"(panarchy OR adaptive cycle) AND resilience"* (N = 278). The data was combined
+with all papers (N = 1923) that cited the inaugural paper that introduced the
+book to the academic community [@Holling:2001cu]. Records with missing abstracts
+were dropped (N = 191), and records with missing year were set to 2020 given
+that they are accepted manuscripts with digital object identifier (doi)
+scheduled to be published later in 2020 or 2021.
 
-To prepare the data for topic modeling, we constructed a document term matrix with documents in rows and words in columns.
-Here our unit of analysis for the document are the abstracts retrieved, and the matrix contains the count of words per document.
-We removed stop words (e.g.
-"the", "a") and digits from the matrix, as well as a list of words that were over represented in our data  and are common in the scientific literature but are unrelated to the papers’ topics (e.g.
-"paper", "study", "aim").
+To prepare the data for topic modeling, we constructed a document term matrix
+with documents in rows and words in columns. Here our unit of analysis for the
+document are the abstracts retrieved, and the matrix contains the count of words
+per document. We removed stop words (e.g. "the", "a") and digits from the
+matrix, as well as a list of words that were over represented in our data and
+are common in the scientific literature but are unrelated to the papers' topics
+(e.g. "paper", "study", "aim").
 
-*Topic models*: are an unsupervised statistical technique to reduce the dimensionality of a corpus of data (typically but not necessarily text) into topics [@Blei:2012dk].
-Here a topic is a latent variable that ranks words with high probability of appearing together within the same document.
-Documents in turn can be described by the probability distribution of a particular set of topics.
-Since they are (posterior) probabilities, the sum of the probability of all words for any given topic should be one, and the sum of the probability of all topics for any given document should also be one.
-An iterative process or algorithm is what allows the model to learn the ranking of words that best explain topics, and the ranking of topics that best explain documents.
+*Topic models*: are an unsupervised statistical technique to reduce the
+dimensionality of a corpus of data (typically but not necessarily text) into
+topics [@Blei:2012dk]. Here a topic is a latent variable that ranks words with
+high probability of appearing together within the same document. Documents in
+turn can be described by the probability distribution of a particular set of
+topics. Since they are (posterior) probabilities, the sum of the probability of
+all words for any given topic should be one, and the sum of the probability of
+all topics for any given document should also be one. An iterative process or
+algorithm is what allows the model to learn the ranking of words that best
+explain topics, and the ranking of topics that best explain documents.
 
-The underlying statistical technique for this machine learning approach is called Latent Dirichlet Allocation (LDA) [@Blei:2003tn].
-It allocates probabilities to latent variables (topics) based on the distribution of words in text data, assuming a multivariate continuous (Dirichlet) distribution.
-We compared three LDA algorithms: correlated topic models (CTM), variational expectation maximization (VEM), and Gibbs sampling (Gibbs), by assessing their performance against their log-likelihood estimation, entropy and perplexity [@Grun:2011tb].
-Entropy is a measure of order or disorder of a system; in the context of topic models it measures how evenly the topic distribution is spread.
-Perplexity measures the uncertainty of predicting a single word, so if the model performance were the same as random, perplexity would approximate the vocabulary size (16 728 words).
-These metrics enable us to choose which algorithm best fits our data, what is the optimal number of topics to fit, and how to avoid overfitting.
+The underlying statistical technique for this machine learning approach is
+called Latent Dirichlet Allocation (LDA) [@Blei:2003tn]. It allocates
+probabilities to latent variables (topics) based on the distribution of words in
+text data, assuming a multivariate continuous (Dirichlet) distribution. We
+compared three LDA algorithms: correlated topic models (CTM), variational
+expectation maximization (VEM), and Gibbs sampling (Gibbs), by assessing their
+performance against their log-likelihood estimation, entropy and perplexity
+[@Grun:2011tb]. Entropy is a measure of order or disorder of a system; in the
+context of topic models it measures how evenly the topic distribution is spread.
+Perplexity measures the uncertainty of predicting a single word, so if the model
+performance were the same as random, perplexity would approximate the vocabulary
+size (N = 16 728 words). These performance metrics enable us to choose which
+algorithm best fits our data, what is the optimal number of topics to fit, and
+how to avoid overfitting.
 
-*Document analysis*: We complemented our unsupervised approach with document analysis [@Bryman:2008va] by coding an additional set of categorical variables for a sample of highly cited papers (N = 41).
-We annotated qualitative aspects such as use of the adaptive cycle, identification of its phases, whether the paper is conceptual, modeling or an empirical study.
-When empirical, we record the temporal and spatial scales of the case study.
-We also identified what aspects of panarchy were most used in the papers.
-For example whether there is emphasis on cross-scale interactions, or poverty and rigidity traps.
-We used text annotations to capture potential criticisms as well as the methods used.
-The code book is available in the supplementary material (archived in [10.6084/m9.figshare.13490919](10.6084/m9.figshare.13490919)).
+*Document analysis*: We complemented our unsupervised approach with document
+analysis [@Bryman:2008va] by coding an additional set of categorical variables
+for a sample of papers (N = 42). Based on the reading of Panarchy, we first
+developed a series of categorical variables that captured the main components of
+the scientific framework. The variables were presented and received feedback at
+the science meeting of the Resilience Alliance in 2019. We used the variables to
+annotate qualitative aspects such as use of the adaptive cycle, identification
+of its phases, whether the paper is conceptual, modeling or an empirical study.
+When empirical, we recorded the temporal and spatial scales of the case study.
+We also identified what aspects of panarchy were most used in the papers. For
+example whether there is emphasis on cross-scale interactions, or poverty and
+rigidity traps. We used text annotations to capture potential criticisms as well
+as the methods used.
 
@@
 
-Panarchy, the book published in 2002, has been cited over 7200 times according to Google Scholar.
-The scientific paper that introduced the book to the scientific community [@Holling:2001cu] has received 1715 citations in the Web of Science, and 1923 in Scopus.
-Roughly half of the citations have come from environmental (28%) and social sciences (22%).
-Computer science (2.1%) and arts and humanities (2.6%) have been the least represented.
+Panarchy, the book published in 2002, has been cited over 7200 times according
+to Google Scholar. The scientific paper that introduced the book to the
+scientific community [@Holling:2001cu] has received, at the time of writing,
+1715 citations in the Web of Science, and 1923 in Scopus. Roughly half of the
+citations have come from environmental (28%) and social sciences (22%). Computer
+science (2.1%) and arts and humanities (2.6%) have been the least represented.
 
+```{=tex}
 \begin{figure*}[ht]
@@
 \end{figure*}
+```
+Comparing algorithm performance revealed that Gibbs sampling was the best fit
+for the data (Fig \ref{fig:algorithm}). As a rule of thumb, an ideal method
+should maximize entropy and the log-likelihood estimation while minimizing
+perplexity [@Grun:2011tb]. Gibbs sampling maximised both entropy and likelihood
+with our data when compared to other alternatives. The second best was the
+variational-expectation maximisation algorithm (VEM) when $\alpha$ was not set
+constant. $\alpha$ is a hyper parameter that weights the evenness of topic
+distribution. A lower $\alpha$ than default values indicates that the documents
+can be described by rather fewer topics, or that its distribution is very
+uneven. In fact, we observed that increasing the number of topics from 5 to 100
+topics did increase entropy at expense of reducing $\alpha$, meaning that
+despite the larger number of topics, the main content of a document is still
+captured by a few of them. The log-likelihood maximization stopped at around
+25-50 topics. Thus, we restricted the rest of our analysis to 25 topics.
 
-Comparing algorithm performance revealed that Gibbs sampling was the best fit for the data (Fig \ref{fig:algorithm}).
-As a rule of thumb, an ideal method should maximize entropy and the log-likelihood estimation while minimizing perplexity [@Grun:2011tb].
-Gibbs sampling maximised both entropy and likelihood with our data when compared to other alternatives.
-The second best was the variational-expectation maximisation algorithm (VEM) when $\alpha$ was not set constant.
-$\alpha$ is a hyper parameter that weights the evenness of topic distribution.
-A lower $\alpha$ than default values indicates that the documents can be described by rather fewer topics, or that its distribution is very uneven.
-In fact, we observed that increasing the number of topics from 5 to 100 topics did increase entropy at expense of reducing $\alpha$, meaning that despite the larger number of topics, the main content of a document is still captured by a few of them.
-The log-likelihood maximization stopped at around 25-50 topics.
-Thus, we restricted the rest of our analysis to 25 topics.
-
+```{=tex}
 \begin{figure*}[ht]
@@
 \end{figure*}
+```
+A topic is a set of words that are ranked according to a probability that they
+represent an underlying content of the document (Fig \ref{fig:topics25}). For
+example, the words "resilience", "vulnerability", "risk" and "disaster" have a
+high probability to capture the content of topic 25. Papers early in our time
+series (2001-3) have high content largely dominated by topics 25 on resilience,
+24 on the adaptive cycle, and 18 on adaptive governance and management. Towards
+the end of the time series, topic 3 on urban systems and 9 on local communities
+and knowledge have spikes up to 9% of the content of each year (2019-2020). For
+comparison, if all topics were equally represented in the content, they would
+have 4% in the corpus (grey line in Fig \ref{fig:topics25}B).
 
-A topic is a set of words that are ranked according to a probability that they represent an underlying content of the document (Fig \ref{fig:topics25}).
-For example, the words "resilience", "vulnerability", “risk” and "disaster" have a high probability to capture the content of topic 25.
-Papers early in our time series (2001-3) have high content largely dominated by topics 25 on resilience, 24 on the adaptive cycle, and 18 on adaptive governance and management.
-Towards the end of the time series, topic 3 on urban systems and 9 on local communities and knowledge have spikes up to 9% of the content of each year (2019-2020).
-For comparison, if all topics were equally represented in the content, they would have 4% in the corpus (grey line in Fig \ref{fig:topics25}B).
+Despite fluctuations, most topics showed a relatively constant level of interest
+over time (Fig \ref{fig:topics25}, Fig \ref{fig:figS1}). We did not observe
+strong trends, but some topics have gained a small amount of attention. For
+example, topic 1 on sustainability, or topic 10 on ecosystem services appeared
+consistently across time. In contrast, research on business innovation (topic
+20), urban infrastructure (topic 3), or archaeology (topic 5) have gained
+attention in recent years. Topic 8 was an outlier, with a selection of papers
+ranking high on content related to network infrastructure and performance,
+possibly from engineering disciplines. It was the only topic with a set of
+papers that was clearly distinct from the rest of the collection, and showed a
+decreasing trend over time.
 
-Despite fluctuations, most topics showed a relatively constant level of interest over time (Fig \ref{fig:topics25}, Fig \ref{fig:figS1}).
-We did not observe strong trends, but some topics have gained a small amount of attention.
-For example, topic 1 on sustainability, or topic 10 on ecosystem services appeared consistently across time.
-In contrast, research on business innovation (topic 20), urban infrastructure (topic 3), or archaeology (topic 5) have gained attention in recent years.
-Topic 8 was an outlier, with a selection of papers ranking high on content related to network infrastructure and performance, possibly from engineering disciplines.
-It was the only topic with a set of papers that was clearly distinct from the rest of the collection, and showed a decreasing trend over time.
+The human-coded document analysis revealed that the most common feature of
+panarchy in the literature was the adaptive cycle (81%, N = 42) followed by
+cross-scale interactions (Fig \ref{fig:qual}). Poverty and rigidity traps were
+less studied features in our sample (14% and 19% respectively), even when
+accounting for slightly different terminology such as "lock-ins". The bulk of
+the papers analyzed were conceptual papers, many without a method or discernible
+research question. Roughly half of the papers were conceptual or only used
+panarchy as a metaphor. Over 40% of papers in our sample went a step further and
+used panarchy as an empirical construct, for example by attempting to identify
+the phases of the adaptive cycle (61%). Out of the 22 empirical cases, 6 were at
+the time scale of centuries, 7 on decades, 6 on years, and 1 in weeks.
+Spatially, 4 were at a city scale, 17 regional, and 1 national. Empirical papers
+were dominated by qualitative methods (77%, N = 22) and were generally
+retrospective historical reviews.
 
-The human-coded document analysis revealed that the most common feature of panarchy in the literature was the adaptive cycle (81%, N = 42) followed by cross-scale interactions (Fig \ref{fig:qual}).
-Poverty and rigidity traps were less studied features in our sample, even when accounting for slightly different terminology such as "lock-ins".
-The bulk of the papers analyzed were conceptual papers, many without a method or discernible research question.
-Roughly half of the papers were conceptual or only used panarchy as a metaphor.
-Over 40% of papers in our sample went a step further and used panarchy as an empirical construct, for example by attempting to identify the phases of the adaptive cycle (61%).
-Out of the 22 empirical cases, 6 were at the time scale of centuries, 7 on decades, 6 on years, and 1 in weeks.
-Spatially, 4 were at a city scale, 17 regional, and 1 national.
-Empirical papers were dominated by qualitative methods (77%, N = 22) and were generally retrospective historical reviews.
+Several of the papers reviewed attempted to identify the adaptive cycle either
+through qualitative or quantitative methods. Identification of adaptive cycles
+has found applications in a wide range of disciplines and research problems,
+from delimitation of periods in archaeology and anthropology [@Redman:2003bd],
+to financial crises in Europe [@Castell:2020dg], or traffic jams in China
+[@Zeng:2020hu]. Recent empirical tests of the adaptive cycle innovate on the
+types of data and methods that one can use. Information transfer methods based
+on entropy have been suggested to approximate relevant components of a system
+and the empirical proxies of the axis where the adaptive cycle unfolds:
+potential, connectedness, and resilience [@Castell:2020dg]. Percolation methods
+combined with big data have been shown to be useful to test hypotheses of regime
+shifts in urban systems (originally proposed in Panarchy [@Gunderson:2002vk]),
+and derive the temporal and spatial scales at which the adaptive cycle emerges
+[@Zeng:2020hu].
 
-The high level of conceptualization but lack of empirically grounded hypotheses or theory testing came across as one of the major limitations.
-Nonetheless, several of the papers reviewed attempted to identify the adaptive cycle either through qualitative or quantitative methods.
-Identification of adaptive cycles has found applications in a wide range of disciplines and research problems, from delimitation of periods in archaeology and anthropology [@Redman:2003bd], to financial crises in Europe [@Castell:2020dg], or traffic jams in China [@Zeng:2020hu].
-Recent empirical tests of the adaptive cycle innovate on the types of data and methods that one can use.
-Information transfer methods based on entropy have been suggested to approximate relevant components of a system and the empirical proxies of the axis where the adaptive cycle unfolds: potential, connectedness, and resilience [@Castell:2020dg].
-Percolation methods combined with big data have been shown to be useful to test hypotheses of regime shifts in urban systems (originally proposed in Panarchy [@Gunderson:2002vk]), and derive the temporal and spatial scales at which the adaptive cycle emerges [@Zeng:2020hu].
-
+```{=tex}
 \begin{figure*}[ht]
@@
 \end{figure*}
-
+```
 # Discussion
 
-Given the strong dominance of conceptual papers, it is not surprising that one of the key frontiers of research is to empirically ground such conceptualizations in a way that one can test hypotheses and advance theory beyond metaphors.
-Metaphors have played an important role as boundary objects enabling interdisciplinary dialogues.
-We observe it on the wide applications that panarchy has inspired, from anthropology to engineering.
-In the empirical realm, the adaptive cycle is the panarchy dimension that researchers engage the most with, but it is dominated by qualitative and retrospective studies.
-That means, we still lack the theory and methods to gather observations and be able to decide in which phase of the adaptive cycle a system currently is, or what is the probability of it transitioning to a new phase?
+The high level of conceptualization but lack of empirically grounded theory
+testing came across as one of the major limitations. Here we do not refer only
+to positivism as a plausible way of theory development. Other equally valid
+approaches exist. From the philosophical point of view of what makes a
+theory a scientific theory, the plurality of approaches should have mechanisms
+in place to avoid confirmation bias, allow the emergence of alternative
+explanations, and identify the circumstances under which a theory may fail
+[@Hansson:2017tr]. Many of the papers we read confirm but not challenged the
+expectations that Panarchy put forward. A few of the most recent work actually
+falsify some of the ideas proposed --e.g. the back loop is faster that the front
+one. We believe more work is needed to better understand the circumstances where
+Panarchy works, and when it falls short. In other words, refine the scientific
+framework towards a scientific theory.
 
-The criticism of over conceptualization is not unique to the literature engaging Panarchy ideas.
-A recent review of sustainability science mapped the different schools of thought that the discipline has developed over the last decades [@Clark:2020jd].
-A similar conclusion was reached, where too many conceptual frameworks have been developed, but fewer empirical attempts try to test the frameworks against data to falsify hypotheses.
-The review emphasizes the problem of measurement and observation [@Clark:2020jd].
-In the context of panarchy, recent work has developed an information theory based approach that enables the identification of adaptive cycles [@Castell:2020dg].
-The authors identify phases of the adaptive cycle  in the European financial crises and in grassland ecosystems, but fail to find support for the difference in speed between the forward and backward loops originally proposed in Panarchy.
+Given the strong dominance of conceptual papers, it is not surprising that one
+of the key frontiers of research is to empirically ground such
+conceptualizations in a way that one can test theoretical expectations beyond
+metaphors. Metaphors have played an important role as boundary objects enabling
+interdisciplinary dialogues. We observe it on the wide applications that
+panarchy has inspired, from anthropology to engineering. In the empirical realm,
+the adaptive cycle is the panarchy dimension that researchers engage the most
+with, but it is dominated by qualitative and retrospective studies. That means,
+we still lack the theory and methods to gather observations and be able to
+decide in which phase of the adaptive cycle a system currently is, or what is
+the probability of it transitioning to a new phase?
 
-The panarchy dimensions that received less attention in our qualitative analysis are creative destruction, rigidity and poverty traps.
-This may be at least partly due to the scope of our data: papers that have cited Panarchy or Holling's [@Gunderson:2002vk; @Holling:2001cu].
-Concepts such as creative destruction and poverty traps precede Panarchy and therefore have been theorized and empirically grounded outside the panarchy stream of thinking.
-For example, the theory of poverty traps dates back to the 1950's in economics, and has received both theoretical development [@Bowles:2006hx] as well as empirical grounding [@Banerjee:2015ja; @Banerjee:2012tq] that has enabled researchers and governments to distinguish what kind of interventions are likely to reduce poverty.
-Here we encourage similar empirical efforts in formalizing theory and empirical support for the adaptive cycles, their cross-scale interactions (remember and revolt), or how the hierarchical nested nature of complex systems enhance or erode resilience in social-ecological systems.
+The criticism of over-conceptualization is not unique to the literature engaging
+Panarchy ideas. Here by over-conceptualization we refer to the problem of
+scientific jargon, where different communities use slightly different wordings
+to refer to the same phenomena; reducing the chances that lessons from a
+disciplines percolates to another. An example from Panarchy is the endorsement
+of ideas such as creative destruction (from economics) or the theory of
+intermediate disturbances (from ecology). The latter has been debunked in the
+ecological literature but might still been used within Panarchy thinking
+[@Fox:2013br]. A recent review of sustainability science mapped the different
+schools of thought that the discipline has developed over the last decades
+[@Clark:2020jd]. A similar conclusion was reached, where too many conceptual
+frameworks have been developed, but fewer empirical attempts try to test the
+frameworks against data to falsify hypotheses. The review emphasizes the problem
+of measurement and observation [@Clark:2020jd]. In the context of panarchy,
+recent work has developed an information theory based approach that enables the
+identification of adaptive cycles [@Castell:2020dg]. The authors identify phases
+of the adaptive cycle in the European financial crises and in grassland
+ecosystems, but fail to find support for the difference in speed between the
+forward and backward loops originally proposed in Panarchy.
 
-This paper aimed to study the use of panarchy related concepts and their evolution since the publication of the book.
-Systematic literature reviews often suffer from the limitation of restricting sample size to a subset of what is readable on the time frame of a project -- between dozens and maybe hundreds of papers.
-Qualitative analysis offers rich insights into the papers but is limited by sample size.
-In our analysis, the qualitative insights could be biased towards a non-random selection of papers that aligned our research interests, or highly cited, review type of work.
-Topic modeling enabled us to complement the analysis to all papers reported in major scientific databases.
-It has the advantage of reproducibility and reduces sample bias, but offers little insights about the dimensions of panarchy used, methods, or criticisms.
-Here we show that both methods combined are a powerful combination.
-Future studies could benefit from including gray literature such as theses, books, non-governmental organizations reports, government agencies reports, or non-english literature.
-Previous studies have shown that the use of full text instead of abstracts can also offer additional insights on the automated analyses, for example in attributing impacts of ecosystem services from regime shifts in social-ecological systems [@Rocha:2015tk].
-Replication attempts do however face the challenge of accessing full text when many of the papers are behind pay-walls.
+The panarchy dimensions that received less attention in our qualitative analysis
+are creative destruction, rigidity and poverty traps. This may be at least
+partly due to the scope of our data: papers that have cited Panarchy or
+Holling's paper [@Gunderson:2002vk; @Holling:2001cu]. Concepts such as creative
+destruction and poverty traps precede Panarchy and therefore have been theorized
+and empirically grounded outside the panarchy stream of thinking. For example,
+the theory of poverty traps dates back to the 1950's in economics, and has
+received both theoretical development [@Bowles:2006hx] as well as empirical
+grounding [@Banerjee:2015ja; @Banerjee:2012tq] that has enabled researchers and
+governments to distinguish what kind of interventions are likely to reduce
+poverty. Here we encourage similar empirical efforts in formalizing theory and
+empirical support for the adaptive cycles, their cross-scale interactions
+(remember and revolt), or how the hierarchical nested nature of complex systems
+enhance or erode resilience in social-ecological systems.
+
+This paper aimed to study the use of panarchy related concepts and their
+evolution since the publication of the book. Systematic literature reviews often
+suffer from the limitation of restricting sample size to a subset of what is
+readable on the time frame of a project -- between dozens and maybe hundreds of
+papers. Qualitative analysis offers rich insights into the papers but is limited
+by sample size. In our analysis, the qualitative insights could be biased
+towards a non-random selection of papers that aligned our research interests, or
+highly cited, review type of work. Topic modeling enabled us to complement the
+analysis to all papers reported in major scientific databases. It has the
+advantage of reproducibility and reduces sample bias, but offers little insights
+about the dimensions of panarchy used, methods, or criticisms. Here we show that
+both methods combined are a powerful combination.
+
+Our choice of methods also has limitations. We could not compare both methods
+since they do not have a common categorical axis on which they can be measured
+back to back. The topics gave us a general but shallow overview, while the
+document analysis provide deep insights on particular work but limited
+generalizations. We, for example, ignored rich social science literature (e.g.
+Anthropology) that is published in books and book chapters, hence not archived
+in scientific search engines. A more inductive approach to the qualitative
+coding could have gave us different insights on the literature outside the
+pre-defined categories we initially shared with the Resilience Alliance.
+
+Future studies could benefit from including gray literature such as theses,
+books, non-governmental organizations reports, government agencies reports, or
+non-english literature. Previous studies have shown that the use of full text
+instead of abstracts can also offer additional insights on the automated
+analyses, for example in attributing impacts of ecosystem services from regime
+shifts in social-ecological systems [@Rocha:2015tk]. Replication attempts do
+however face the challenge of accessing full text when many of the papers are
+behind pay-walls.
 
@@
 
-We reviewed the academic literature of the last two decades to trace how Panarchy related concepts have evolved, in what type of problems they have been applied and found useful, and what remains as the key frontiers of research.
-Despite a growing body of literature, no topic seems to dominate the academic production of scholars using these concepts.
-The feature most used is the adaptive cycle, and the problems where it is found most useful is in studies with a historical perspective, either in short time horizons such as natural resource management, urban development, or conservation; all the way to archeology and anthropology studies on the time horizon of centuries to millennia.
-Hierarchies in scale or maladaptive traps have received comparatively less attention.
-The frontiers of research point out to the problem of measurement and prediction.
-For example, how to observe and approximate resilience, connectedness and potential in the present; or how likely a current social-ecological system is to undergo a transition in the phases of the adaptive cycle, or a “memory” or “revolt” type of cross-scale interaction.
-While our survey of the literature offers a few options to empirically ground panarchy concepts, we believe the problem of measurement and prediction are fertile ground for future research efforts.
+We reviewed the academic literature of the last two decades to trace how
+Panarchy related concepts have evolved, in what type of problems they have been
+applied and found useful, and what remains as the key frontiers of research.
+Despite a growing body of literature, no topic seems to dominate the academic
+production of scholars using these concepts. The feature most used is the
+adaptive cycle, and the problems where it is found most useful is in studies
+with a historical perspective, either in short time horizons such as natural
+resource management, urban development, or conservation; all the way to
+archeology and anthropology studies on the time horizon of centuries to
+millennia. Hierarchies in scale or maladaptive traps have received comparatively
+less attention. The frontiers of research point out to the problem of
+measurement and prediction. For example, how to observe and approximate
+resilience, connectedness and potential in the present; or how likely a current
+social-ecological system is to undergo a transition in the phases of the
+adaptive cycle, or a "memory" or "revolt" type of cross-scale interaction. While
+our survey of the literature offers a few options to empirically ground panarchy
+concepts, we believe the problem of measurement and prediction are fertile
+ground for future research efforts.
 
@@
 
-JCR acknowledges financial support from Formas research grant (942-2015-731), the Stockholm Resilience Centre, and Stockholm University.
+JCR acknowledges financial support from Formas research grant (942-2015-731),
+the Stockholm Resilience Centre, and Stockholm University.
+
+# Data availability
+
+All data and code to replicate the analysis are available at:
+<https://github.com/juanrocha/panarchy>. The code book and database are
+available in thea public online repository
+([https://doi.org/10.6084/m9.figshare.13490919](doi.org/10.6084/m9.figshare.13490919)).
 
@@
 
-<div id="refs"></div>
+::: {#refs}
+:::
 
@@
 
+```{=tex}
 \renewcommand\thefigure{S\arabic{figure}}
@@
 \setcounter{figure}{0}
-
+```
+```{=tex}
 \begin{figure*}[h]
@@
 \includegraphics[width = 6in, height = 5in]{SM_topics25_timeline}
-\caption{\textbf{Proportion of papers per year per topic.} }
+\caption{\textbf{Proportion of papers per year per topic.} Each line represent the average proportion of content for a given year over all papers published in the same year.}
 \label{fig:figS1}
 \end{figure*}
+```
+
+```{=tex}
+\begin{figure*}[h]
+\centering
+\includegraphics[width = 6in, height = 5in]{sm2}
+\caption{\textbf{Proportion of topics per paper.} Each dot is a paper with its respective proportion of topic per year.
+A loess fit (blue lines) show the lack of trends over time for most topics.
+The color of topics correspond to Fig \ref{fig:topics25}}
+\label{fig:figS2}
+\end{figure*}
+```

### 2021-05-19 03-figures.R

@@ -195,18 +195,32 @@ g_stack <-  df_documents %>%
 
 # + 
 #     theme(axis.text.x = element_text(size = 4))
+sm2 <- df_documents %>% 
+    left_join(
+        dat %>% 
+            select(abstract, title, year) %>%
+            filter(!is.na(abstract)) %>%
+            unique(),
+        by = c("document" = "title") ) %>%
+    mutate(year = as.numeric(year)) %>% 
+    filter(year > 2000, year <2021) %>%
+    ggplot(aes(year, gamma) )+
+    geom_jitter(
+        aes(color = as.factor(topic), fill = as.factor(topic), alpha = gamma),
+        size = 1, show.legend = FALSE) +
+    geom_smooth(method = "loess") +
+    scale_alpha_continuous(range = c(0.1,1)) +
+    facet_wrap(.~topic, ncol = 5) +
+    labs(y = "Relative proportion of topics", x = "Year") +
+    #theme_classic(base_size = 6) 
+    theme_light(base_size = 6) +
+    theme(axis.text.x = element_text(size = 4))
 
-    # ggplot(aes(year, gamma) )+
-    # geom_jitter(
-    #     aes(color = as.factor(topic), fill = as.factor(topic), alpha = gamma),
-    #     size = 1, show.legend = FALSE) +
-    # geom_smooth(method = "loess") +
-    # scale_alpha_continuous(range = c(0.1,1)) +
-    # facet_wrap(.~topic, ncol = 5) +
-    # #labs(tag = "B") +
-    # theme_light(base_size = 6) + 
-    # theme(axis.text.x = element_text(size = 4))
-
+ggsave(
+    plot = sm2,
+    filename = "sm2.png", path = "paper/figures",
+    device = "png", width = unit(5,"in"), height = unit(5,"in"),
+    dpi = 500)
 
 (g_tl | g_prop | g_stack) / g_25 + plot_layout(heights = unit(c(1.5,4), "in"))
 
@@ -238,6 +252,7 @@ qual_summary <- qual %>%
     add_column(var_type = c(
         rep("panarchy feature", 2), rep("paper type", 5), 
         rep("panarchy feature", 4),rep("paper type", 2))) %>% 
+    mutate(feature = str_replace_all(feature, "idenfity phases", "identify phases")) %>% 
     mutate(feature = as_factor(feature)) %>%
[…]

### 2021-06-14 paper/paper_panarchy.Rmd
added suggestions by co-authors and corrected refs format

@@
 In 2002 Holling and colleagues proposed the concept of Panarchy, which presented social-ecological systems as an interacting set of adaptive cycles, each of which is produced by the dynamic tensions between novelty and efficiency at multiple scales.
-Initially introduced as a conceptual framework and set of metaphors, panarchy has gained the attention of scholars across many disciplines and its ideas continue to inspire further conceptual developments.
+Initially introduced as a conceptual framework and set of metaphors, panarchy has gained the attention of scholars across many disciplines, and its ideas continue to inspire further conceptual developments.
 Almost twenty years after this concept was introduced we review how it has been used, tested, extended and revised.
@@
 bibliography: panarchy.bib
-# csl: nature.csl
+csl: "/Users/juanrocha/Documents/styles/ecology.csl"
 output:
+  rtf_document:
+    toc: no
   pdf_document:
@@
     latex_engine: xelatex
-  rtf_document:
-    toc: no
 citecolor: blue
@@
 - \usepackage[margin=2.5cm]{geometry}
-
-## For Arxiv submission, use pdflatex as engine, comment out font, increase font size to 11, delete the package fontspec and the graphics path.
-Compile in latex so the bibliography is part of the tex file produced.
 ---
@@
 Resilience Alliance understood change in social-ecological systems. The concept
-of Panarchy was a key focus on this influential book. In the late 1990s and
+of Panarchy was a key focus of this influential book. In the late 1990s and
 early 2000s, the Resilience Alliance was a productive, innovative, and highly
-collaborative group of interdisciplinary scientists that focused on bridging
-social and natural sciences as well as theory and practice. They were addressing
-social-ecological problems by combining insights from the social and natural
-sciences, as well as the arts and humanities [@Parker:2012eo].
+collaborative group of interdisciplinary scientists that focused on addressing social-ecological problems.
+They did this by combining insights from the social and natural sciences, arts, and humanities, as well as by bridging theory and practice
+ [@Parker:2012eo].
 
@@
 is little space for experimentation and innovation. Examples include systems
-where ecological resilience has been expensively replaced by artificial
+where ecological resilience has been extensively replaced by artificial
 processes to maintain the system such as levees, flood barriers, or chemical
@@
 Panarchy offers a rich conceptual framework for understanding environmental
-problems. While inspired by a few mathematical constructs (e.g. cycles, traps,
+problems. While inspired by several mathematical constructs (e.g. cycles, traps,
 scaling laws), it is general enough to invite scholars from multiple
@@
 Here we review the academic literature of the last two decades to trace how
-these concepts have evolved, in what type of problems they have been applied and
-found useful, and finally what remains as the key frontiers of research.
+the ideas proposed in Panarchy have evolved, to what type of problems they have been applied and
+found useful, and finally, what key research frontiers remain.
 
@@
 with documents in rows and words in columns. Here our unit of analysis for the
-document are the abstracts retrieved, and the matrix contains the count of words
-per document. We removed stop words (e.g. "the", "a") and digits from the
+document are the abstracts retrieved, and the matrix contains the count of the number of occurrences of each word per abstract. We removed stop words (e.g. "the", "a") and digits from the
 matrix, as well as a list of words that were over represented in our data and
@@
 strong trends, but some topics have gained a small amount of attention. For
-example, topic 1 on sustainability, or topic 10 on ecosystem services appeared
+example, topic 1 on sustainability, and topic 10 on ecosystem services appeared
 consistently across time. In contrast, research on business innovation (topic
-20), urban infrastructure (topic 3), or archaeology (topic 5) have gained
+20), urban infrastructure (topic 3), and archaeology (topic 5) have gained
 attention in recent years. Topic 8 was an outlier, with a selection of papers
@@
 accounting for slightly different terminology such as "lock-ins". The bulk of
-the papers analyzed were conceptual papers, many without a method or discernible
+the papers analyzed were conceptual papers, many without defined methods or discernible
 research question. Roughly half of the papers were conceptual or only used
@@
 Spatially, 4 were at a city scale, 17 regional, and 1 national. Empirical papers
-were dominated by qualitative methods (77%, N = 22) and were generally
+predominantly used qualitative methods (77%, N = 22) and were generally
 retrospective historical reviews.
@@
 from delimitation of periods in archaeology and anthropology [@Redman:2003bd],
-to financial crises in Europe [@Castell:2020dg], or traffic jams in China
-[@Zeng:2020hu]. Recent empirical tests of the adaptive cycle innovate on the
-types of data and methods that one can use. Information transfer methods based
+to financial crises in Europe [@Castell:2020dg], to traffic jams in China
+[@Zeng:2020hu]. Some recent empirical tests of the adaptive cycle used innovative
+types of data and methods. For example, information transfer methods based
 on entropy have been suggested to approximate relevant components of a system
-and the empirical proxies of the axis where the adaptive cycle unfolds:
+and the empirical proxies of the axes within which the adaptive cycle unfolds:
 potential, connectedness, and resilience [@Castell:2020dg]. Percolation methods
-combined with big data have been shown to be useful to test hypotheses of regime
+combined with big data have been proven useful for testing hypotheses of regime
 shifts in urban systems (originally proposed in Panarchy [@Gunderson:2002vk]),
-and derive the temporal and spatial scales at which the adaptive cycle emerges
+and deriving the temporal and spatial scales at which the adaptive cycle emerges
 [@Zeng:2020hu].
@@
 
-The high level of conceptualization but lack of empirically grounded theory
-testing came across as one of the major limitations. Here we do not refer only
-to positivism as a plausible way of theory development. Other equally valid
+The high level of conceptualization but lack of empirically grounded testing of panarchy theory
+came across as one of the major limitations. Here we do not refer to positivism as the only plausible way of theory development, but acknowledge that other equally valid
 approaches exist. From the philosophical point of view of what makes a
@@
 explanations, and identify the circumstances under which a theory may fail
-[@Hansson:2017tr]. Many of the papers we read confirm but not challenged the
-expectations that Panarchy put forward. A few of the most recent work actually
+[@Hansson:2017tr]. Many of the papers we read confirmed but not challenge the
+expectations that Panarchy put forward, though a few of the most recent works did
 falsify some of the ideas proposed --e.g. the back loop is faster that the front
-one. We believe more work is needed to better understand the circumstances where
-Panarchy works, and when it falls short. In other words, refine the scientific
-framework towards a scientific theory.
+one [@Castell:2020dg]. We believe more work is needed to better understand the circumstances where
+Panarchy works, and when it falls short. In other words, to refine the scientific
+framework into a scientific theory.
 
-Given the strong dominance of conceptual papers, it is not surprising that one
+Given the strong predominance of conceptual papers, it is not surprising that one
 of the key frontiers of research is to empirically ground such
@@
 metaphors. Metaphors have played an important role as boundary objects enabling
-interdisciplinary dialogues. We observe it on the wide applications that
-panarchy has inspired, from anthropology to engineering. In the empirical realm,
-the adaptive cycle is the panarchy dimension that researchers engage the most
-with, but it is dominated by qualitative and retrospective studies. That means,
-we still lack the theory and methods to gather observations and be able to
-decide in which phase of the adaptive cycle a system currently is, or what is
-the probability of it transitioning to a new phase?
+interdisciplinary dialogues. We observed this on the wide variety of applications that
+panarchy has inspired, in fields from anthropology to engineering. In the empirical realm,
+the adaptive cycle is the dimension of panarchy with which researchers engage the most, but it is dominated by qualitative and retrospective studies. That means,
+we still lack the theory and methods to gather observations and identify a system's current adaptive cycle phase, or the probability of it transitioning to a new phase.
 
-The criticism of over-conceptualization is not unique to the literature engaging
+The criticism of over-conceptualization is not unique to the literature that engage with
 Panarchy ideas. Here by over-conceptualization we refer to the problem of
 scientific jargon, where different communities use slightly different wordings
-to refer to the same phenomena; reducing the chances that lessons from a
+to refer to the same phenomenon; reducing the chances that lessons from a
 disciplines percolates to another. An example from Panarchy is the endorsement
@@
 schools of thought that the discipline has developed over the last decades
-[@Clark:2020jd]. A similar conclusion was reached, where too many conceptual
-frameworks have been developed, but fewer empirical attempts try to test the
-frameworks against data to falsify hypotheses. The review emphasizes the problem
+[@Clark:2020jd] and reached a similar conclusion: too many conceptual
+frameworks have been developed, but there have been too few empirical attempts to test the
+frameworks against data and falsify hypotheses. The review also emphasizes the problem
 of measurement and observation [@Clark:2020jd]. In the context of panarchy,
@@
 The panarchy dimensions that received less attention in our qualitative analysis
-are creative destruction, rigidity and poverty traps. This may be at least
+included creative destruction, rigidity and poverty traps. This may be at least
 partly due to the scope of our data: papers that have cited Panarchy or
@@
 This paper aimed to study the use of panarchy related concepts and their
-evolution since the publication of the book. Systematic literature reviews often
-suffer from the limitation of restricting sample size to a subset of what is
-readable on the time frame of a project -- between dozens and maybe hundreds of
-papers. Qualitative analysis offers rich insights into the papers but is limited
-by sample size. In our analysis, the qualitative insights could be biased
-towards a non-random selection of papers that aligned our research interests, or
-highly cited, review type of work. Topic modeling enabled us to complement the
-analysis to all papers reported in major scientific databases. It has the
-advantage of reproducibility and reduces sample bias, but offers little insights
-about the dimensions of panarchy used, methods, or criticisms. Here we show that
+evolution since the publication of the book. Systematic literature reviews are often restricted to a limited subset of the literature -- based on what is
+readable on the time frame of a project -- of dozens to maybe hundreds of
+papers. Qualitative analysis offers rich insights into the papers reviewed but is limited
+by sample size in its ability to draw general conclusions. In our analysis, the qualitative insights could be biased
+towards a non-random selection of papers that aligned our research interests or
+highly cited review papers. Topic modeling enabled us to complement the
+analysis and include all papers indexed by major scientific databases. It has the
+advantage of reproducibility and reduces sample bias, but offers limited insights
+into the dimensions of panarchy used, methods, or criticisms. Here we show that
 both methods combined are a powerful combination.
 
-Our choice of methods also has limitations. We could not compare both methods
-since they do not have a common categorical axis on which they can be measured
+Our choice of methods does however have some limitations. We could not directly compare the two methods
+since they do not share a common categorical axis on which they can be measured
 back to back. The topics gave us a general but shallow overview, while the
 document analysis provide deep insights on particular work but limited
-generalizations. We, for example, ignored rich social science literature (e.g.
-Anthropology) that is published in books and book chapters, hence not archived
-in scientific search engines. A more inductive approach to the qualitative
+generalizations. We also, for example, ignored rich areas of social science research (e.g.
+Anthropology) that is published in books and book chapters, that are not archived
+by scientific search engines. Additionally, a more inductive approach to the qualitative
 coding could have gave us different insights on the literature outside the
@@
 books, non-governmental organizations reports, government agencies reports, or
-non-english literature. Previous studies have shown that the use of full text
+non-english literature. This would let us see the usage of Panarchy concepts in more applied settings such as management and decision making. Previous studies have also shown that the use of full text
 instead of abstracts can also offer additional insights on the automated
@@
 production of scholars using these concepts. The feature most used is the
-adaptive cycle, and the problems where it is found most useful is in studies
-with a historical perspective, either in short time horizons such as natural
-resource management, urban development, or conservation; all the way to
-archeology and anthropology studies on the time horizon of centuries to
+adaptive cycle, and it appears to have been found most useful is in studies
+with a historical perspective, both in short time studies in natural
+resource management, urban development, and conservation; as well as 
+archeological and anthropological studies that span centuries to
 millennia. Hierarchies in scale or maladaptive traps have received comparatively
-less attention. The frontiers of research point out to the problem of
-measurement and prediction. For example, how to observe and approximate
-resilience, connectedness and potential in the present; or how likely a current
-social-ecological system is to undergo a transition in the phases of the
-adaptive cycle, or a "memory" or "revolt" type of cross-scale interaction. While
-our survey of the literature offers a few options to empirically ground panarchy
-concepts, we believe the problem of measurement and prediction are fertile
+less attention. The frontiers of research point to the problems of
+measurement and prediction. For example, it is difficult to observe and approximate
+resilience, connectedness and potential in the present; or to determine the likelihood that a 
+social-ecological system will transition to a different phases of the
+adaptive cycle, or experience a "memory" or "revolt" type of cross-scale interaction. While
+our survey of the literature identified a few promising pathways towards empirically grounding panarchy
+concepts, we believe the problem of measurement and prediction remain fertile
 ground for future research efforts.

### 2021-06-15 paper/paper_panarchy.Rmd
Review 1 round

@@
 output:
-  rtf_document:
-    toc: no
   pdf_document:
@@
     latex_engine: xelatex
+  rtf_document:
+    toc: no
 citecolor: blue

### 2021-09-15 paper/paper_panarchy.Rmd
Second round of review, updated analysis with lemmatization.

@@
 Challenges remain in empirically grounding the metaphor, but recent theoretical and empirical work offer some avenues for future research.
-bibliography: panarchy.bib
+bibliography: [panarchy.bib, packages.bib]
 csl: "/Users/juanrocha/Documents/styles/ecology.csl"
@@
 
+```{r message = FALSE, warnings=FALSE, include = FALSE}
+library(here)
+library(tidytext)
+library(tidyverse)
+library(tm)
+library (topicmodels)
+library(lda)
+
+load(here("data/dtm.Rdata"))
+load(here("data/qual_coding.Rdata"))
+```
+
 # Introduction
@@
 early 2000s, the Resilience Alliance was a productive, innovative, and highly
-collaborative group of interdisciplinary scientists that focused on addressing social-ecological problems.
-They did this by combining insights from the social and natural sciences, arts, and humanities, as well as by bridging theory and practice
- [@Parker:2012eo].
+collaborative group of interdisciplinary scientists that focused on addressing
+social-ecological problems. They did this by combining insights from the social
+and natural sciences, arts, and humanities, as well as by bridging theory and
+practice [@Parker:2012eo].
 
@@
 attention on processes of destruction and reorganization, which had been
-neglected in comparison to those of growth and conservation. Considering these
-processes provides a more complex way to understand system dynamics.
+neglected in comparison to those of growth and conservation.
 
@@
 key components [@Holling:2002wu]. Resilience researchers proposed that tensions
-between demands for organization and efficiency versus demands for novelty and
-diversity drive adaptive cycle type dynamics in many different types of complex
+between demands for organization and efficiency, versus demands for novelty and
+diversity, drive adaptive cycle dynamics in many different types of complex
 systems [@Holling:2002wu].
@@
 adaptive change. The adaptive cycle is hypothesized to exist within a
-three-dimensional space defined by three properties: potential, connectivity,
+three-dimensional space defined by three properties: potential, connectedness,
 and resilience [@Gunderson:2002vk]. Potential refers to the capital available to
 the system, for example nutrients and carbon captured by a forest, or human
-capital - skills and knowledge accumulated to run the economy. Connectivity is a
-proxy of the structure of the system; it is the network of interactions and
+capital - skills and knowledge accumulated to run the economy. Connectedness is
+a proxy of the structure of the system; it is the network of interactions and
 strength between its elements. Resilience is the capacity of any system to
@@
 Poverty traps are described in panarchy as maladaptive states where potential,
-connectivity and resilience are low. Poverty traps are a series of feedback
+connectedness and resilience are low. Poverty traps are a series of feedback
 mechanisms that reinforce impoverished states [@Bowles:2006hx; @Maru:2012dm;
@@
 ecologists have demonstrated the utility of the framework to their area of
-research. However, the extent to which these concepts have gained empirical
-support beyond the metaphor, or how they have evolved over time, remain open
-questions. For example, the adaptive cycle is a useful metaphor to look back at
-history and organize events and periods around phases, but it remains
-challenging to identify the phase of a system in real time, or to use panarchy
-to project future trajectories.
+research.
 
-Here we review the academic literature of the last two decades to trace how
-the ideas proposed in Panarchy have evolved, to what type of problems they have been applied and
-found useful, and finally, what key research frontiers remain.
+However, the extent to which these concepts have gained empirical support beyond
+the metaphor, or how they have evolved over time, remain open questions. For
+example, the adaptive cycle is a useful metaphor to look back at history and
+organize events and periods around phases, but it remains challenging to
+identify the phase of a system in real time, or to use panarchy to project
+future trajectories. Scientific theories are bodies of knowledge that answer the
+question why? The answer implies causal reasoning. Regardless of discipline, a
+theory should therefore provide a series of testable expectations that help
+distinguishing the situations when it is a good explanation of reality, versus
+situations when it falls short. The practice of science implies avoiding
+confirmation bias and testing when explanations are valid representations of
+reality. For panarchy to become a useful theory in sustainability science, it
+needs to stand the rigor of testing, and being able to explain phenomena in
+social-ecological systems across different contexts. We identify progress and
+knowledge gaps in the scientific practice inspired by panarchy.
+
+Here we review the academic literature of the last two decades to trace how the
+ideas proposed in Panarchy have evolved, to what type of problems they have been
+applied and found useful, and finally, what key research frontiers remain.
 
@@
 topic modeling [@Blei:2012dk; @Griffiths:2004ey] with human-coded document
-analysis [@Bryman:2008va].
+analysis [@Bryman:2008va]. We first describe the data curation and how it was
+used in the topic models. Then we introduce the qualitative approach.
 
@@
 that they are accepted manuscripts with digital object identifier (doi)
-scheduled to be published later in 2020 or 2021.
+scheduled to be published later in 2020 or 2021. Our final sample is
+`r dim(dtm)[1]` documents.
 
 To prepare the data for topic modeling, we constructed a document term matrix
-with documents in rows and words in columns. Here our unit of analysis for the
-document are the abstracts retrieved, and the matrix contains the count of the number of occurrences of each word per abstract. We removed stop words (e.g. "the", "a") and digits from the
-matrix, as well as a list of words that were over represented in our data and
-are common in the scientific literature but are unrelated to the papers' topics
-(e.g. "paper", "study", "aim").
+with documents in rows (N=`r dim(dtm)[1]`) and words in columns
+(N=`r dim(dtm)[2]`). Here our unit of analysis for the document are the
+abstracts retrieved, and the matrix contains the count of the number of
+occurrences of each word per abstract. We removed stop words (e.g. "the", "a")
+and digits from the matrix, as well as a list of words that were over
+represented in our data and are common in the scientific literature but are
+unrelated to the papers' topics (e.g. "paper", "study", "aim"). The lexicon of
+stop words has 1149 entries, curated by various research groups, and accessible
+in the `R` language under the packages `tm` and `tidytext` [@R-topicmodels;
+@R-tm; @R-NLP; @tm2008; @R-tidytext], for example by `tidytext::stop_words`. The
+data was also lemmatized using the hash lemmas dataset available at
+[\<http://www.lexiconista.com\>](http://www.lexiconista.com/). Lemmatization is
+the process of simplifying words to their root lemmas (e.g. *teaching* and
+*taught* are lemmatized to *teach*).
 
@@
 performance were the same as random, perplexity would approximate the vocabulary
-size (N = 16 728 words). These performance metrics enable us to choose which
-algorithm best fits our data, what is the optimal number of topics to fit, and
-how to avoid overfitting.
+size (N = `r dim(dtm)[2]` words). These performance metrics enable us to choose
+which algorithm best fits our data, what is the optimal number of topics to fit,
+and how to avoid overfitting.
 
@@
 analysis [@Bryman:2008va] by coding an additional set of categorical variables
-for a sample of papers (N = 42). Based on the reading of Panarchy, we first
-developed a series of categorical variables that captured the main components of
-the scientific framework. The variables were presented and received feedback at
-the science meeting of the Resilience Alliance in 2019. We used the variables to
-annotate qualitative aspects such as use of the adaptive cycle, identification
-of its phases, whether the paper is conceptual, modeling or an empirical study.
-When empirical, we recorded the temporal and spatial scales of the case study.
+for a sample of papers (N = `r nrow(qual)`). The papers read are part of the
+corpus used for topic modelling. We read papers avoiding biasing our selection
+to high cited papers, specific journals, older papers or particular disciplines
+or methods.
+
+Based on the reading of Panarchy, we first developed a series of categorical
+variables that captured the main components of the scientific framework. The
+variables were presented and received feedback at the science meeting of the
+Resilience Alliance in 2019. We used the variables to annotate qualitative
+aspects such as use of the adaptive cycle, identification of its phases, whether
+the paper is conceptual, modeling or an empirical study. When empirical, we
+recorded the temporal and spatial scales of the case study. We also
+distinguished empirical cases from empirical constructs. An empirical case is
+when a study is trying to test an idea or proposition put forward by Panarchy
+(e.g. a cross-scale interaction, or that the back loop is faster than the forth
+loop). An empirical construct is when Panarchy is used as inspiration to
+interpret a case study, typically with a historical component to it (e.g.
+identify the phases of adaptive cycle); but it does not find support for or
+against a Panarchy preposition. That is, there was no mechanism in place to
+avoid confirmation bias.
+
 We also identified what aspects of panarchy were most used in the papers. For
@@
 rigidity traps. We used text annotations to capture potential criticisms as well
-as the methods used.
+as the methods used. We did a preliminary scoping and coding of papers, and
+through a iterative process we kept track of potential fields under reviewed by
+our corpus, or simply absent of it. We coded papers until we reached conceptual
+saturation, that is, additional papers were not proving additional categories of
+the already coded. The code book is available at
+[doi.org/10.6084/m9.figshare.13490919](http://doi.org/10.6084/m9.figshare.13490919).
 
@@
 Increasing the number of topics (from 5-100) shows that $\alpha$ decreases, suggesting that despite the larger number of topics, a few of them suffice to describe most papers (B).
-Log-likelihood is maximized for 50 topics followed closely with 25.
-While 50 topics is marginally better, for visualization purposes we restrict our analysis to 25.
+Log-likelihood is maximized for 50 topics followed closely with 25 and 100.
+While 50 topics is marginally better.
+We choose 25 because the $\alpha$ parameter suggest preferring lower numbers, or a more parsimonious solution to avoid over fitting.
 Note that perplexity cannot be calculated for Gibbs sampling, hence the missing value in A and absence in B.}
@@
 The proportion of papers per year (B) and the relative proportion of topic content per year (C) do not show strong trends for the time window with most papers (gray area in A).
+In (B), a reference dotted horizontal line marks the level where all topics are equally dominant.
 Each topic is summarized in (D) by the top 10-words that best describe them according to the posterior probability of our model fit.
-See FigS1 for a panel of time series per topics}
+Fig \ref{fig:figS1} expands (B) presenting time series for each topic.
+All time series exclude 2021 to account for the artificial decline of papers at the end of the time series as shown in A.}
 \label{fig:topics25}
@@
 represent an underlying content of the document (Fig \ref{fig:topics25}). For
-example, the words "resilience", "vulnerability", "risk" and "disaster" have a
-high probability to capture the content of topic 25. Papers early in our time
-series (2001-3) have high content largely dominated by topics 25 on resilience,
-24 on the adaptive cycle, and 18 on adaptive governance and management. Towards
-the end of the time series, topic 3 on urban systems and 9 on local communities
-and knowledge have spikes up to 9% of the content of each year (2019-2020). For
-comparison, if all topics were equally represented in the content, they would
-have 4% in the corpus (grey line in Fig \ref{fig:topics25}B).
+example, the words "resilience", "adaptive", or "capacity" have a high
+probability to capture the content of topic 14. Papers early in our time series
+(2001-3) have high content largely dominated by topics 14 on resilience, 9 on
+resource management, and 23 on cross-scale interactions. Towards the end of the
+time series, topic 7 on urban systems and 25 on local communities and knowledge
+have spikes up to 9% of the content of each year (2019-2020). For comparison, if
+all topics were equally represented in the content, they would have 4% in the
+corpus (grey line in Fig \ref{fig:topics25}B). The topics presented lack names
+for reproducibility purposes. While here we present e.g. topic 7 as the urban
+topic, if a researcher replicates the analysis that person will find the same
+list of words and probability distributions listed under a topic with a
+different number. That is because the labeling of topics is random and we do not
+want to bias the reader to over interpret the word patterns found.
 
+```{=tex}
+\begin{figure*}[h]
+\centering
+\includegraphics[width = 6in, height = 5in]{SM_topics25_timeline}
+\caption{\textbf{Proportion of papers per year per topic.} Each line represent the average proportion of content for a given year over all papers published in the same year.
+Fig \ref{fig:figS2} complements with the proportion of topics for each paper in our sample.}
+\label{fig:figS1}
+\end{figure*}
+```
 Despite fluctuations, most topics showed a relatively constant level of interest
-over time (Fig \ref{fig:topics25}, Fig \ref{fig:figS1}). We did not observe
-strong trends, but some topics have gained a small amount of attention. For
-example, topic 1 on sustainability, and topic 10 on ecosystem services appeared
-consistently across time. In contrast, research on business innovation (topic
-20), urban infrastructure (topic 3), and archaeology (topic 5) have gained
-attention in recent years. Topic 8 was an outlier, with a selection of papers
-ranking high on content related to network infrastructure and performance,
-possibly from engineering disciplines. It was the only topic with a set of
-papers that was clearly distinct from the rest of the collection, and showed a
-decreasing trend over time.
+over time (Fig \ref{fig:topics25}, \ref{fig:figS1}). We did not observe strong
+trends, but some topics have gained a small amount of attention. For example,
+topic 11 on sustainability, and topic 15 on ecosystem services appeared
+consistently across time. In contrast, research on innovation (topic 18), urban
+infrastructure (topic 7), and local communities (topic 13) have gained attention
+in recent years. Topic 6 was an outlier, with a selection of papers ranking high
+on content related to network infrastructure and performance, possibly from
+engineering disciplines. It was the only topic with a set of papers that was
+clearly distinct from the rest of the collection, and showed a decreasing trend
+over time.
 
 The human-coded document analysis revealed that the most common feature of
-panarchy in the literature was the adaptive cycle (81%, N = 42) followed by
-cross-scale interactions (Fig \ref{fig:qual}). Poverty and rigidity traps were
-less studied features in our sample (14% and 19% respectively), even when
-accounting for slightly different terminology such as "lock-ins". The bulk of
-the papers analyzed were conceptual papers, many without defined methods or discernible
-research question. Roughly half of the papers were conceptual or only used
-panarchy as a metaphor. Over 40% of papers in our sample went a step further and
-used panarchy as an empirical construct, for example by attempting to identify
-the phases of the adaptive cycle (61%). Out of the 22 empirical cases, 6 were at
-the time scale of centuries, 7 on decades, 6 on years, and 1 in weeks.
-Spatially, 4 were at a city scale, 17 regional, and 1 national. Empirical papers
-predominantly used qualitative methods (77%, N = 22) and were generally
-retrospective historical reviews.
+panarchy in the literature was the adaptive cycle (81%, N = `r nrow(qual)`)
+followed by cross-scale interactions (Fig \ref{fig:qual}). Poverty and rigidity
+traps were less studied features in our sample (14% and 19% respectively), even
+when accounting for slightly different terminology such as "lock-ins". 52.4% of
+the papers analyzed were conceptual papers, many without defined methods or
+discernible research question. Roughly half (47%) of the papers were conceptual
+or only used panarchy as a metaphor. Over 40% of papers in our sample went a
+step further and used panarchy as an empirical construct, for example by
+attempting to identify the phases of the adaptive cycle (61%). Out of the 22
+empirical cases, 6 were at the time scale of centuries, 7 on decades, 6 on
+years, and 1 in weeks. Spatially, 4 were at a city scale, 17 regional, and 1
+national. Empirical papers predominantly used qualitative methods (77%, N = 22)
+and were generally retrospective historical reviews.
 
-Several of the papers reviewed attempted to identify the adaptive cycle either
-through qualitative or quantitative methods. Identification of adaptive cycles
-has found applications in a wide range of disciplines and research problems,
-from delimitation of periods in archaeology and anthropology [@Redman:2003bd],
-to financial crises in Europe [@Castell:2020dg], to traffic jams in China
-[@Zeng:2020hu]. Some recent empirical tests of the adaptive cycle used innovative
-types of data and methods. For example, information transfer methods based
-on entropy have been suggested to approximate relevant components of a system
-and the empirical proxies of the axes within which the adaptive cycle unfolds:
-potential, connectedness, and resilience [@Castell:2020dg]. Percolation methods
-combined with big data have been proven useful for testing hypotheses of regime
-shifts in urban systems (originally proposed in Panarchy [@Gunderson:2002vk]),
-and deriving the temporal and spatial scales at which the adaptive cycle emerges
-[@Zeng:2020hu].
+Several of the papers reviewed (81%) attempted to identify the adaptive cycle
+either through qualitative or quantitative methods. Identification of adaptive
+cycles has found applications in a wide range of disciplines and research
+problems, from delimitation of periods in archaeology and anthropology
+[@Redman:2003bd], to financial crises in Europe [@Castell:2020dg], to traffic
+jams in China [@Zeng:2020hu]. Some recent empirical tests of the adaptive cycle
+used innovative types of data and methods. For example, information transfer
+methods based on entropy have been suggested to approximate relevant components
+of a system and the empirical proxies of the axes within which the adaptive
+cycle unfolds: potential, connectedness, and resilience [@Castell:2020dg].
+Percolation methods combined with big data have been proven useful for testing
+hypotheses of regime shifts in urban systems (originally proposed in Panarchy
+[@Gunderson:2002vk]), and deriving the temporal and spatial scales at which the
+adaptive cycle emerges [@Zeng:2020hu].
 
@@
 
-The high level of conceptualization but lack of empirically grounded testing of panarchy theory
-came across as one of the major limitations. Here we do not refer to positivism as the only plausible way of theory development, but acknowledge that other equally valid
-approaches exist. From the philosophical point of view of what makes a
-theory a scientific theory, the plurality of approaches should have mechanisms
-in place to avoid confirmation bias, allow the emergence of alternative
-explanations, and identify the circumstances under which a theory may fail
-[@Hansson:2017tr]. Many of the papers we read confirmed but not challenge the
-expectations that Panarchy put forward, though a few of the most recent works did
-falsify some of the ideas proposed --e.g. the back loop is faster that the front
-one [@Castell:2020dg]. We believe more work is needed to better understand the circumstances where
-Panarchy works, and when it falls short. In other words, to refine the scientific
-framework into a scientific theory.
-
-Given the strong predominance of conceptual papers, it is not surprising that one
-of the key frontiers of research is to empirically ground such
+Given the strong predominance of conceptual papers, it is not surprising that
+one of the key frontiers of research is to empirically ground such
 conceptualizations in a way that one can test theoretical expectations beyond
 metaphors. Metaphors have played an important role as boundary objects enabling
-interdisciplinary dialogues. We observed this on the wide variety of applications that
-panarchy has inspired, in fields from anthropology to engineering. In the empirical realm,
-the adaptive cycle is the dimension of panarchy with which researchers engage the most, but it is dominated by qualitative and retrospective studies. That means,
-we still lack the theory and methods to gather observations and identify a system's current adaptive cycle phase, or the probability of it transitioning to a new phase.
+interdisciplinary dialogues. We observed this on the wide variety of
+applications that panarchy has inspired, in fields from anthropology to
+engineering. In the empirical realm, the adaptive cycle is the dimension of
+panarchy with which researchers engage the most, but it is dominated by
+qualitative and retrospective studies. That means, we still lack the theory and
+methods to gather observations and identify a system's current adaptive cycle
+phase, or the probability of it transitioning to a new phase.
 
-The criticism of over-conceptualization is not unique to the literature that engage with
-Panarchy ideas. Here by over-conceptualization we refer to the problem of
-scientific jargon, where different communities use slightly different wordings
-to refer to the same phenomenon; reducing the chances that lessons from a
-disciplines percolates to another. An example from Panarchy is the endorsement
+The high level of conceptualization but lack of empirically grounded testing of
+panarchy theory came across as one of the major limitations, given our sample of
+papers reviewed (N=`r nrow(qual)`). What makes a theory a scientific theory is
+that the plurality of scientific approaches (e.g. positivism, constructivism)
+should have mechanisms in place to avoid confirmation bias, allowing the
+emergence of alternative explanations, and identify the circumstances under
+which a theory may fail [@Hansson:2017tr]. Many of the papers we read confirmed
+but did not challenge the expectations that Panarchy put forward, though a few
+of the most recent works did falsify some of the ideas proposed --e.g. the back
+loop is faster that the front one [@Castell:2020dg]. We believe more work is
+needed to better understand the circumstances where Panarchy works, and when it
+falls short. In other words, to refine the scientific framework into a
+scientific theory.
+
+The criticism of over-conceptualization is not unique to the literature that
+engages with Panarchy ideas. Here by over-conceptualization we refer to the
+problem of scientific jargon, where different communities use slightly different
+wordings to refer to the same phenomenon; reducing the chances that lessons from
+a discipline percolates to another. An example from Panarchy is the endorsement
 of ideas such as creative destruction (from economics) or the theory of
@@
 schools of thought that the discipline has developed over the last decades
-[@Clark:2020jd] and reached a similar conclusion: too many conceptual
-frameworks have been developed, but there have been too few empirical attempts to test the
-frameworks against data and falsify hypotheses. The review also emphasizes the problem
-of measurement and observation [@Clark:2020jd]. In the context of panarchy,
-recent work has developed an information theory based approach that enables the
-identification of adaptive cycles [@Castell:2020dg]. The authors identify phases
-of the adaptive cycle in the European financial crises and in grassland
-ecosystems, but fail to find support for the difference in speed between the
-forward and backward loops originally proposed in Panarchy.
+[@Clark:2020jd] and reached a similar conclusion: too many conceptual frameworks
+have been developed, but there have been too few empirical attempts to test the
+frameworks against data and falsify hypotheses. The review also emphasizes the
+problem of measurement and observation [@Clark:2020jd]. In the context of
+panarchy, recent work has developed an information theory based approach that
+enables the identification of adaptive cycles [@Castell:2020dg]. The authors
+identify phases of the adaptive cycle in the European financial crises and in
+grassland ecosystems, but fail to find support for the difference in speed
+between the forward and backward loops originally proposed in Panarchy.
 
@@
 partly due to the scope of our data: papers that have cited Panarchy or
-Holling's paper [@Gunderson:2002vk; @Holling:2001cu]. Concepts such as creative
-destruction and poverty traps precede Panarchy and therefore have been theorized
-and empirically grounded outside the panarchy stream of thinking. For example,
-the theory of poverty traps dates back to the 1950's in economics, and has
-received both theoretical development [@Bowles:2006hx] as well as empirical
-grounding [@Banerjee:2015ja; @Banerjee:2012tq] that has enabled researchers and
+Holling's paper [@Gunderson:2002vk; @Holling:2001cu], and a relatively small
+sample size (N=`r nrow(qual)`). Concepts such as creative destruction and
+poverty traps precede Panarchy and therefore have been theorized and empirically
+grounded outside the panarchy stream of thinking. For example, the theory of
+poverty traps dates back to the 1950's in economics, and has received both
+theoretical development [@Bowles:2006hx] as well as empirical grounding
+[@Banerjee:2015ja; @Banerjee:2012tq] that has enabled researchers and
 governments to distinguish what kind of interventions are likely to reduce
@@
 
+Sustainability science is defined by the problems it addresses rather than the
+disciplines it employs [@Anonymous:2016fv]. Our topic modeling results helped in
+identifying some of the problems that the community is addressing and how they
+have changed over time. Some of these problems include climate change (topic
+24), urban development and disaster response (7), conservation and landscape
+management trade-offs (5), natural resource management (9), food production (3),
+policy tools and plans (16), community based management (13) or knowledge and
+local social practices (25. We also observed topics closely related to methods
+or approaches to study social-ecological systems such as modeling (19,21),
+networks (6), or ecosystem services (15). Topic modeling can be used to further
+investigate particular academic communities, identify key landmark papers in
+certain topics, or identify the network of key collaborators and experts.
+
 This paper aimed to study the use of panarchy related concepts and their
-evolution since the publication of the book. Systematic literature reviews are often restricted to a limited subset of the literature -- based on what is
+evolution since the publication of the book. Systematic literature reviews are
+often restricted to a limited subset of the literature -- based on what is
 readable on the time frame of a project -- of dozens to maybe hundreds of
-papers. Qualitative analysis offers rich insights into the papers reviewed but is limited
-by sample size in its ability to draw general conclusions. In our analysis, the qualitative insights could be biased
-towards a non-random selection of papers that aligned our research interests or
-highly cited review papers. Topic modeling enabled us to complement the
-analysis and include all papers indexed by major scientific databases. It has the
-advantage of reproducibility and reduces sample bias, but offers limited insights
-into the dimensions of panarchy used, methods, or criticisms. Here we show that
-both methods combined are a powerful combination.
+papers. Qualitative analysis offers rich insights into the papers reviewed but
+is limited by sample size in its ability to draw general conclusions. In our
+analysis, the qualitative insights could be biased towards a non-random
+selection of papers that aligned our research interests or highly cited review
+papers. Topic modeling enabled us to complement the analysis and include all
+papers indexed by major scientific databases. It has the advantage of
+reproducibility and reduces sample bias, but offers limited insights into the
+dimensions of panarchy used, methods, or criticisms. Here we show that both
+methods combined are a powerful combination.
 
-Our choice of methods does however have some limitations. We could not directly compare the two methods
-since they do not share a common categorical axis on which they can be measured
-back to back. The topics gave us a general but shallow overview, while the
-document analysis provide deep insights on particular work but limited
-generalizations. We also, for example, ignored rich areas of social science research (e.g.
-Anthropology) that is published in books and book chapters, that are not archived
-by scientific search engines. Additionally, a more inductive approach to the qualitative
-coding could have gave us different insights on the literature outside the
-pre-defined categories we initially shared with the Resilience Alliance.
+Our choice of methods does however have some limitations. We could not directly
+compare the two methods since they do not share a common categorical axis on
+which they can be measured back to back. The topics gave us a general but
+shallow overview, while the document analysis provide deep insights on
+particular work but limited generalizations. We also, for example, ignored rich
+areas of social science research (e.g. Anthropology) that is published in books
+and book chapters, that are not archived by scientific search engines.
+Additionally, a more inductive approach to the qualitative coding could have
+given us different insights on the literature outside the pre-defined categories
+we initially shared with the Resilience Alliance.
 
@@
 books, non-governmental organizations reports, government agencies reports, or
-non-english literature. This would let us see the usage of Panarchy concepts in more applied settings such as management and decision making. Previous studies have also shown that the use of full text
-instead of abstracts can also offer additional insights on the automated
-analyses, for example in attributing impacts of ecosystem services from regime
-shifts in social-ecological systems [@Rocha:2015tk]. Replication attempts do
-however face the challenge of accessing full text when many of the papers are
-behind pay-walls.
+non-English literature. This would let us see the usage of Panarchy concepts in
+more applied settings such as management and decision making. Previous studies
+have also shown that the use of full text instead of abstracts can also offer
+additional insights on the automated analyses, for example in attributing
+impacts of ecosystem services from regime shifts in social-ecological systems
+[@Rocha:2015tk]. Replication attempts do however face the challenge of accessing
+full text when many of the papers are behind pay-walls.
 
@@
 production of scholars using these concepts. The feature most used is the
-adaptive cycle, and it appears to have been found most useful is in studies
-with a historical perspective, both in short time studies in natural
-resource management, urban development, and conservation; as well as 
-archeological and anthropological studies that span centuries to
-millennia. Hierarchies in scale or maladaptive traps have received comparatively
-less attention. The frontiers of research point to the problems of
-measurement and prediction. For example, it is difficult to observe and approximate
-resilience, connectedness and potential in the present; or to determine the likelihood that a 
-social-ecological system will transition to a different phases of the
-adaptive cycle, or experience a "memory" or "revolt" type of cross-scale interaction. While
-our survey of the literature identified a few promising pathways towards empirically grounding panarchy
-concepts, we believe the problem of measurement and prediction remain fertile
-ground for future research efforts.
+adaptive cycle. It have been found most useful in studies with a historical
+perspective, both in short time studies in natural resource management, urban
+development, and conservation; as well as archeological and anthropological
+studies that span centuries to millennia. Hierarchies in scale or maladaptive
+traps have received comparatively less attention. The frontiers of research
+point to the problems of measurement and prediction. For example, it is
+difficult to observe and approximate resilience, connectedness and potential in
+the present. How to determine the likelihood that a social-ecological system
+will transition to a different phase of the adaptive cycle, or experience a
+"memory" or "revolt" type of cross-scale interaction are still open questions.
+While our survey of the literature identified a few promising pathways towards
+empirically grounding panarchy concepts, we believe the problem of measurement
+and prediction remain fertile ground for future research efforts.
 
@@
 JCR acknowledges financial support from Formas research grant (942-2015-731),
-the Stockholm Resilience Centre, and Stockholm University.
+the Stockholm Resilience Centre, and Stockholm University. We acknowledge the
+feedback of the Resilience Alliance and encouragement in pursuing this project.
 
@@
 <https://github.com/juanrocha/panarchy>. The code book and database are
-available in thea public online repository
+available in the public online repository
 ([https://doi.org/10.6084/m9.figshare.13490919](doi.org/10.6084/m9.figshare.13490919)).
@@
 ```{=tex}
-\renewcommand\thefigure{S\arabic{figure}}
-\renewcommand\thetable{S\arabic{table}}
+\renewcommand\thefigure{A\arabic{figure}}
+\renewcommand\thetable{A\arabic{table}}
 \setcounter{table}{0}
@@
 ```
-```{=tex}
-\begin{figure*}[h]
-\centering
-\includegraphics[width = 6in, height = 5in]{SM_topics25_timeline}
-\caption{\textbf{Proportion of papers per year per topic.} Each line represent the average proportion of content for a given year over all papers published in the same year.}
-\label{fig:figS1}
-\end{figure*}
-```
-
 ```{=tex}

### 2021-09-15 01-read_data.R

@@ -6,17 +6,19 @@
 # libraries for modeling
 library(tidytext)
 library(tidyverse)
-
-
+library(tictoc)
+library(bib2df) # update to the github version, otherwise errors importing
 
 files <- list.files("data/") %>% str_subset(pattern = ".bib")
 files <- paste0("data/", files) 
 
-dat <- map(.x = files, .f = function(x){
+tic()
+dat <- map(.x = files[1:3], .f = function(x){ #exclude Gunderson file
     x %>% 
         bib2df::bib2df(separate_names = TRUE) %>%
         janitor::clean_names()
 })
+toc() #6s
 
 dat <- bind_rows(dat)
 
@@ -37,10 +39,11 @@ dat %>% pull(abstract) %>% unique() %>% length() # 2192
 
 ## extra words
 too_words <- tibble(
-    word = c("paper", "study", "aim", "aims", "objective", "purpose")
+    word = c("paper", "study", "aim", "aims", "objective", "purpose", "elsevier", "taylor", "francis")
 )
 
 ## creates the document term matrix
+tic()
 dtm <- dat %>% 
     select(abstract, title, year) %>%
     filter(!is.na(abstract)) %>%
@@ -50,10 +53,11 @@ dtm <- dat %>%
     anti_join(stop_words) %>% 
     anti_join(too_words) %>%
     filter(!str_detect(word, "[:digit:]")) %>%
+    mutate(word = textstem::lemmatize_words(word)) %>% 
     group_by(title) %>%
     count(word, sort = TRUE) %>%
     cast_dtm(document = title, term = word, value = n)
-
+toc() #3.5s
 dtm
 
 save(dtm, file = "data/dtm.Rdata")

### 2021-09-15 02-model.R

@@ -9,31 +9,31 @@ library(tidyverse)
 library(tm)
 library (topicmodels)
 library(lda)
-
+library(tictoc)
 load("data/dtm.Rdata")
 
 SEED <- 2010
 
 ## Choosing best algorithm
 k <- 10
-
+tic()
 tset.TM <- list (
     VEM0 = LDA(dtm, k=k, control = list ( seed = SEED)),
     VEM_fixed= LDA(dtm, k=k, control= list (estimate.alpha = F, seed = SEED)),
     Gibbs = LDA (dtm, k=k, method ="Gibbs", control = list (seed = SEED, burnin= 1000, thin = 100, iter= 1000)),
     CTM = CTM (dtm, k=k, control = list(seed = SEED, var= list (tol= 10^-4), em= list (tol = 10^-3))))
-
+toc() # 164s
 sapply (tset.TM[1:3], slot, "alpha")
 
 #Finding number of topics
 k <- c(5,10,25,50,100)
-
+tic()
 topicNumber.TM <- map(
     .x = k,
     .f = function(x) {
     LDA(dtm, k = x, control= list (seed = SEED), method = "Gibbs")
     })
-
+toc() # 314s
 
 save(tset.TM, topicNumber.TM, file = "data/models_gibbs.RData")
 

### 2021-09-15 03-figures.R

@@ -167,7 +167,7 @@ g_prop <- df_documents %>%
     geom_line(aes(color = as.factor(topic)), show.legend = FALSE, size = 0.25) +
     geom_hline(yintercept = 0.04, color = "gray50", size = 0.5, linetype = 2 ) +
     labs(y = "Proportion of papers", x = "Year", tag = "B") +
-    facet_wrap(.~topic, ncol = 5) +
+    #facet_wrap(.~topic, ncol = 5) +
     theme_light(base_size = 6) 
 
 ## alternative with stacked lines
@@ -215,7 +215,7 @@ sm2 <- df_documents %>%
     #theme_classic(base_size = 6) 
     theme_light(base_size = 6) +
     theme(axis.text.x = element_text(size = 4))
-
+sm2
 ggsave(
     plot = sm2,
     filename = "sm2.png", path = "paper/figures",
@@ -266,6 +266,8 @@ qual_summary %>%
     theme_light(base_size = 6)
 
 ggsave("paper/figures/fig4_qual_results.png", device = "png", width = unit(5,"in"), height = unit(2.5,"in"))
+
+save(qual, file = "data/qual_coding.RData")
 
 #### Fig 1 ####
 

### 2021-09-15 04-packages_bibliography.R

@@ -1,0 +1,12 @@
+## This are the key packages used in the analysis:
+
+library(tidytext)
+library(tidyverse)
+library(tm)
+library (topicmodels)
+library(lda)
+
+pkgs <- c("tidytext", "tidyverse", "tm", "topicmodels", "lda") 
+
+## automatically create a bib datasbase of R packages:
+knitr::write_bib( c(.packages(), pkgs), "packages.bib")

### 2021-10-07 paper/paper_panarchy.Rmd
edits second round of review

@@
 
-The Panarchy concept builds on Holling's adaptive cycle [@biosphere:hJ4c_Ua8] by
-extending the idea across spatial and temporal scales. It remains a boundary
+The Panarchy concept builds on Holling's adaptive cycle [@biosphere:hJ4c_Ua8]
+by extending the idea across spatial and temporal scales. It remains a boundary
 object that has inspired research topics, enabled collaborations and nurtured
@@
 similar form, or may transition to a new type of system with new boundaries and
-key components [@Holling:2002wu]. Resilience researchers proposed that tensions
-between demands for organization and efficiency, versus demands for novelty and
-diversity, drive adaptive cycle dynamics in many different types of complex
-systems [@Holling:2002wu].
+key components [@Holling:2002wu]. Resilience researchers proposed that
+tensions between demands for organization and efficiency, versus demands for
+novelty and diversity, drive adaptive cycle dynamics in many different types of
+complex systems [@Holling:2002wu].
 
@@
 three-dimensional space defined by three properties: potential, connectedness,
-and resilience [@Gunderson:2002vk]. Potential refers to the capital available to
-the system, for example nutrients and carbon captured by a forest, or human
+and resilience [@Gunderson:2002vk]. Potential refers to the capital available
+to the system, for example nutrients and carbon captured by a forest, or human
 capital - skills and knowledge accumulated to run the economy. Connectedness is
@@
 literature that has used or referenced works that trace back to Panarchy
-[@Gunderson:2002vk]. We extracted complete records from the Scopus database that
-matched the search for *"panarchy OR adaptive cycle"* (N = 595), or the search
-*"(panarchy OR adaptive cycle) AND resilience"* (N = 278). The data was combined
-with all papers (N = 1923) that cited the inaugural paper that introduced the
-book to the academic community [@Holling:2001cu]. Records with missing abstracts
-were dropped (N = 191), and records with missing year were set to 2020 given
-that they are accepted manuscripts with digital object identifier (doi)
-scheduled to be published later in 2020 or 2021. Our final sample is
+[@Gunderson:2002vk]. We extracted complete records from the Scopus database
+that matched the search for *"panarchy OR adaptive cycle"* (N = 595), or the
+search *"(panarchy OR adaptive cycle) AND resilience"* (N = 278). The data was
+combined with all papers (N = 1923) that cited the inaugural paper that
+introduced the book to the academic community [@Holling:2001cu]. Records with
+missing abstracts were dropped (N = 191), and records with missing year were set
+to 2020 given that they are accepted manuscripts with digital object identifier
+(doi) scheduled to be published later in 2020 or 2021. Our final sample is
 `r dim(dtm)[1]` documents.
@@
 data was also lemmatized using the hash lemmas dataset available at
-[\<http://www.lexiconista.com\>](http://www.lexiconista.com/). Lemmatization is
+[\<http://www.lexiconista.com>](http://www.lexiconista.com/). Lemmatization is
 the process of simplifying words to their root lemmas (e.g. *teaching* and
@@
 dimensionality of a corpus of data (typically but not necessarily text) into
-topics [@Blei:2012dk]. Here a topic is a latent variable that ranks words with
-high probability of appearing together within the same document. Documents in
-turn can be described by the probability distribution of a particular set of
+topics [@Blei:2012dk]. Here a topic is a latent variable that ranks words
+with high probability of appearing together within the same document. Documents
+in turn can be described by the probability distribution of a particular set of
 topics. Since they are (posterior) probabilities, the sum of the probability of
@@
 *Document analysis*: We complemented our unsupervised approach with document
-analysis [@Bryman:2008va] by coding an additional set of categorical variables
-for a sample of papers (N = `r nrow(qual)`). The papers read are part of the
-corpus used for topic modelling. We read papers avoiding biasing our selection
-to high cited papers, specific journals, older papers or particular disciplines
-or methods.
+analysis [@Bryman:2008va] by coding an additional set of categorical
+variables for a sample of papers (N = `r nrow(qual)`). The papers read are part
+of the corpus used for topic modelling. We read papers avoiding biasing our
+selection to high cited papers, specific journals, older papers or particular
+disciplines or methods.
 
@@
 as the methods used. We did a preliminary scoping and coding of papers, and
-through a iterative process we kept track of potential fields under reviewed by
+through an iterative process we kept track of potential fields under review by
 our corpus, or simply absent of it. We coded papers until we reached conceptual
@@
 should maximize entropy and the log-likelihood estimation while minimizing
-perplexity [@Grun:2011tb]. Gibbs sampling maximised both entropy and likelihood
-with our data when compared to other alternatives. The second best was the
-variational-expectation maximisation algorithm (VEM) when $\alpha$ was not set
-constant. $\alpha$ is a hyper parameter that weights the evenness of topic
-distribution. A lower $\alpha$ than default values indicates that the documents
-can be described by rather fewer topics, or that its distribution is very
-uneven. In fact, we observed that increasing the number of topics from 5 to 100
-topics did increase entropy at expense of reducing $\alpha$, meaning that
+perplexity [@Grun:2011tb]. Gibbs sampling maximised both entropy and
+likelihood with our data when compared to other alternatives. The second best
+was the variational-expectation maximisation algorithm (VEM) when $\alpha$ was
+not set constant. $\alpha$ is a hyper parameter that weights the evenness of
+topic distribution. A lower $\alpha$ than default values indicates that the
+documents can be described by rather fewer topics, or that its distribution is
+very uneven. In fact, we observed that increasing the number of topics from 5 to
+100 topics did increase entropy at expense of reducing $\alpha$, meaning that
 despite the larger number of topics, the main content of a document is still
@@
 problems, from delimitation of periods in archaeology and anthropology
-[@Redman:2003bd], to financial crises in Europe [@Castell:2020dg], to traffic
-jams in China [@Zeng:2020hu]. Some recent empirical tests of the adaptive cycle
-used innovative types of data and methods. For example, information transfer
-methods based on entropy have been suggested to approximate relevant components
-of a system and the empirical proxies of the axes within which the adaptive
-cycle unfolds: potential, connectedness, and resilience [@Castell:2020dg].
-Percolation methods combined with big data have been proven useful for testing
-hypotheses of regime shifts in urban systems (originally proposed in Panarchy
-[@Gunderson:2002vk]), and deriving the temporal and spatial scales at which the
-adaptive cycle emerges [@Zeng:2020hu].
+[@Redman:2003bd], systemic interactions between forest-related sectors in Sweden [@moen2010interlocking], financial crises in Europe [@Castell:2020dg], to
+traffic jams in China [@Zeng:2020hu].
+The adaptive cycle has been used as a model to structure rich empirical information about participation in adaptive management and, based on that analysis, to develop recommendations for more flexible stakeholder engagement [@stringer2006unpacking].
+The adaptive cycle has also been used to provide a more mechanistic description of the history of a variety of specific social-ecological systems, including the Galapagos Islands [@gonzalez2008rethinking], the Western Australian Agricultural Region [@allison2004resilience], the Erhai Lake watershed in southwest China [@dearing2008landscape], and the cities of Cambridge and Swansea in the UK [@simmie2010economic].
+The adaptive cycle has also been used to describe more general social and ecological processes, such as lake eutrophication [@carpenter2001metaphor] and the development of regional industrial clusters [@martin2011conceptualizing].
+
+Recent empirical tests of the
+adaptive cycle used innovative types of data and methods. For example,
+information transfer methods based on entropy have been suggested to approximate
+relevant components of a system and the empirical proxies of the axes within
+which the adaptive cycle unfolds: potential, connectedness, and resilience
+[@Castell:2020dg]. Percolation methods combined with big data have been proven
+useful for testing hypotheses of regime shifts in urban systems (originally
+proposed in Panarchy [@Gunderson:2002vk]), and deriving the temporal and
+spatial scales at which the adaptive cycle emerges [@Zeng:2020hu].
 
@@
 emergence of alternative explanations, and identify the circumstances under
-which a theory may fail [@Hansson:2017tr]. Many of the papers we read confirmed
-but did not challenge the expectations that Panarchy put forward, though a few
-of the most recent works did falsify some of the ideas proposed --e.g. the back
-loop is faster that the front one [@Castell:2020dg]. We believe more work is
-needed to better understand the circumstances where Panarchy works, and when it
-falls short. In other words, to refine the scientific framework into a
-scientific theory.
+which a theory may fail [@Hansson:2017tr]. Many of the papers we read
+confirmed but did not challenge the expectations that Panarchy put forward,
+though a few of the most recent works did falsify some of the ideas proposed
+--e.g. the back loop is faster than the front one [@Castell:2020dg]. We
+believe more work is needed to better understand the circumstances where
+Panarchy works, and when it falls short. In other words, to refine the
+scientific framework into a scientific theory.
 
@@
 intermediate disturbances (from ecology). The latter has been debunked in the
-ecological literature but might still been used within Panarchy thinking
+ecological literature but might still have been used within Panarchy thinking
 [@Fox:2013br]. A recent review of sustainability science mapped the different
 schools of thought that the discipline has developed over the last decades
-[@Clark:2020jd] and reached a similar conclusion: too many conceptual frameworks
-have been developed, but there have been too few empirical attempts to test the
-frameworks against data and falsify hypotheses. The review also emphasizes the
-problem of measurement and observation [@Clark:2020jd]. In the context of
-panarchy, recent work has developed an information theory based approach that
-enables the identification of adaptive cycles [@Castell:2020dg]. The authors
-identify phases of the adaptive cycle in the European financial crises and in
-grassland ecosystems, but fail to find support for the difference in speed
-between the forward and backward loops originally proposed in Panarchy.
+[@Clark:2020jd] and reached a similar conclusion: too many conceptual
+frameworks have been developed, but there have been too few empirical attempts
+to test the frameworks against data and falsify hypotheses. The review also
+emphasizes the problem of measurement and observation [@Clark:2020jd]. In the
+context of panarchy, recent work has developed an information theory based
+approach that enables the identification of adaptive cycles
+[@Castell:2020dg]. The authors identify phases of the adaptive cycle in the
+European financial crises and in grassland ecosystems, but fail to find support
+for the difference in speed between the forward and backward loops originally
+proposed in Panarchy.
 
@@
 Sustainability science is defined by the problems it addresses rather than the
-disciplines it employs [@Anonymous:2016fv]. Our topic modeling results helped in
-identifying some of the problems that the community is addressing and how they
-have changed over time. Some of these problems include climate change (topic
-24), urban development and disaster response (7), conservation and landscape
-management trade-offs (5), natural resource management (9), food production (3),
-policy tools and plans (16), community based management (13) or knowledge and
-local social practices (25. We also observed topics closely related to methods
-or approaches to study social-ecological systems such as modeling (19,21),
-networks (6), or ecosystem services (15). Topic modeling can be used to further
-investigate particular academic communities, identify key landmark papers in
-certain topics, or identify the network of key collaborators and experts.
+disciplines it employs [@Anonymous:2016fv]. Our topic modeling results helped
+in identifying some of the problems that the community is addressing and how
+they have changed over time. Some of these problems include climate change
+(topic 24), urban development and disaster response (7), conservation and
+landscape management trade-offs (5), natural resource management (9), food
+production (3), policy tools and plans (16), community based management (13) or
+knowledge and local social practices (25). We also observed topics closely
+related to methods or approaches to study social-ecological systems such as
+modeling (19,21), networks (6), or ecosystem services (15). Topic modeling can
+be used to further investigate particular academic communities, identify key
+landmark papers in certain topics, or identify the network of key collaborators
+and experts.
 
@@
 analysis, the qualitative insights could be biased towards a non-random
-selection of papers that aligned our research interests or highly cited review
+selection of papers that aligned with our research interests or highly cited review
 papers. Topic modeling enabled us to complement the analysis and include all
@@
 particular work but limited generalizations. We also, for example, ignored rich
-areas of social science research (e.g. Anthropology) that is published in books
+areas of social science research (e.g. Anthropology) often published in books
 and book chapters, that are not archived by scientific search engines.
@@
 impacts of ecosystem services from regime shifts in social-ecological systems
-[@Rocha:2015tk]. Replication attempts do however face the challenge of accessing
-full text when many of the papers are behind pay-walls.
+[@Rocha:2015tk]. Replication attempts do however face the challenge of
+accessing full text when many of the papers are behind pay-walls.
 

### 2022-03-07 paper/paper_panarchy.Rmd
implementing suggestions from reviewers and co-authors

@@
 author: |
-    | \small Juan Rocha^1,2,3^, Linda Luvuno^4^, Jesse Rieb^5^, Erin Crockett^6^,  Katja Malmborg^1^, Michael Schoon^7^, Garry Peterson^1^
+    | \small Juan C. Rocha^1,2,3^, Linda Luvuno^4^, Jesse Rieb^5^, Erin Crockett^6^,  Katja Malmborg^1^, Michael Schoon^7^, Garry D. Peterson^1^
     | \small
@@
   How do social-ecological systems change over time?
-In 2002 Holling and colleagues proposed the concept of Panarchy, which presented social-ecological systems as an interacting set of adaptive cycles, each of which is produced by the dynamic tensions between novelty and efficiency at multiple scales.
+In 2002 Holling and colleagues proposed the concept of panarchy, which presented social-ecological systems as an interacting set of adaptive cycles, each of which is produced by the dynamic tensions between novelty and efficiency at multiple scales.
 Initially introduced as a conceptual framework and set of metaphors, panarchy has gained the attention of scholars across many disciplines, and its ideas continue to inspire further conceptual developments.
@@
 output:
+  rtf_document:
+    toc: no
   pdf_document:
@@
 Resilience Alliance understood change in social-ecological systems. The concept
-of Panarchy was a key focus of this influential book. In the late 1990s and
+of panarchy was a key focus of this influential book. In the late 1990s and
 early 2000s, the Resilience Alliance was a productive, innovative, and highly
@@
 and natural sciences, arts, and humanities, as well as by bridging theory and
-practice [@Parker:2012eo].
-
-The Panarchy concept builds on Holling's adaptive cycle [@biosphere:hJ4c_Ua8]
-by extending the idea across spatial and temporal scales. It remains a boundary
+practice [@Parker:2012eo].Panarchy remains a boundary
 object that has inspired research topics, enabled collaborations and nurtured
-new scientific frameworks. The ideas put forward have been applied in field
+new scientific frameworks [@Parker:2012eo]. The ideas put forward have been applied in field
 studies, archaeology, mathematical models, participatory work, and scenario
-development. Panarchy has inspired resilience assessments and guided decision
+development [@gunderson2022applied]. Panarchy has inspired resilience assessments and guided decision
 making. In this article we pay tribute to the book by studying how the concepts
@@
 
-Panarchy proposes that it is useful to conceptualize systems in terms of
+The panarchy concept builds on Holling's adaptive cycle [@biosphere:hJ4c_Ua8]
+by extending the idea across spatial and temporal scales. Panarchy proposes that it is useful to conceptualize systems in terms of
 interacting adaptive cycles. The adaptive cycle was an idea first proposed by
@@
 
-Panarchy is not a theory of what it is, but a metaphor of what might be
-[@Gunderson:2002vk]. It is not a predictive tool, but aims to understand
-adaptive change. The adaptive cycle is hypothesized to exist within a
+"_Panarchy is not a theory of what it is, but a metaphor of what might be_"
+[@Gunderson:2002vk 32]. It is not a predictive tool, but aims to understand
+adaptive change. The adaptive cycle is proposed to exist within a
 three-dimensional space defined by three properties: potential, connectedness,
@@
 \centering
-\includegraphics[width = 6in, height = 2.5in]{fig1_panarchy}
+\includegraphics[width = 6in, height = 2in]{fig1_panarchy}
 \caption{\textbf{Panarchy} is an heuristic of nested adaptive cycles that serves to represent a variety of systems and environmental problems.
-Adaptive cycles (A) at different scales of the hierarchy (B) can be connected through remember and revolt cross-scale interactions.}
+Adaptive cycles (A) at different scales of the hierarchy (B) can be connected through remember and revolt cross-scale interactions.
+(C) relates the adaptive cycle to the potential, connectedness and resilience axis, reproduced from Panarchy edited by Lance Gunderson and C.S.
+Holling, copyright \textcopyright 2002 Island Press.
+Reproduced with permission of Island Press, Washington, DC.}
 \label{fig:fig1}
@@
 collapse, and recovery that are common to populations, ecological communities,
-markets, or political organizations. While each subsystem in the hierarchy can
-be at a different phase of the adaptive cycle, these subsystems can influence
-each other through cross-scale interactions called *"revolt"* and *"remember"*
+markets, or political organizations.
+While each subsystem in the hierarchy can be at a different phase of the adaptive cycle, these subsystems can particularly influence each other through cross-scale interactions called *“revolt”*, where release in a smaller, faster sub-system can trigger release in a larger, slower sub-system,  and *“remember”*, in which the structure of a larger-slower sub-system shapes the dynamics of reorganization of a smaller, faster sub-system
 [@Gunderson:2002vk].
 
-The three-dimensional space has corners with attractors that can cause the
-adaptive cycle to stagnate: poverty and rigidity traps [@Gunderson:2002vk].
+The three-dimensional space defined by potential, connectedness and resilience, has corners with attractors that can derail the
+adaptive cycle: poverty and rigidity traps [@Gunderson:2002vk].
 Poverty traps are described in panarchy as maladaptive states where potential,
@@
 
-Panarchy offers a rich conceptual framework for understanding environmental
+Panarchy offers a rich conceptual framework for interpreting environmental
 problems. While inspired by several mathematical constructs (e.g. cycles, traps,
@@
 research questions and applied problems. As such, it is useful as a boundary
-object and can be used empirically or metaphorically. The book itself presents a
+object and can be used empirically or metaphorically. The book presents a
 series of case studies where geographers, economists, political scientists, and
@@
 
-However, the extent to which these concepts have gained empirical support beyond
-the metaphor, or how they have evolved over time, remain open questions. For
-example, the adaptive cycle is a useful metaphor to look back at history and
-organize events and periods around phases, but it remains challenging to
-identify the phase of a system in real time, or to use panarchy to project
-future trajectories. Scientific theories are bodies of knowledge that answer the
+Scientific theories are bodies of knowledge that answer the
 question why? The answer implies causal reasoning. Regardless of discipline, a
@@
 confirmation bias and testing when explanations are valid representations of
-reality. For panarchy to become a useful theory in sustainability science, it
-needs to stand the rigor of testing, and being able to explain phenomena in
-social-ecological systems across different contexts. We identify progress and
-knowledge gaps in the scientific practice inspired by panarchy.
-
+reality.
+In this paper we explore how the conceptual framework of panarchy has been applied, developed and tested over the past twenty years.
+For panarchy to become a useful theory in sustainability science, it
+needs to be able to explain phenomena in
+social-ecological systems across different contexts, in a way tah tis testable, and withstands the rigour of those tests.
 Here we review the academic literature of the last two decades to trace how the
-ideas proposed in Panarchy have evolved, to what type of problems they have been
-applied and found useful, and finally, what key research frontiers remain.
+ideas proposed in panarchy have evolved. We conclude by suggesting research frontiers for future work.
 
@@
 *Data*: We used the Web of Science, Scopus, and GoogleScholar to survey academic
-literature that has used or referenced works that trace back to Panarchy
+literature that has used or referenced works that trace back to panarchy
 [@Gunderson:2002vk]. We extracted complete records from the Scopus database
@@
 
-Based on the reading of Panarchy, we first developed a series of categorical
+Based on the reading of panarchy, we first developed a series of categorical
 variables that captured the main components of the scientific framework. The
@@
 distinguished empirical cases from empirical constructs. An empirical case is
-when a study is trying to test an idea or proposition put forward by Panarchy
+when a study is trying to test an idea or proposition put forward by panarchy
 (e.g. a cross-scale interaction, or that the back loop is faster than the forth
-loop). An empirical construct is when Panarchy is used as inspiration to
+loop). An empirical construct is when panarchy is used as inspiration to
 interpret a case study, typically with a historical component to it (e.g.
 identify the phases of adaptive cycle); but it does not find support for or
-against a Panarchy preposition. That is, there was no mechanism in place to
+against a panarchy preposition. That is, there was no mechanism in place to
 avoid confirmation bias.
@@
 through an iterative process we kept track of potential fields under review by
-our corpus, or simply absent of it. We coded papers until we reached conceptual
-saturation, that is, additional papers were not proving additional categories of
-the already coded. The code book is available at
+our corpus, or simply absent from it. We coded papers until we reached conceptual
+saturation, that is, additional papers were not proving additional categories. The code book is available at
 [doi.org/10.6084/m9.figshare.13490919](http://doi.org/10.6084/m9.figshare.13490919).
@@
 
-Panarchy, the book published in 2002, has been cited over 7200 times according
-to Google Scholar. The scientific paper that introduced the book to the
+Panarchy, the book published in 2002, has been cited over 8600 times according
+to Google Scholar (by March 2022). The scientific paper that introduced the book to the
 scientific community [@Holling:2001cu] has received, at the time of writing,
-1715 citations in the Web of Science, and 1923 in Scopus. Roughly half of the
+1853 citations in the Web of Science, and 2197 in Scopus. Roughly half of the
 citations have come from environmental (28%) and social sciences (22%). Computer
@@
 ```
-Comparing algorithm performance revealed that Gibbs sampling was the best fit
+
+Topic modelling was done using Gibbs sampling. In our comparison of alternative Latent Dirichlet Allocation algorithms Gibbs sampling provided the best fit
 for the data (Fig \ref{fig:algorithm}). As a rule of thumb, an ideal method
@@
 list of words and probability distributions listed under a topic with a
-different number. That is because the labeling of topics is random and we do not
-want to bias the reader to over interpret the word patterns found.
+different number. That is because the labeling of topics is random and we urge readers not to over interpret the word patterns found.
 
@@
 The human-coded document analysis revealed that the most common feature of
-panarchy in the literature was the adaptive cycle (81%, N = `r nrow(qual)`)
+panarchy in the literature was the adaptive cycle (81%, 34 out of N = `r nrow(qual)`)
 followed by cross-scale interactions (Fig \ref{fig:qual}). Poverty and rigidity
 traps were less studied features in our sample (14% and 19% respectively), even
-when accounting for slightly different terminology such as "lock-ins". 52.4% of
+when accounting for slightly different terminology such as "lock-in". About half (52.4% or 22/42) of
 the papers analyzed were conceptual papers, many without defined methods or
-discernible research question. Roughly half (47%) of the papers were conceptual
-or only used panarchy as a metaphor. Over 40% of papers in our sample went a
+a clear research question. Roughly half (20/42) of the papers used panarchy as a metaphor. About 40% of papers in our sample (17/42) went a
 step further and used panarchy as an empirical construct, for example by
-attempting to identify the phases of the adaptive cycle (61%). Out of the 22
+attempting to identify the phases of the adaptive cycle (26/42). Off the 22
 empirical cases, 6 were at the time scale of centuries, 7 on decades, 6 on
 years, and 1 in weeks. Spatially, 4 were at a city scale, 17 regional, and 1
-national. Empirical papers predominantly used qualitative methods (77%, N = 22)
-and were generally retrospective historical reviews.
+national. Empirical papers predominantly used qualitative methods (17)
+and were primarily retrospective historical reviews.
 
-Several of the papers reviewed (81%) attempted to identify the adaptive cycle
-either through qualitative or quantitative methods. Identification of adaptive
+Most of the papers reviewed (81%, 34 of N = 42) identified adaptive cycles
+either through scientific methods, or as analogy or framework to make sense of a narrative. Identification of adaptive
 cycles has found applications in a wide range of disciplines and research
@@
 [@Castell:2020dg]. Percolation methods combined with big data have been proven
-useful for testing hypotheses of regime shifts in urban systems (originally
-proposed in Panarchy [@Gunderson:2002vk]), and deriving the temporal and
+useful for testing hypotheses of regime shifts in urban systems, an idea originally
+proposed in panarchy [@Gunderson:2002vk]; and deriving the temporal and
 spatial scales at which the adaptive cycle emerges [@Zeng:2020hu].
@@
 
-Given the strong predominance of conceptual papers, it is not surprising that
-one of the key frontiers of research is to empirically ground such
-conceptualizations in a way that one can test theoretical expectations beyond
-metaphors. Metaphors have played an important role as boundary objects enabling
-interdisciplinary dialogues. We observed this on the wide variety of
-applications that panarchy has inspired, in fields from anthropology to
-engineering. In the empirical realm, the adaptive cycle is the dimension of
-panarchy with which researchers engage the most, but it is dominated by
-qualitative and retrospective studies. That means, we still lack the theory and
-methods to gather observations and identify a system's current adaptive cycle
-phase, or the probability of it transitioning to a new phase.
-
-The high level of conceptualization but lack of empirically grounded testing of
-panarchy theory came across as one of the major limitations, given our sample of
-papers reviewed (N=`r nrow(qual)`). What makes a theory a scientific theory is
-that the plurality of scientific approaches (e.g. positivism, constructivism)
-should have mechanisms in place to avoid confirmation bias, allowing the
-emergence of alternative explanations, and identify the circumstances under
-which a theory may fail [@Hansson:2017tr]. Many of the papers we read
-confirmed but did not challenge the expectations that Panarchy put forward,
-though a few of the most recent works did falsify some of the ideas proposed
---e.g. the back loop is faster than the front one [@Castell:2020dg]. We
-believe more work is needed to better understand the circumstances where
-Panarchy works, and when it falls short. In other words, to refine the
-scientific framework into a scientific theory.
-
-The criticism of over-conceptualization is not unique to the literature that
-engages with Panarchy ideas. Here by over-conceptualization we refer to the
-problem of scientific jargon, where different communities use slightly different
-wordings to refer to the same phenomenon; reducing the chances that lessons from
-a discipline percolates to another. An example from Panarchy is the endorsement
-of ideas such as creative destruction (from economics) or the theory of
-intermediate disturbances (from ecology). The latter has been debunked in the
-ecological literature but might still have been used within Panarchy thinking
-[@Fox:2013br]. A recent review of sustainability science mapped the different
-schools of thought that the discipline has developed over the last decades
-[@Clark:2020jd] and reached a similar conclusion: too many conceptual
-frameworks have been developed, but there have been too few empirical attempts
-to test the frameworks against data and falsify hypotheses. The review also
-emphasizes the problem of measurement and observation [@Clark:2020jd]. In the
-context of panarchy, recent work has developed an information theory based
-approach that enables the identification of adaptive cycles
-[@Castell:2020dg]. The authors identify phases of the adaptive cycle in the
-European financial crises and in grassland ecosystems, but fail to find support
-for the difference in speed between the forward and backward loops originally
-proposed in Panarchy.
-
-The panarchy dimensions that received less attention in our qualitative analysis
-included creative destruction, rigidity and poverty traps. This may be at least
-partly due to the scope of our data: papers that have cited Panarchy or
-Holling's paper [@Gunderson:2002vk; @Holling:2001cu], and a relatively small
-sample size (N=`r nrow(qual)`). Concepts such as creative destruction and
-poverty traps precede Panarchy and therefore have been theorized and empirically
-grounded outside the panarchy stream of thinking. For example, the theory of
-poverty traps dates back to the 1950's in economics, and has received both
-theoretical development [@Bowles:2006hx] as well as empirical grounding
-[@Banerjee:2015ja; @Banerjee:2012tq] that has enabled researchers and
-governments to distinguish what kind of interventions are likely to reduce
-poverty. Here we encourage similar empirical efforts in formalizing theory and
-empirical support for the adaptive cycles, their cross-scale interactions
-(remember and revolt), or how the hierarchical nested nature of complex systems
-enhance or erode resilience in social-ecological systems.
+We examined how the conceptual framework of panarchy has been developed and tested over the past twenty years.
+We did this using a topic modelling to analyze a large number of papers and a focused review of a selection of panarchy papers.
+We find that panarchy has been used an interpretative tool across a wide variety of topics, but has there has been little effort to substantially operationalize or test panarchy.
 
 Sustainability science is defined by the problems it addresses rather than the
-disciplines it employs [@Anonymous:2016fv]. Our topic modeling results helped
+disciplines it employs [@Anonymous:2016fv].
+This goal of this paper was to explore how the ideas proposed in panarchy have evolved, to what type of problems they have been applied and found useful, and finally, what key research frontiers remain.
+Our topic modeling results helped
 in identifying some of the problems that the community is addressing and how
@@
 This paper aimed to study the use of panarchy related concepts and their
-evolution since the publication of the book. Systematic literature reviews are
+evolution since the publication of the book [@Gunderson:2002vk]. Systematic literature reviews are
 often restricted to a limited subset of the literature -- based on what is
@@
 reproducibility and reduces sample bias, but offers limited insights into the
-dimensions of panarchy used, methods, or criticisms. Here we show that both
-methods combined are a powerful combination.
+dimensions of panarchy used, methods, or criticisms. By combining methods, as we did in this paper, we believe literature reviews can become more powerful.
 
 Our choice of methods does however have some limitations. We could not directly
-compare the two methods since they do not share a common categorical axis on
-which they can be measured back to back. The topics gave us a general but
-shallow overview, while the document analysis provide deep insights on
-particular work but limited generalizations. We also, for example, ignored rich
+compare the two methods since they do not share a common categorical axis that we could use to translate between methods.
+Topic modelling contributes understanding of what topics --sets of words-- panarchy has been associated with; while our document analysis contribute in-depth understanding on how people have interpreted and use these concepts.
+The topics gave us a broad but
+shallow overview, while the document analysis provides deep insights on
+particular work but limited generalizations. Our method also ignored rich
 areas of social science research (e.g. Anthropology) often published in books
 and book chapters, that are not archived by scientific search engines.
-Additionally, a more inductive approach to the qualitative coding could have
+Finally, an iterative inductive approach to the qualitative coding could have
 given us different insights on the literature outside the pre-defined categories
@@
 
-Future studies could benefit from including gray literature such as theses,
+Future studies of panarchy, or other concepts, can benefit from our experience.
+Panarchy is a collection of concepts, some of them more used and developed than others such as the adaptive cycle.
+Future reviews would benefit from including gray literature such as theses,
 books, non-governmental organizations reports, government agencies reports, or
-non-English literature. This would let us see the usage of Panarchy concepts in
-more applied settings such as management and decision making. Previous studies
-have also shown that the use of full text instead of abstracts can also offer
+non-English literature. This would let us see the usage of panarchy concepts in
+more applied settings such as management and decision making. The use of full text instead of abstracts can also offer
 additional insights on the automated analyses, for example in attributing
 impacts of ecosystem services from regime shifts in social-ecological systems
-[@Rocha:2015tk]. Replication attempts do however face the challenge of
-accessing full text when many of the papers are behind pay-walls.
+[@Rocha:2015tk].
+While databases are improving, full-text analysis are only possible for open-access journals, and text analysis is difficult across multiple languages.
+Databases are increasing their coverage of books and reports, which suggest our approach may be able to access a broader set of literature in the future.
+
+Our review of panarchy papers (N = `r nrow(qual)`) revealed a high level of conceptualization but lack of theory development or testing.
+Conceptual frameworks, such as panarchy, have an important role as boundary objects that enable interdisciplinary dialogues.
+The usefulness of panarchy in this role is revealed by how it has been used to address a variety of issues in fields ranging from anthropology to engineering.
+Many of papers reviewed used the adaptive cycle to interpret past events, but did not engage in measurement or observation to test adaptive cycles in the present or potential future trajectories, nor other features of panarchy such as poverty or rigidity traps, or revolt and remember cross-scale interactions.
+There remains a lack of testable mechanistic theories to explain the dynamics of panarchy, and consequently there is a lack of methods to measure or observe it in real social-ecological systems.
+
+The high level of conceptualization but lack of theory testing came across as one of the major limitations, given our sample of
+papers reviewed (N=`r nrow(qual)`). What makes a theory a scientific theory is
+that the plurality of scientific approaches (e.g. positivism, constructivism)
+should have mechanisms in place to avoid confirmation bias, allowing the
+emergence of alternative explanations, and identify the circumstances under
+which a theory may fail [@Hansson:2017tr]. Many of the papers we read
+confirmed but did not challenge the expectations that panarchy put forward,
+though a few of the most recent works did falsify some of the ideas proposed
+--e.g. the back loop is faster than the front one [@Castell:2020dg]. We
+believe more work is needed to better understand the circumstances where
+panarchy works, and when it falls short. In other words, to refine the
+scientific framework into a scientific theory.
+
+The criticism of over-conceptualization is not unique to the literature that
+engages with panarchy ideas. Here by over-conceptualization we refer to the
+problem of scientific jargon, where different communities use slightly different
+wordings to refer to the same phenomenon; reducing the chances that lessons from
+a discipline percolates to another. An example from panarchy is the endorsement
+of ideas such as creative destruction (from economics) or the theory of
+intermediate disturbances (from ecology). The latter has been debunked in the
+ecological literature
+[@Fox:2013br]. A recent review of sustainability science mapped the different
+schools of thought that the discipline has developed over the last decades
+[@Clark:2020jd] and reached a similar conclusion: too many conceptual
+frameworks have been developed, but there have been too few empirical attempts
+to test the frameworks against data and falsify hypotheses. The review also
+emphasizes the problem of measurement and observation [@Clark:2020jd]. In the
+context of panarchy, recent work has developed an information theory based
+approach that enables the identification of adaptive cycles
+[@Castell:2020dg]. The authors identify phases of the adaptive cycle in the
+European financial crises and in grassland ecosystems, but fail to find support
+for the difference in speed between the forward and backward loops originally
+proposed in panarchy. Lack of support for the same ideas was also found in studies of urban traffic [@Zeng:2020hu].
+
+The panarchy dimensions that received less attention in our qualitative analysis
+included creative destruction, rigidity and poverty traps. This may be at least
+partly due to the scope of our data: papers that have cited panarchy or
+Holling's paper [@Gunderson:2002vk; @Holling:2001cu], and a relatively small
+sample size (N=`r nrow(qual)`). Concepts such as creative destruction and
+poverty traps precede panarchy and therefore have been theorized and empirically
+grounded outside the panarchy stream of thinking. For example, the theory of
+poverty traps dates back to the 1950's in economics, and has received both
+theoretical development [@Bowles:2006hx] as well as empirical grounding
+[@Banerjee:2015ja; @Banerjee:2012tq] that has enabled researchers and
+governments to distinguish what kind of interventions are likely to reduce
+poverty.  As researchers have done with poverty traps, we believe it would be fruitful to propose and evaluate potential mechanisms that could produce adaptive cycles, their cross-scale interactions
+(remember and revolt), or how the hierarchical nested nature of complex systems
+enhance or erode resilience in social-ecological systems.
 
@@
 
-We reviewed the academic literature of the last two decades to trace how
-Panarchy related concepts have evolved, in what type of problems they have been
-applied and found useful, and what remains as the key frontiers of research.
-Despite a growing body of literature, no topic seems to dominate the academic
-production of scholars using these concepts. The feature most used is the
-adaptive cycle. It have been found most useful in studies with a historical
-perspective, both in short time studies in natural resource management, urban
-development, and conservation; as well as archeological and anthropological
-studies that span centuries to millennia. Hierarchies in scale or maladaptive
-traps have received comparatively less attention. The frontiers of research
-point to the problems of measurement and prediction. For example, it is
-difficult to observe and approximate resilience, connectedness and potential in
-the present. How to determine the likelihood that a social-ecological system
-will transition to a different phase of the adaptive cycle, or experience a
-"memory" or "revolt" type of cross-scale interaction are still open questions.
-While our survey of the literature identified a few promising pathways towards
-empirically grounding panarchy concepts, we believe the problem of measurement
-and prediction remain fertile ground for future research efforts.
+In this paper we explore how the conceptual framework of panarchy has been used, developed and tested over the past twenty years.
+Despite a growing body of literature, no topic appears to dominate papers using panarchy concepts.
+The adaptive cycle is the most widely used concept, and it has been most used to interpret histories that range from the dynamics of natural resource management, urban development, to archeological and anthropological studies that span millennia.
+Cross-scale dynamics and structure and traps have received less attention.
+Operationalzing panarchy theory remains a challenge.
+How to measure and forecast system behaviour in ways that is useful for panarchy concepts remains an open question.
+Addressing these issues would be very useful by developing tools to answer critical questions about the dynamics of resilience across linked social-ecological systems.
+For example, how do current changes in the global food system alter its resilience, connectedness and potential.
+How do identify whether a social-ecological system is entering a different phase of the adaptive cycle?
+What type of ‘remember’ or ‘revolt’ dynamics are shaping the current food systems of Stockholm or Bogota?
+While our survey of the literature identified a few promising pathways, developing ways to operationalize theories that explain the dynamics of resilience is vitally needed to confront the challenges of creating a sustainable, just, world.
 
@@
 the Stockholm Resilience Centre, and Stockholm University. We acknowledge the
-feedback of the Resilience Alliance and encouragement in pursuing this project.
+feedback of the Resilience Alliance and encouragement in pursuing this project. GDP acknowledges the support of the Rasmussen foundation.
 

### 2022-03-07 03-figures.R

@@ -5,7 +5,7 @@ library (topicmodels)
 library(lda)
 library(patchwork)
 library(png)
-
+library(gg3D)
 
 load("data/models_gibbs.RData")
 
@@ -261,7 +261,7 @@ qual_summary %>%
     geom_col() + 
     facet_wrap(~var_type, scales = "free_y") +
     scale_y_continuous(labels = scales::percent) +
-    scale_x_reordered() +
+    tidytext::scale_x_reordered() +
     coord_flip() + labs(x = "", y = "") +
     theme_light(base_size = 6)
 
@@ -273,24 +273,81 @@ save(qual, file = "data/qual_coding.RData")
 
 p1 <- ggplot(data = data_frame(x = 0, y= 0), aes(x,y)) + geom_blank() +
     labs(tag = "A") +
-    theme_void() +
+    theme_void(base_size = 6) +
     annotation_custom(
         grob = grid::rasterGrob( image = readPNG(
-            "paper/figures/adaptive_cycle.png"), interpolate = TRUE),
+            "paper/extra_figs/adaptive_cycle.png"), interpolate = TRUE),
         xmin = -Inf, xmax = Inf, ymin = -Inf, ymax = Inf
     )
 p2 <- ggplot(data = data_frame(x = 0, y= 0), aes(x,y)) + geom_blank() +
     labs(tag = "B") +
-    theme_void() +
+    theme_void(base_size = 6) +
     annotation_custom(
         grob = grid::rasterGrob( image = readPNG(
-            "paper/figures/panarchy.png"), interpolate = TRUE),
+            "paper/extra_figs/panarchy.png"), interpolate = TRUE),
         xmin = -Inf, xmax = Inf, ymin = -Inf, ymax = Inf
     )
 
-p1 | p2 + plot_layout(widths = 2.5, heights = 2.5)
 
-ggsave(filename = "paper/figures/fig1.png",
-       plot = last_plot(), width = 6, height = 2.5, device = "png", dpi = 400)
+p3 <- ggplot(data = data_frame(x = 0, y= 0), aes(x,y)) + geom_blank() +
+    labs(tag = "C") +
+    theme_void(base_size = 6) +
+    annotation_custom(
+        grob = grid::rasterGrob( image = readPNG(
+            "paper/extra_figs/Figure_2-2.png"), interpolate = TRUE),
[…]
